### 2024-11-21 analysis_plan.qmd
Analysis plan for pre-registration added.

---
title: "Data Synthesis Plan (Meta-Analysis)"
format: pdf
bibliography: 
  mylibrary.bib
csl: apa.csl
---

Statistical analysis of the data extracted will be performed in R, (v 4.3.3; R Core Team, https://www.r-project.org/) and RStudio (v 2023.06.1; Posit, https://posit.co/). All code utilised for data preparation and analyses are available in the corresponding GitHub repository <https://github.com/jamessteeleii/pcos_ree_meta>. We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: [ADD LINKS](ADD LINK).

The proposed analysis plan for data synthesis is part of the pre-registration on PROSPERO (update with registration number once complete). Given our research question our analysis aimed at parameter estimation [@cummingNewStatisticsWhy2014] within a Bayesian meta-analytic framework [@kruschkeBayesianNewStatistics2018]. For all analyses model parameter estimates and their precision, along with conclusions based upon them, will be interpreted continuously and probabilistically, considering data quality, plausibility of effect, and previous literature, all within the context of each model. The `renv` package [@usheyRenvProjectEnvironments2023] is used for package version reproducibility and a function based analysis pipeline using the `targets` package [@landauTargetsDynamicFunctionOriented2023] is employed (the analysis pipeline can be viewed by downloading the R Project and running the function `targets::tar_visnetwork()`). Effect sizes and their variances are all calculated using the `metafor` packages `escalc()` function [@viechtbauerMetaforMetaAnalysisPackage2023]. The main package `brms` [@burknerBrmsBayesianRegression2023] is used in fitting all the Bayesian meta-analysis models. Prior and posterior draws are taken using `tidybayes` [@kayTidybayesTidyData2023] and `marginaleffects` [@arel-bundockMarginaleffectsPredictionsComparisons2023] packages. All visualisations are created using `ggplot2` [@wickhamGgplot2CreateElegant2023], `tidybayes`, and the `patchwork` [@pedersenPatchworkComposerPlots2023] packages. Where data to be extracted from included studies is reported in plots only we will use the `juicr` package to extract this data [@lajeunesseJuicrAutomatedManual2021] and the reproducible reports for this will be found in the github repository.

As noted, we adopt a Bayesian approach to the present meta-analysis and examine both contrast-based models of studies with pairwise effects (i.e., relative contrasts between non-PCOS and PCOS arms), in addition to arm-based models. We adopt an arm-based multiple condition comparison (i.e., network) type model given that we know there will studies of women with PCOS both with, and without, a non-PCOS control arm [@hongBayesianMissingData2016]. In typical contrast-based meta-analyses data is limited to the effect sizes for paired contrasts between arms and thus studies that include both arms (i.e., relative effects between non-PCOS control vs PCOS arms); however, in arm-based analyses the data are the absolute effects within each arm and information is borrowed across studies to enable both within condition absolute, and between condition relative contrasts to be estimated. From this we focus on reporting the between condition relative contrast for non-PCOS control vs PCOS arms. The arm-based models, which are specified along with their priors in the `R/functions.R` file in `brms` syntax and explantory comments, will be our primary model and contrast-based models with weakly regularising default priors will also be presented in order to examine the sensitivity of findings to the inclusion of indirect evidence from single arm studies and prior information on overall inferences. The arm-based models are parameterised such that priors can be set informatively on the non-PCOS arms as the model intercept (and its random effect variance) and these are based on estimates from previous large scale studies (see comments in `R/functions.R`) of the resting energy expenditure of non-PCOS women, and the coefficient reflecting the relative contrast between non-PCOS controls and PCOS arms was set to be weakly informative based on the minimum and maximum ranges of resting energy expenditure as reflecting the possible range of differences that could in principle exist between two conditions. The priors for the variance effects (noted below) were similarly set based on these previous studies.

We will examine both raw mean effects (i.e., for the arm-based model the absolute mean resting energy expenditure in kcals per day for each arm, and for the contrast-based model the absolute mean difference in resting energy expenditure in kcals per day between non-PCOS controls and PCOS arms) in addition to the variance effects (i.e., for the arm-based model the absolute standard deviation in resting energy expenditure in kcals per day for each arm, and for the contrast-based model the absolute difference in standard deviation in resting energy expenditure in kcals per day between non-PCOS controls and PCOS arms). For the variance effects the arm-based models will model the natural log transformation of the standard deviation as the effect outcome and include the natural log transformation of the mean (modelled with measurement error based on the natural log transformation of the standard error of the mean for each arm) to adjust for mean-variance relationships, and for the contrast based model will utilise the log coefficient of variation effect size statistic which also adjusts for mean-variance relationships.

For all models we will examine trace plots along with $\hat{R}$ values to examine whether chains have converged, and posterior predictive checks for each model to understand the model implied distributions. From each model we will obtain draws from the posterior distributions for the conditional absolute estimates for each condition (i.e., non-PCOS controls and PCOS arms) by study, the global grand mean absolute estimates for each condition, and the between condition relative contrast for non-PCOS controls vs PCOS arms in order to present probability density functions visually, and also to calculate mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for each estimate. These will give us the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles.

# References

### 2024-11-21 R/functions.R

@@ -1,0 +1,307 @@
+# Functions to create initial example data using studies identified in https://macrofactorapp.com/pcos-bmr/ 
+# This data will be used for setting up the analysis pipelines and will be replaced by the final extracted data after systematic review
+  create_pairwise_data <- function() {
+    
+    # First we create a tibble with the data from the pairwise studies of both PCOS and controls (CONT)
+    dat <- tibble(
+      study = c("Segal, 1990",
+                "Robinson, 1992",
+                "Cosar, 2008",
+                "Georgopoulos, 2009",
+                "Graff, 2013",
+                "Larsson, 2015",
+                "Doh, 2016"),
+      n_PCOS = c(10,14,31,91,61,72,14),
+      n_CONT = c(11,14,29,48,44,30,10),
+      m_PCOS = c(1508,1624,1167,1446,1469,1411,1272),
+      m_CONT = c(1558,1633,1046,1841,1453,1325,1240),
+      s_PCOS = c(168,137,371,725,227,229,167),
+      s_CONT = c(186,215,296,305,249,193,216)
+    )
+    return(dat)
+  }
+  
+  calculate_pairwise_mean_effects <- function(dat) {
+    dat <- escalc(measure = "MD",
+                  m1i = m_PCOS,
+                  m2i = m_CONT,
+                  sd1i = s_PCOS,
+                  sd2i = s_CONT,
+                  n1i = n_PCOS,
+                  n2i = n_CONT,
+                  data = dat)
+    
+    return(dat)
+  }
+  
+  calculate_pairwise_variance_effects <- function(dat) {
+    dat <- escalc(measure = "CVR",
+                  m1i = m_PCOS,
+                  m2i = m_CONT,
+                  sd1i = s_PCOS,
+                  sd2i = s_CONT,
+                  n1i = n_PCOS,
+                  n2i = n_CONT,
+                  data = dat)
+    
+    return(dat)
+  }
+  
+  create_arm_data <- function(dat) {
+    
+    # First we pivot the pairwise data to long form (and add the se column)
+    dat2 <- dat |>
+      pivot_longer(
+        2:7,
+        cols_vary = "slowest",
+        names_to = c(".value", "cond"),
+        names_pattern = "(..)(....)"
+      ) |>
+      rename(
+        n = "n_",
+        m = "m_",
+        sd = "s_"
+      ) |>
+      mutate(
+        se = NA
+      )
+    
[…]

### 2024-11-21 R/quick_look.R

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(metafor)
+library(brms)
 
 dat <- tibble(
   study = c("Segal, 1990",
@@ -10,25 +11,34 @@ dat <- tibble(
             "Larsson, 2015",
             "Doh, 2016"),
   n_PCOS = c(10,14,31,91,61,72,14),
-  n_CON = c(11,14,29,48,44,30,10),
+  n_CONT = c(11,14,29,48,44,30,10),
   m_PCOS = c(1508,1624,1167,1446,1469,1411,1272),
-  m_CON = c(1558,1633,1046,1841,1453,1325,1240),
+  m_CONT = c(1558,1633,1046,1841,1453,1325,1240),
   s_PCOS = c(168,137,371,725,227,229,167),
-  s_CON = c(186,215,296,305,249,193,216)
+  s_CONT = c(186,215,296,305,249,193,216)
 )
 
 dat <- escalc(measure = "MD",
               m1i = m_PCOS,
-              m2i = m_CON,
+              m2i = m_CONT,
               sd1i = s_PCOS,
-              sd2i = s_CON,
+              sd2i = s_CONT,
               n1i = n_PCOS,
-              n2i = n_CON,
+              n2i = n_CONT,
               data = dat)
 
-paired_model <- rma(yi, vi, data = dat)
+# fit brms model
 
-forest(paired_model, slab = study)
+paired_model <- brm(yi | se(sqrt(vi)) ~ 1 + (1 | study),
+                    data = dat,
+                    chains = 4,
+                    cores = 4,
+                    seed = 1988,
+                    warmup = 2000,
+                    iter = 8000)
+
+plot(paired_model)
+pp_check(paired_model)
 
 
 # dat_var <- escalc(measure = "CVR",
@@ -48,7 +58,7 @@ dat2 <- dat |>
     2:7,
     cols_vary = "slowest",
     names_to = c(".value", "cond"),
-    names_pattern = "(..)(...)"
+    names_pattern = "(..)(....)"
   ) |>
   rename(
     n = "n_",
@@ -56,10 +66,6 @@ dat2 <- dat |>
     sd = "s_"
     ) |>
   mutate(
-    cond = case_when(
-      cond == "PCO" ~ "PCOS",
-      .default = cond
-    ),
     se = NA
   )
   
@@ -75,8 +81,8 @@ dat_arm <- bind_rows(
 ) |>
   mutate(
[…]

### 2024-11-21 _targets.R

@@ -5,11 +5,17 @@
 
 # Load packages required to define the pipeline:
 library(targets)
-# library(tarchetypes) # Load other packages as needed.
+library(tarchetypes)
+# Load other packages as needed.
 
 # Set target options:
 tar_option_set(
-  packages = c("tibble") # Packages that your targets need for their tasks.
+  # Packages that your targets need for their tasks.
+  packages = c("tidyverse", 
+               "metafor",
+               "rstan",
+               "brms",
+               "tidybayes") 
   # format = "qs", # Optionally set the default storage format. qs is fast.
   #
   # Pipelines that take a long time to run may benefit from
@@ -45,8 +51,8 @@ tar_option_set(
 )
 
 # Run the R scripts in the R/ folder with your custom functions:
-tar_source()
-# tar_source("other_functions.R") # Source other scripts as needed.
+tar_source("R/functions")
+# tar_source("R/other_functions.R") # Source other scripts as needed.
 
 # Replace the target list below with your own:
 list(

### 2024-11-21 analysis_plan.qmd
Grateful report link on analysis plan

@@
 All code utilised for data preparation and analyses are available in the corresponding GitHub repository <https://github.com/jamessteeleii/pcos_ree_meta>.
-We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: [ADD LINKS](ADD LINK).
+We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: [https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf](https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf).
 

### 2025-10-14 methods_results.qmd
Added to pipeline esitmation of missing BMIs, heights, weights.

---
title: "Methods and Results"
format: docx
bibliography: 
  mylibrary.bib
csl: apa.csl
execute: 
  echo: false
  message: false
  warning: false
---
```{r}
library(kableExtra)
library(tidybayes)

targets::tar_config_set(store = here::here('_targets'))

```

# Methods
This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024. The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".

## Search Strategy
PubMed, Web of Science, and MEDLINE databases were searched using the following Boolean search string: (("Basal Metabolic Rate"[MeSH] OR "Energy Metabolism"[MeSH] OR "Resting Metabolic Rate" OR RMR OR "Resting Energy Expenditure" OR REE OR "Basal Metabolic Rate" OR BMR OR "resting energy" OR "basal energy expenditure") AND ("Polycystic Ovary Syndrome"[MeSH] OR "Polycystic Ovary Syndrome" OR PCOS OR "Polycystic Ovarian Disease" OR "Stein-Leventhal Syndrome")). Searches were limited to publications up until May 2025 when the search was completed, limited to English language articles, and Rayyan was used to manage the search and screening process.

## Eligibility Criteria
The condition being studied was PCOS and we included observational cross-sectional design studies, in addition to studies of intervention where REE was reported for the population (and if included, the comparator) condition of interest. Following the PICO framework our eligibility criteria were as follows:

* Population
  + Inclusion criteria:
    - Women
    - 18-65 y
    - With or without insulin resistance (IS)
  + Exclusion criteria:
    -Studies using invalid or non-standard methods for measuring REE (e.g., predicted REE from body composition)

* Intervention(s) or exposure(s)
  + Otherwise healthy women with PCOS

* Comparator(s) or control(s)
  + Otherwise healthy control women without PCOS

* Outcome
 + Resting energy expenditure (REE) measured via multiple methods including direct/indirect calorimetry, doubly labelled water.

Studies using invalid or non-standard methods for measuring BMR (e.g., predicted BMR from body composition) were excluded.

## Data extraction (selection and coding)
Bibliometric data including authors, journal, and title of article were extracted. Descriptive statistics for age, body mass, fat mass, fat free mass, height, BMI, race, physical activity levels, country of investigation, information regarding glucose/insulin regulation and insulin resistance status (where available) were extracted for each arm within each study in addition to sample size. Descriptive characteristics were then tabulated across studies for reporting.

The method of measuring REE was extracted and the units of measurement for which REE was reported. For each arm, and observation time point if multiple observations reported (e.g., before and after an intervention), depending on what was reported by the authors we extracted the means, medians, standard deviations, standard errors, lower and upper range values, and interquartile range for the unadjusted and/or body mass adjusted and/or fat free mass adjusted REE values. Where REE values adjusted for body mass and/or fat free mass were reported we used the reported body mass and/or fat free mass mean values for that arm to convert them to unadjusted REE values (i.e., multiplied them by body mass and/or fat free mass mean values). Where means and/or standard deviations were missing the latter were either calculated from standard errors and sample size, or all both were estimated from lower and upper range, interquartile range, median, and sample size depending on the information available using the methods of Wan et al. [-@wanEstimatingSampleMean2014]. All REE values were converted to kcal/day.

<!-- ## Risk of bias (quality) assessment -->
<!-- The assessment will be performed at the study level with the Joanna Briggs Institute (JBI) Critical Appraisal Checklist for Analytical Cross Sectional Studies, which assesses the methodological quality of a study and determines the extent to which a study has addressed the possibility of bias in its design, conduct and analysis. -->

## Statistical Analysis

Statistical analysis of the data extracted was be performed in R, (v 4.3.3; R Core Team, https://www.r-project.org/) and RStudio (v 2023.06.1; Posit, https://posit.co/). All code utilised for data preparation, transformations, analyses, plotting, and reporting are available in the corresponding GitHub repository <https://github.com/jamessteeleii/pcos_ree_meta>. We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf>. The statistical analysis plan was linked in our pre-registration (PROSPERO: [CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) and available at the accompanying github repository. Any deviations from the pre-registration are noted below.

Given our research question our analysis was aimed at parameter estimation [@cummingNewStatisticsWhy2014] within a Bayesian meta-analytic framework [@kruschkeBayesianNewStatistics2018]. For all analyses model parameter estimates and their precision, along with conclusions based upon them, are interpreted continuously and probabilistically, considering data quality, plausibility of effect, and previous literature, all within the context of each model. The `renv` package [@usheyRenvProjectEnvironments2023] was used for package version reproducibility and a function based analysis pipeline using the `targets` package [@landauTargetsDynamicFunctionOriented2023] was employed (the analysis pipeline can be viewed by downloading the R Project and running the function `targets::tar_visnetwork()`). Effect sizes and their variances were all calculated using the `metafor` packages `escalc()` function [@viechtbauerMetaforMetaAnalysisPackage2023]. The main package `brms` [@burknerBrmsBayesianRegression2023] was used in fitting all the Bayesian meta-analysis models. Prior and posterior draws were taken using `marginaleffects` [@arel-bundockMarginaleffectsPredictionsComparisons2023] and `tidybayes` [@kayTidybayesTidyData2023] packages. All visualisations are created using `ggplot2` [@wickhamGgplot2CreateElegant2023], `tidybayes`, and the `patchwork` [@pedersenPatchworkComposerPlots2023] packages.

### Main Pre-registered Models

We adopted an arm-based multiple condition comparison (i.e., network) type model given that the studies included had arms of women with PCOS both with, and without, a non-PCOS control arm [@hongBayesianMissingData2016], and also in some cases multiple observations of REE in the different arms included in the study. In typical contrast-based meta-analyses data is limited to the effect sizes for paired contrasts between arms and thus studies that include both arms (i.e., relative effects between non-PCOS control vs PCOS arms); however, in arm-based analyses the data are the absolute effects within each arm and information is borrowed across studies to enable both within condition absolute, and between condition relative contrasts to be estimated. From this model we focus on reporting the global grand mean estimate for the fixed between condition relative contrast for non-PCOS control vs PCOS arms as our primary estimand of interest (i.e., $\beta_1$ in both mean and standard deviation models). We examined both raw mean REE (i.e., the absolute mean REE in kcals per day for each arm) in addition to the between person standard deviation in REE (i.e., the absolute standard deviation in REE in kcals per day for each arm). Both models were multilevel in that they included nested random intercepts for both study, and arm within study, and in addition and in deviation from our pre-registration we also included lab as a random intercept as in some cases we had multiple studies from the same lab or research group included. Lastly, the inclusion of a random intercept for each effect size was also accidentally omitted from our pre-registration and so this is also included in the model.

#### Mean REE Model

The main model for mean REE with $\textrm{cond}$ representing the condition (either control or PCOS) was as follows:
$$
\begin{align*}
\hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]

\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl]
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[10pt]

\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim 
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}

\text{, for lab i = 1,} \dots \text{,I} \\[12pt]

\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim 
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix} 
\text{, for study j = 1,} \dots \text{,J} \\[12pt]

\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}) 
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}) 
\text{, for effect l = 1,} \dots \text{,L} \\[12pt]

\end{align*}
$$

where $\hat{\theta}_{ijkl}$ is the $l\textrm{th}$ mean REE estimate from the $k\textrm{th}$ arm, for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma^2_{ijkl}$ is the corresponding sampling error for that estimate. The random intercepts for the $i\textrm{th}$ lab, $j\textrm{th}$ study, $k\textrm{th}$ arm, and $l\textrm{th}$ mean REE estimate are $\alpha_{0,lab[i]}$, $\alpha_{0,study[j]}$, $\alpha_{0,arm[k]}$, and $\alpha_{0,effect[l]}$ respectively each with standard deviation of $\sigma^2_{0,lab[i]}$, $\sigma^2_{0,study[j]}$, $\sigma^2_{0,arm[k]}$, and $\sigma^2_{0,effect[l]}$. The parameter $\beta_0$ represents the fixed effect estimate of REE for control conditions and $\beta_1$ the fixed effect estimate for the offset from this for the PCOS conditions (i.e., the difference between conditions). The estimated offset was allowed to vary across both labs and studies each reflected by $\alpha_{1,lab[i]}$ and $\alpha_{1,study[j]}$ respectively, and these effects were also modelled as correlated with the corresponding random intercepts with covariance $\Sigma_{lab}$ and $\Sigma_{study}$, and $\mathrm{corr}_{\mathrm{lab}}$ and $\mathrm{corr}_{\mathrm{study}}$ correlation matrices.

The priors for this model were as follows:

$$
\begin{align*}
\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]

\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]

\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{align*}
$$

where the prior for $\beta_{0}$, which corresponded to the model intercept and mean REE in the control condition, was set based on meta-analysis of the mean REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this. The prior for the fixed effect $\beta_{1}$, reflecting the difference between control and PCOS conditions was set based on a wide range of possible values considering the minimum and and maximum values of the ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We then set a prior that permits values approximately across this range of values with the majority of it's mass centred around zero. The remaining random effects were set based on the default weakly regularising priors for `brms` and scaled to the expected response values using a $\textrm{half-student-}t$ distribution with 3 degrees of freedom and $\mu=0$, and both correlation matrices $\textrm{corr}_{lab}$ and $\textrm{corr}_{study}$ were set with an $\textrm{LKJcorr(1)}$ distribution.

#### Standard Deviation of REE Model

The main model for the standard deviation of REE with $\textrm{cond}$ representing the condition (either control or PCOS) was as follows:

$$
\begin{align*}
\log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]

\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]} 
       + \beta_2\,\tilde{m}_{[ijkl]}
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[8pt]

m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt]
\tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt]

\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim 
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}

\text{, for lab i = 1,} \dots \text{,I} \\[12pt]

\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim 
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix} 
\text{, for study j = 1,} \dots \text{,J} \\[12pt]

\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}) 
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}) 
\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
\end{align*}
$$
where $\log(\hat{\theta}_{ijkl})$ is the $l\textrm{th}$ natural logarithm of the standard deviation of REE estimate from the $k\textrm{th}$ arm, for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma^2_{ijkl}$ is the corresponding sampling error for that estimate. The random intercepts for the $i\textrm{th}$ lab, $j\textrm{th}$ study, $k\textrm{th}$ arm, and $l\textrm{th}$ standard deviation of REE estimate are $\alpha_{0,lab[i]}$, $\alpha_{0,study[j]}$, $\alpha_{0,arm[k]}$, and $\alpha_{0,effect[l]}$ respectively each with standard deviation of $\sigma^2_{0,lab[i]}$, $\sigma^2_{0,study[j]}$, $\sigma^2_{0,arm[k]}$, and $\sigma^2_{0,effect[l]}$. The parameter $\beta_0$ represents the fixed effect estimate of standard deviation of REE for control conditions and $\beta_1$ the fixed effect estimate for the offset from this for the PCOS conditions (i.e., the difference between conditions). The estimated offset was allowed to vary across both labs and studies each reflected by $\alpha_{1,lab[i]}$ and $\alpha_{1,study[j]}$ respectively, and these effects were also modelled as correlated with the corresponding random intercepts with covariance $\Sigma_{lab}$ and $\Sigma_{study}$, and $\mathrm{corr}_{\mathrm{lab}}$ and $\mathrm{corr}_{\mathrm{study}}$ correlation matrices. Finally, $\beta_{2}$ represents the fixed effect of the natural logarithm of the corresponding mean REE estimate $\tilde{m}$ which is modelled as estimated with measurement error i.e., $m$ represents the point estimate for the $l\textrm{th}$ natural logarithm of the mean REE estimate from the $k\textrm{th}$ arm, for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2$ is the corresponding sampling error for that estimate.

The priors for this model were as follows:
$$
\begin{align*}
\beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt]
\beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt]

\text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt]
\text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt]

\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt]

\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{align*}

$$
where the prior for $\beta_{0}$, which corresponded to the model intercept and standard deviation of REE in the control condition, was again set based on meta-analysis of the standard deviation of REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this. The prior for the fixed effect $\beta_{1}$, reflecting the difference between control and PCOS conditions was set based on a wide range of possible values. Given that in many cases of variables in the field there is an approximate relationship of ~1 for the natural logarithm of the standard deviation conditioned upon the natural logarithm of the mean [@steeleMetaanalysisVariationSport2023a] we set this prior to reflect the range of differences on the on the log scale (i.e., $\log(1584)$). We then set a prior that permits values approximately across this range of values with the majority of it's mass centred around zero. The prior for the fixed effect $\beta_{2}$, reflecting the relationship between the natural logarithm of the mean REE with the natural logarithm of the standard deviation of REE, was set it to be weakly informative centred on zero with a wide scale to indicate uncertainty in this outcome specifically despite the typical relationship close to ~1. Priors for the measurement error of the natural logarithm of the mean REE estimate $\tilde{m}$ were again based upon meta-analysis of the aforementioned studies, though measurement error has to be positive, as does the corresponding standard deviation of this error, so we set these to conservative wide $\textrm{half-student-}t$ distributions. The remaining random effects were set based on the default weakly regularising priors for `brms` and scaled to the expected response values using a $\textrm{half-student-}t$ distribution with 3 degrees of freedom and $\mu=0$, and both correlation matrices $\textrm{corr}_{lab}$ and $\textrm{corr}_{study}$ were set with an $\textrm{LKJcorr(1)}$ distribution.

#### Post-processing of models
For both models we examined trace plots along with $\hat{R}$ values to examine whether chains have converged, and posterior predictive checks for each model to understand the model implied distributions. From each model we  took draws from the posterior distributions for the conditional absolute estimates for each condition (i.e., controls and PCOS) by study incorporating random effects, the global grand mean absolute estimates for each condition ignoring random effects, and the global grand mean between condition relative contrast for controls vs PCOS conditions ignoring random effects. The between condition relative contrast for controls vs PCOS conditions corresponded to $\beta{1}$ in each model and was our primary estimand of interest; for the mean REE model this corresponded to the absolute difference in mean REE, and for the standard deviation of REE model this corresponded to the natural logarithm of the ratio of standard deviations of REE which was exponentiated (note, all log standard deviation of REE model estimates were exponentiated back to the original scale to aid interpretibility). We present the full probability density functions for posterior visually, and also to calculate mean and 95% quantile intervals (QI: i.e., 'credible' or 'compatibility' intervals) for each estimate providing the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles given our priors and data.

### Sensitivity analysis using pairwise contrast based models

By way of sensitivity analysis we also conducted pairwise contrast based models where we limited the included effects to those extracted from studies including only a directly comparable control and PCOS arm at baseline. These models were both as follows:

$$
\begin{align*}
\hat{\theta}_{ij} &\sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij}) \\[6pt]

\mu_{ij} &= \beta_0
       + \alpha_{0,\mathrm{lab}[i]}
       + \alpha_{0,\mathrm{study}[j]} \\[10pt]

\alpha_{0,\mathrm{lab}[i]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}}) 
\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
\alpha_{0,\mathrm{study}[j]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}}) 
\text{, for study j = 1,} \dots \text{,J} \\[12pt]

\end{align*}
$$

where $\hat{\theta}_{ijkl}$ is the pairwise effect size, either the mean difference in REE or the log coefficient of variation ratio (calculated as PCOS vs control), for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma^2_{ijkl}$ is the corresponding sampling error for that effect size estimate. The random intercepts for the $i\textrm{th}$ lab, $j\textrm{th}$ study are $\alpha_{0,lab[i]}$ and $\alpha_{0,study[j]}$ respectively each with standard deviation of $\sigma^2_{0,lab[i]}$, $\sigma^2_{0,study[j]}$. The parameter $\beta_0$ represents the fixed effect estimate of the pairwise effect size i.e., the pooled estimate of the contrast between conditions. The priors for $\beta_{0}$ in these models were set as default weakly regularising which is set on an intercept that results when internally centering all population-level predictors around zero to improve sampling efficiency and scaled to the expected response values using a $\textrm{student-}t$ distribution; for the mean difference in REE this was $\textrm{student-}t(3,\ 5.5,\ 50.2)$ and for the log coefficient of variance ratio this was $\textrm{student-}t(3,\ -0.1,\ 2.5)$. The random effects were set similarly scaled to the expected response values but using a $\textrm{half-student-}t$ distribution centred on zero. From these models we calculated the mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for the $\beta_0$ (comparable to the $\beta_1$ from the corresponding mean and standard devation of REE arm-based models) for each effect size providing the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles given our priors and data.

# Results
```{r}
targets::tar_load(main_arm_data_imputed_demographics)

main_arm_data_imputed_demographics_baseline <- main_arm_data_imputed_demographics |> 
  filter(timepoint == "baseline")

no_conditions <- main_arm_data_imputed_demographics_baseline |>
  group_by(cond) |>
  count()

no_participants <- main_arm_data_imputed_demographics_baseline |>
  group_by(cond) |>
  summarise(
    min_n = min(n),
    median_n = median(n),
    max_n = max(n),
    total_n = sum(n)
  )

```
A total of `r length(unique(main_arm_data_imputed_demographics$study)` studies from `r length(unique(main_arm_data_imputed_demographics$lab)` labs including `r length(unique(main_arm_data_imputed_demographics$arm)` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`). Descriptive characteristics of the arms and participants in these studies are reported in @tbl-demographics.

```{r}
#| label: tbl-demographics 
#| tbl-cap-location: "top"
#| tbl-cap: Descriptive characteristics of arms and participants for included studies.

main_arm_data_imputed_demographics_baseline |> 
  mutate(across(where(is.numeric), round, 2)) |>
  mutate(
    Age = paste0(m_age, " (", sd_age, ")"),
    `Body mass` = paste0(m_body_mass, " (", sd_body_mass, ")"),
    Height = paste0(m_height, " (", sd_height, ")"),
    BMI = paste0(m_bmi, " (", sd_bmi, ")"),
    `Fat mass` = paste0(m_fat_mass, " (", sd_fat_mass, ")"),
    `Fat free mass` = paste0(m_fat_free_mass, " (", sd_fat_free_mass, ")")
  ) |>
  select(authors, year, title, cond,
         Age, 
         `Body mass`,
         Height,
         BMI,
         `Fat mass`,
         `Fat free mass`,
         race,
         physical_activity_level,
         country,
         -c(m_age, sd_age, 
         m_body_mass, sd_body_mass,
         m_height, sd_height,
         m_bmi, sd_bmi,
         m_fat_mass, sd_fat_mass,
         m_fat_free_mass, sd_fat_free_mass),
         insulin_resistant_description,
         m_body_mass_estim,
         m_height_estim,
         m_bmi_estim
         )|>
  arrange(year) |>
  kable() |>
  collapse_rows(columns = 1, valign = "top") |>
  # footnote(general = c("ST = self-talk", "CON = non-intervention control")) |>
  row_spec(0, bold = TRUE) |>
  kable_classic(full_width = FALSE) 
```

## Mean REE Model Results
```{r}

targets::tar_load(
  c(main_arm_mean_effects_preds_condition,
    main_arm_mean_effects_contrast_condition,
    combined_mean_plot)
)

main_arm_mean_effects_preds_condition <- main_arm_mean_effects_preds_condition |>
    group_by(cond) |>
    mean_qi(draw)

main_arm_mean_effects_contrast_condition <- main_arm_mean_effects_contrast_condition |>
    mean_qi(draw)

```
The main model for mean REE resulted in a posterior distribution for the contrast between control and PCOS conditions with a mean point estimate of `r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day with a 95% quantile interval ranging from  `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)` kcal/day suggesting there is a 95% probability that the true difference lies between these values given our priors and the data from included studies. The corresponding conditional estimates for the control condition and PCOS condition respectively were `r round(main_arm_mean_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[1])` kcal/day] and `r round(main_arm_mean_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[2])` kcal/day]. These results including the full visualisation of the posterior distribution, in addition to the conditional estimates by study, can be seen in @fig-mean-REE.

```{r}
#| label: fig-mean-REE 
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.

combined_mean_plot
```

## Standard Deviation of REE Model Results
```{r}

targets::tar_load(
  c(main_arm_variance_effects_preds_condition,
    main_arm_variance_effects_contrast_condition,
    combined_variance_plot)
)

main_arm_variance_effects_preds_condition <- main_arm_variance_effects_preds_condition |>
    group_by(cond) |>
    mean_qi(draw)

main_arm_variance_effects_contrast_condition <- main_arm_variance_effects_contrast_condition |>
    mean_qi(draw)

```

The main model for the between participant standard deviation of REE resulted in a posterior distribution for the contrast ratio between control and PCOS conditions with a variance point estimate of `r round(main_arm_variance_effects_contrast_condition$draw)` with a 95% quantile interval ranging from `r round(main_arm_variance_effects_contrast_condition$.lower)` to `r round(main_arm_variance_effects_contrast_condition$.upper)` suggesting there is a 95% probability that the true ratio of standard deviations lies between these values given our priors and the data from included studies. The corresponding conditional estimates for the standard deviations of the control condition and PCOS condition respectively were `r round(main_arm_variance_effects_preds_condition$draw[1])` [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` to `r round(main_arm_variance_effects_preds_condition$.upper[1])` ] and `r round(main_arm_variance_effects_preds_condition$draw[2])` [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` to `r round(main_arm_variance_effects_preds_condition$.upper[2])` ]. These results including the full visualisation of the posterior distribution, in addition to the conditional estimates by study, can be seen in @fig-variance-REE.

```{r}
#| label: fig-variance-REE 
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.

combined_variance_plot
```

### 2025-10-14 R/functions/main_data_functions.R

@@ -109,7 +109,8 @@ prepare_data <- function(file) {
         .default = sd_ffm_adjusted
       )
     ) |>
-    ungroup()
+    ungroup() |>
+    rowid_to_column("effect")
   
   return(data)
 }
@@ -139,6 +140,53 @@ calculate_arm_effects <- function(data) {
   return(data)
 }
 
+impute_demographic_estimates <- function(data) {
+  data <- data |>
+    
+    # convert height to cm for reporting
+    mutate(m_height = case_when(
+      m_height < 100 ~ m_height*100
+      ),
+      sd_height = case_when(
+        m_height < 100 ~ sd_height*100
+      )
+    ) |>
+    
+    # impute estimates of missing weight/height/bmi and note where estimate
+    mutate(
+      m_body_mass_estim = case_when(
+        is.na(m_body_mass) & !is.na(m_height) & !is.na(m_bmi) ~ "y",
+        .default = "n"
+      ),
+      m_height_estim = case_when(
+        is.na(m_height) & !is.na(m_body_mass) & !is.na(m_bmi) ~ "y",
+        .default = "n"
+      ),
+      m_bmi_estim = case_when(
+        is.na(m_bmi) & !is.na(m_body_mass) & !is.na(m_height) ~ "y",
+        .default = "n"
+      ),
+    ) |>
+    
+    mutate(
+      m_body_mass = case_when(
+        is.na(m_body_mass) & !is.na(m_height) & !is.na(m_bmi) ~ m_bmi * (m_height/100)^2,
+        .default = m_body_mass
+      ),
+      m_height = case_when(
+        is.na(m_height) & !is.na(m_body_mass) & !is.na(m_bmi) ~ (sqrt(m_body_mass/m_bmi))*100,
+        .default = m_height
+      ),
+      m_bmi = case_when(
+        is.na(m_bmi) & !is.na(m_body_mass) & !is.na(m_height) ~ m_body_mass/(m_height/100)^2,
+        .default = m_bmi
+      ),
+    )
+  
+  return(data)
+    
+}
+
 # Pairwise data preparation for sensitivity analysis
 
 prepare_pairwise_data <- function(data) {
@@ -155,7 +203,6 @@ prepare_pairwise_data <- function(data) {
   
   return(wide_data)
 }
-
 
 calculate_pairwise_effects <- function(data) {
   

### 2025-10-14 R/functions/model_functions.R

@@ -62,7 +62,7 @@ set_prior_arm_mean_effects <- function() {
 
 fit_arm_mean_effects_model <- function(data, prior) {
   
-  arm_model <- brm(yi_mean | se(sqrt(vi_mean)) ~ 0 + Intercept + cond + (1 + cond | lab) + (1 + cond | study) + (1 | arm),
+  arm_model <- brm(yi_mean | se(sqrt(vi_mean)) ~ 0 + Intercept + cond + (1 + cond | lab) + (1 + cond | study) + (1 | arm) + (1|effect),
                    data = data,
                    prior = prior,
                    chains = 4,
@@ -167,7 +167,7 @@ set_prior_arm_variance_effects <- function() {
 
 fit_arm_variance_effects_model <- function(data, prior) {
   
-  arm_model <- brm(yi_sd | se(sqrt(vi_sd)) ~ 0 + Intercept + cond + me(log_yi_mean, se_log_yi_mean) + (1 + cond | lab) + (1 + cond | study) + (1 | arm),
+  arm_model <- brm(yi_sd | se(sqrt(vi_sd)) ~ 0 + Intercept + cond + me(log_yi_mean, se_log_yi_mean) + (1 + cond | lab) + (1 + cond | study) + (1 | arm) + (1|effect),
                    data = data,
                    prior = prior,
                    chains = 4,
@@ -279,7 +279,7 @@ plot_meta_mean_pred <- function(preds) {
                        is.numeric, round, 2),
       aes(
         label = glue::glue("{round(draw)} [{round(.lower)}, {round(.upper)}]"),
-        x = draw, y = 0.1
+        x = draw, y = 0.2
       ),
       size = 3
     ) +
@@ -328,10 +328,10 @@ plot_study_mean_pred <- function(preds, data) {
                        is.numeric, round, 2),
       aes(
         label = glue::glue("{cond}: {round(draw)} [{round(.lower)}, {round(.upper)}]"),
-        x = 2250, y =reorder(study_label, draw), group = cond
+        x = Inf, y =reorder(study_label, draw), group = cond
       ),
       size = 2, position = position_dodge(width = 0.75),
-      hjust = "inward"
+      hjust = 1.1
     ) +
     # Add individual study data
     geom_point(
@@ -343,6 +343,7 @@ plot_study_mean_pred <- function(preds, data) {
       alpha = 0.75,
[…]

### 2025-10-14 R/quick_look.R

@@ -1,165 +1,0 @@
-library(tidyverse)
-library(metafor)
-library(brms)
-
-dat <- tibble(
-  study = c("Segal, 1990",
-            "Robinson, 1992",
-            "Cosar, 2008",
-            "Georgopoulos, 2009",
-            "Graff, 2013",
-            "Larsson, 2015",
-            "Doh, 2016"),
-  n_PCOS = c(10,14,31,91,61,72,14),
-  n_CONT = c(11,14,29,48,44,30,10),
-  m_PCOS = c(1508,1624,1167,1446,1469,1411,1272),
-  m_CONT = c(1558,1633,1046,1841,1453,1325,1240),
-  s_PCOS = c(168,137,371,725,227,229,167),
-  s_CONT = c(186,215,296,305,249,193,216)
-)
-
-dat <- escalc(measure = "MD",
-              m1i = m_PCOS,
-              m2i = m_CONT,
-              sd1i = s_PCOS,
-              sd2i = s_CONT,
-              n1i = n_PCOS,
-              n2i = n_CONT,
-              data = dat)
-
-# fit brms model
-
-paired_model <- brm(yi | se(sqrt(vi)) ~ 1 + (1 | study),
-                    data = dat,
-                    chains = 4,
-                    cores = 4,
-                    seed = 1988,
-                    warmup = 2000,
-                    iter = 8000)
-
-plot(paired_model)
-pp_check(paired_model)
-
-
-# dat_var <- escalc(measure = "CVR",
-#               m1i = m_PCOS,
-#               m2i = m_CON,
-#               sd1i = sd_PCOS,
-#               sd2i = sd_CON,
-#               n1i = n_PCOS,
-#               n2i = n_CON,
-#               data = dat)
-# 
-# forest(rma(yi, vi, data = dat_var), slab = study)
-
-
-dat2 <- dat |>
-  pivot_longer(
-    2:7,
-    cols_vary = "slowest",
-    names_to = c(".value", "cond"),
-    names_pattern = "(..)(....)"
-  ) |>
-  rename(
-    n = "n_",
-    m = "m_",
-    sd = "s_"
-    ) |>
-  mutate(
-    se = NA
-  )
-  
-dat_arm <- bind_rows(
-  dat2, tibble(
[…]

### 2025-10-14 R/quick_look_main.R

@@ -1,437 +1,0 @@
-library(tidyverse)
-library(patchwork)
-library(metafor)
-library(brms)
-library(marginaleffects)
-library(tidybayes)
-
-data <- read_csv("data/studies_data.csv") |>
-  
-  # convert all to kcal
-  mutate(
-    across(
-      # apply to all numeric columns except 'unit'
-      .cols = c(mean:iqr_ffm_adjusted),
-      .fns = ~ case_when(
-        units == "kj"   ~ .x / 4.184,     # 1 kcal = 4.184 kJ
-        units == "MJ"   ~ .x * 1000 / 4.184, # 1 MJ = 1000 kJ
-        units == "kcal/kg"  ~ .x * m_body_mass, # note this is just for Pohlmeier et al.
-        .default = .x,             # already kcal      
-        )
-    )
-  ) |>
-  
-  # convert standard errors to standard deviations
-  mutate(
-    sd = case_when(
-     is.na(sd) == TRUE ~ se * sqrt(n),
-     .default = sd
-    ),
-    sd_bm_adjusted  = case_when(
-      is.na(sd_bm_adjusted) == TRUE ~ se_bm_adjusted  * sqrt(n),
-      .default = sd_bm_adjusted 
-    ),
-    sd_ffm_adjusted = case_when(
-      is.na(sd_ffm_adjusted) == TRUE ~ se_ffm_adjusted * sqrt(n),
-      .default = sd_ffm_adjusted
-    )
-  ) |>
-  
-  # estimate means and standard deviations from range, iqr, median, and sample size 
-  # (see DOI: 10.1186/1471-2288-14-135)
-  rowwise() |>
-  mutate(
-    scenario = case_when(
-      # unadjusted REEs
-      !is.na(lower_range) & !is.na(upper_range) & is.na(iqr) ~ 1,  # Scenario 1
-      is.na(lower_range) & is.na(upper_range) & !is.na(iqr)  ~ 2,  # Scenario 2
-      !is.na(lower_range) & !is.na(upper_range) & !is.na(iqr) ~ 3, # Scenario 3
-      
-      # bm adjusted REEs
-      !is.na(lower_range_bm_adjusted) & !is.na(upper_range_bm_adjusted) & is.na(iqr_bm_adjusted) ~ 1,  # Scenario 1
-      is.na(lower_range_bm_adjusted) & is.na(upper_range_bm_adjusted) & !is.na(iqr_bm_adjusted)  ~ 2,  # Scenario 2
[…]

### 2025-10-14 _targets.R

@@ -126,6 +126,11 @@ list(
     calculate_arm_effects(main_arm_data)
   ),
   
+  tar_target(
+    main_arm_data_imputed_demographics,
+    impute_demographic_estimates(main_arm_data)
+  ),
+  
   # Fitting main analysis models
   
   tar_target(
@@ -206,6 +211,14 @@ list(
   ),
   
   tar_target(
+    combined_mean_plot_tiff,
+    ggsave("plots/combine_mean_plot.tiff",
+           plot = combined_mean_plot,
+           dpi = 300, device = "tiff",
+           w = 10, h = 5)
+  ),
+  
+  tar_target(
     meta_variance_pred_plot,
     plot_meta_variance_pred(main_arm_variance_effects_preds_condition)
   ),
@@ -226,6 +239,14 @@ list(
     combine_variance_plots(meta_variance_pred_plot,
                            meta_study_variance_pred_plot,
                        meta_variance_contrast_plot)
+  ),
+  
+  tar_target(
+    combined_variance_plot_tiff,
+    ggsave("plots/combine_variance_plot.tiff",
+           plot = combined_variance_plot,
+           dpi = 300, device = "tiff",
+           w = 10, h = 5)
   ),
   
   #### Pairwise sensitivity analysis ----

### 2025-10-14 methods_results.qmd
methods/results finished, need to fix descriptives table and equations to docx

@@
 ```{r}
-library(kableExtra)
+library(tidyverse)
+library(flextable)
 library(tidybayes)
@@
 The main model for mean REE with $\textrm{cond}$ representing the condition (either control or PCOS) was as follows:
+<!-- $$ -->
+<!-- \begin{align*} -->
+<!-- \hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt] -->
+
+<!-- \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl] -->
+<!--        + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} -->
+<!--        + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} -->
+<!--        + \alpha_{0,\mathrm{arm}[k]} -->
+<!--        + \alpha_{0,\mathrm{effect}[l]} \\[10pt] -->
+
+<!-- \begin{pmatrix} -->
+<!-- \alpha_{0,\mathrm{lab}[i]} \\[4pt] -->
+<!-- \alpha_{1,\mathrm{lab}[i]} -->
+<!-- \end{pmatrix} -->
+<!-- &\sim  -->
+<!-- \mathcal{N}\!\left( -->
+<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
+<!-- \Sigma_{\mathrm{lab}} -->
+<!-- \right), -->
+<!-- \quad -->
+<!-- \Sigma_{\mathrm{lab}} = -->
+<!-- \begin{pmatrix} -->
+<!-- \sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt] -->
+<!-- \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}} -->
+<!-- \end{pmatrix}  -->
+
+<!-- \text{, for lab i = 1,} \dots \text{,I} \\[12pt] -->
+
+<!-- \begin{pmatrix} -->
+<!-- \alpha_{0,\mathrm{study}[j]} \\[4pt] -->
+<!-- \alpha_{1,\mathrm{study}[j]} -->
+<!-- \end{pmatrix} -->
+<!-- &\sim  -->
+<!-- \mathcal{N}\!\left( -->
+<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
+<!-- \Sigma_{\mathrm{study}} -->
+<!-- \right), -->
+<!-- \quad -->
+<!-- \Sigma_{\mathrm{study}} = -->
+<!-- \begin{pmatrix} -->
+<!-- \sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt] -->
+<!-- \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}} -->
+<!-- \end{pmatrix}  -->
+<!-- \text{, for study j = 1,} \dots \text{,J} \\[12pt] -->
+
+<!-- \alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})  -->
+<!-- \text{, for arm k = 1,} \dots \text{,K} \\[12pt] -->
+<!-- \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})  -->
+<!-- \text{, for effect l = 1,} \dots \text{,L} \\[12pt] -->
+
+<!-- \end{align*} -->
+<!-- $$ -->
+
 $$
-\begin{align*}
-\hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
+\hat{\theta}_{ijkl} \sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl})
+$$
 
-\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl]
+$$
+\mu_{ijkl} = \beta_0 + \beta_1\,\text{cond}_{ijkl} 
 + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} 
 + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} 
-       + \alpha_{0,\mathrm{arm}[k]}
-       + \alpha_{0,\mathrm{effect}[l]} \\[10pt]
++ \alpha_{0,\mathrm{arm}[k]} + \alpha_{0,\mathrm{effect}[l]}
+$$
 
+$$
 \begin{pmatrix}
-\alpha_{0,\mathrm{lab}[i]} \\[4pt]
+\alpha_{0,\mathrm{lab}[i]} \\
 \alpha_{1,\mathrm{lab}[i]}
 \end{pmatrix}
-&\sim 
-\mathcal{N}\!\left(
-\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
+\sim 
+\mathcal{N}\Bigg(
+\begin{pmatrix} 0 \\ 0 \end{pmatrix},
 \Sigma_{\mathrm{lab}}
-\right),
+\Bigg), 
 \quad
@@
 \begin{pmatrix}
-\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
-\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
-\end{pmatrix}
+\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} \\
+\rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
+\end{pmatrix},
+\text{ for lab } i = 1, \dots, I
+$$
 
-\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
-
+$$
 \begin{pmatrix}
-\alpha_{0,\mathrm{study}[j]} \\[4pt]
+\alpha_{0,\mathrm{study}[j]} \\
 \alpha_{1,\mathrm{study}[j]}
 \end{pmatrix}
-&\sim 
-\mathcal{N}\!\left(
-\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
+\sim 
+\mathcal{N}\Bigg(
+\begin{pmatrix} 0 \\ 0 \end{pmatrix},
 \Sigma_{\mathrm{study}}
-\right),
+\Bigg), 
 \quad
@@
 \begin{pmatrix}
-\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
-\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
-\end{pmatrix} 
-\text{, for study j = 1,} \dots \text{,J} \\[12pt]
+\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} \\
+\rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
+\end{pmatrix},
+\text{ for study } j = 1, \dots, J
+$$
 
-\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}) 
-\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
-\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}) 
-\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
+$$
+\alpha_{0,\mathrm{arm}[k]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}), 
+\text{ for arm } k = 1, \dots, K
+$$
 
-\end{align*}
+$$
+\alpha_{0,\mathrm{effect}[l]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}), 
+\text{ for effect } l = 1, \dots, L
 $$
@@
 
+<!-- $$ -->
+<!-- \begin{align*} -->
+<!-- \beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt] -->
+<!-- \beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt] -->
+
+<!-- \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt] -->
+<!-- \sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,  -->
+<!-- \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt] -->
+
+<!-- \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt] -->
+<!-- \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1) -->
+<!-- \end{align*} -->
+<!-- $$ -->
+
 $$
-\begin{align*}
-\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
-\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]
+\beta_0 \sim \text{Student-}t(3, 1441.81, 84.56) \\
+\beta_1 \sim \text{Student-}t(3, 0, 200)
+$$
 
-\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
+$$
+\sigma^2_{0,\mathrm{study}} \sim \text{Half-student-}t(3, 149.89, 82.91) \\
 \sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
-\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]
+\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}},\, \sigma^2_{0,\mathrm{effect}} \sim \text{Half-student-}t(3, 0, 112.4)
+$$
 
-\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
-\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
-\end{align*}
+$$
+\mathrm{corr}_{\mathrm{lab}} \sim \mathrm{LKJ}(1) \\
+\mathrm{corr}_{\mathrm{study}} \sim \mathrm{LKJ}(1)
 $$
@@
 
+<!-- $$ -->
+<!-- \begin{align*} -->
+<!-- \log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt] -->
+
+<!-- \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]}  -->
+<!--        + \beta_2\,\tilde{m}_{[ijkl]} -->
+<!--        + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} -->
+<!--        + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} -->
+<!--        + \alpha_{0,\mathrm{arm}[k]} -->
+<!--        + \alpha_{0,\mathrm{effect}[l]} \\[8pt] -->
+
+<!-- m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt] -->
+<!-- \tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt] -->
+
+<!-- \begin{pmatrix} -->
+<!-- \alpha_{0,\mathrm{lab}[i]} \\[4pt] -->
+<!-- \alpha_{1,\mathrm{lab}[i]} -->
+<!-- \end{pmatrix} -->
+<!-- &\sim  -->
+<!-- \mathcal{N}\!\left( -->
+<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
+<!-- \Sigma_{\mathrm{lab}} -->
+<!-- \right), -->
+<!-- \quad -->
+<!-- \Sigma_{\mathrm{lab}} = -->
+<!-- \begin{pmatrix} -->
+<!-- \sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt] -->
+<!-- \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}} -->
+<!-- \end{pmatrix}  -->
+
+<!-- \text{, for lab i = 1,} \dots \text{,I} \\[12pt] -->
+
+<!-- \begin{pmatrix} -->
+<!-- \alpha_{0,\mathrm{study}[j]} \\[4pt] -->
+<!-- \alpha_{1,\mathrm{study}[j]} -->
+<!-- \end{pmatrix} -->
+<!-- &\sim  -->
+<!-- \mathcal{N}\!\left( -->
+<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
+<!-- \Sigma_{\mathrm{study}} -->
+<!-- \right), -->
+<!-- \quad -->
+<!-- \Sigma_{\mathrm{study}} = -->
+<!-- \begin{pmatrix} -->
+<!-- \sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt] -->
+<!-- \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}} -->
+<!-- \end{pmatrix}  -->
+<!-- \text{, for study j = 1,} \dots \text{,J} \\[12pt] -->
+
+<!-- \alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})  -->
+<!-- \text{, for arm k = 1,} \dots \text{,K} \\[12pt] -->
+<!-- \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})  -->
+<!-- \text{, for effect l = 1,} \dots \text{,L} \\[12pt] -->
+<!-- \end{align*} -->
+<!-- $$ -->
+
 $$
-\begin{align*}
-\log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
+\log(\hat{\theta}_{ijkl}) \sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl})
+$$
 
-\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]} 
+$$
+\mu_{ijkl} = \beta_0 + \beta_1\,\text{cond}_{[ijkl]} 
 + \beta_2\,\tilde{m}_{[ijkl]}
@@
 + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} 
-       + \alpha_{0,\mathrm{arm}[k]}
-       + \alpha_{0,\mathrm{effect}[l]} \\[8pt]
++ \alpha_{0,\mathrm{arm}[k]} + \alpha_{0,\mathrm{effect}[l]}
+$$
 
-m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt]
-\tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt]
+$$
+m = \log(y_{i,\mathrm{mean,[ijkl]}})
+$$
 
+$$
+\tilde{m} \sim \mathcal{N}(m, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2)
+$$
+
+$$
 \begin{pmatrix}
-\alpha_{0,\mathrm{lab}[i]} \\[4pt]
+\alpha_{0,\mathrm{lab}[i]} \\
 \alpha_{1,\mathrm{lab}[i]}
 \end{pmatrix}
-&\sim 
-\mathcal{N}\!\left(
-\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
+\sim 
+\mathcal{N}\Bigg(
+\begin{pmatrix} 0 \\ 0 \end{pmatrix},
 \Sigma_{\mathrm{lab}}
-\right),
+\Bigg),
 \quad
@@
 \begin{pmatrix}
-\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
-\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
-\end{pmatrix}
+\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} \\
+\rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
+\end{pmatrix},
+\text{ for lab } i = 1, \dots, I
+$$
 
-\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
-
+$$
 \begin{pmatrix}
-\alpha_{0,\mathrm{study}[j]} \\[4pt]
+\alpha_{0,\mathrm{study}[j]} \\
 \alpha_{1,\mathrm{study}[j]}
 \end{pmatrix}
-&\sim 
-\mathcal{N}\!\left(
-\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
+\sim 
+\mathcal{N}\Bigg(
+\begin{pmatrix} 0 \\ 0 \end{pmatrix},
 \Sigma_{\mathrm{study}}
-\right),
+\Bigg),
 \quad
@@
 \begin{pmatrix}
-\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
-\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
-\end{pmatrix} 
-\text{, for study j = 1,} \dots \text{,J} \\[12pt]
+\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} \\
+\rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
+\end{pmatrix},
+\text{ for study } j = 1, \dots, J
+$$
 
-\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}) 
-\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
-\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}) 
-\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
-\end{align*}
 $$
+\alpha_{0,\mathrm{arm}[k]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}), 
+\text{ for arm } k = 1, \dots, K
+$$
+
+$$
+\alpha_{0,\mathrm{effect}[l]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}), 
+\text{ for effect } l = 1, \dots, L
+$$
+
 where $\log(\hat{\theta}_{ijkl})$ is the $l\textrm{th}$ natural logarithm of the standard deviation of REE estimate from the $k\textrm{th}$ arm, for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma^2_{ijkl}$ is the corresponding sampling error for that estimate.
@@
 The priors for this model were as follows:
-$$
-\begin{align*}
-\beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt]
-\beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt]
-\beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt]
+<!-- $$ -->
+<!-- \begin{align*} -->
+<!-- \beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt] -->
+<!-- \beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt] -->
+<!-- \beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt] -->
 
-\text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt]
-\text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt]
+<!-- \text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt] -->
+<!-- \text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt] -->
 
-\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt]
-\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
-\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt]
+<!-- \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt] -->
+<!-- \sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, -->
+<!-- \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt] -->
 
-\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
-\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
-\end{align*}
+<!-- \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt] -->
+<!-- \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1) -->
+<!-- \end{align*} -->
+
+<!-- $$ -->
 
 $$
+\beta_0 \sim \text{Student-}t(3, 5.54, 0.80) \\
+\beta_1 \sim \text{Student-}t(3, 0, 5.3) \\
+\beta_2 \sim \text{Student-}t(3, 0, 2.5)
+$$
+
+$$
+\text{mean}(\tilde{m}) \sim \text{Half-student-}t(3, 7.28, 0.62) \\
+\text{sd}(\tilde{m}) \sim \text{Half-student-}t(3, 0, 5)
+$$
+
+$$
+\sigma^2_{0,\mathrm{study}} \sim \text{Half-student-}t(3, 1.08, 1.06) \\
+\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
+\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}},\, \sigma^2_{0,\mathrm{effect}} 
+\sim \text{Half-student-}t(3, 0, 2.5)
+$$
+
+$$
+\mathrm{corr}_{\mathrm{lab}} \sim \mathrm{LKJ}(1) \\
+\mathrm{corr}_{\mathrm{study}} \sim \mathrm{LKJ}(1)
+$$
+
 where the prior for $\beta_{0}$, which corresponded to the model intercept and standard deviation of REE in the control condition, was again set based on meta-analysis of the standard deviation of REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this.
@@
 
+<!-- $$ -->
+<!-- \begin{align*} -->
+<!-- \hat{\theta}_{ij} &\sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij}) \\[6pt] -->
+
+<!-- \mu_{ij} &= \beta_0 -->
+<!--        + \alpha_{0,\mathrm{lab}[i]} -->
+<!--        + \alpha_{0,\mathrm{study}[j]} \\[10pt] -->
+
+<!-- \alpha_{0,\mathrm{lab}[i]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}})  -->
+<!-- \text{, for lab i = 1,} \dots \text{,I} \\[12pt] -->
+<!-- \alpha_{0,\mathrm{study}[j]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}})  -->
+<!-- \text{, for study j = 1,} \dots \text{,J} \\[12pt] -->
+
+<!-- \end{align*} -->
+<!-- $$ -->
+
 $$
-\begin{align*}
-\hat{\theta}_{ij} &\sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij}) \\[6pt]
+\hat{\theta}_{ij} \sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij})
+$$
 
-\mu_{ij} &= \beta_0
-       + \alpha_{0,\mathrm{lab}[i]}
-       + \alpha_{0,\mathrm{study}[j]} \\[10pt]
+$$
+\mu_{ij} = \beta_0 + \alpha_{0,\mathrm{lab}[i]} + \alpha_{0,\mathrm{study}[j]}
+$$
 
-\alpha_{0,\mathrm{lab}[i]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}}) 
-\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
-\alpha_{0,\mathrm{study}[j]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}}) 
-\text{, for study j = 1,} \dots \text{,J} \\[12pt]
+$$
+\alpha_{0,\mathrm{lab}[i]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}}), 
+\text{ for lab } i = 1, \dots, I
+$$
 
-\end{align*}
+$$
+\alpha_{0,\mathrm{study}[j]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}}), 
+\text{ for study } j = 1, \dots, J
 $$
@@
 ```
-A total of `r length(unique(main_arm_data_imputed_demographics$study)` studies from `r length(unique(main_arm_data_imputed_demographics$lab)` labs including `r length(unique(main_arm_data_imputed_demographics$arm)` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
-Descriptive characteristics of the arms and participants in these studies are reported in @tbl-demographics.
-
-```{r}
-#| label: tbl-demographics 
-#| tbl-cap-location: "top"
-#| tbl-cap: Descriptive characteristics of arms and participants for included studies.
-
-main_arm_data_imputed_demographics_baseline |> 
-  mutate(across(where(is.numeric), round, 2)) |>
-  mutate(
-    Age = paste0(m_age, " (", sd_age, ")"),
-    `Body mass` = paste0(m_body_mass, " (", sd_body_mass, ")"),
-    Height = paste0(m_height, " (", sd_height, ")"),
-    BMI = paste0(m_bmi, " (", sd_bmi, ")"),
-    `Fat mass` = paste0(m_fat_mass, " (", sd_fat_mass, ")"),
-    `Fat free mass` = paste0(m_fat_free_mass, " (", sd_fat_free_mass, ")")
-  ) |>
-  select(authors, year, title, cond,
-         Age, 
-         `Body mass`,
-         Height,
-         BMI,
-         `Fat mass`,
-         `Fat free mass`,
-         race,
-         physical_activity_level,
-         country,
-         -c(m_age, sd_age, 
-         m_body_mass, sd_body_mass,
-         m_height, sd_height,
-         m_bmi, sd_bmi,
-         m_fat_mass, sd_fat_mass,
-         m_fat_free_mass, sd_fat_free_mass),
-         insulin_resistant_description,
-         m_body_mass_estim,
-         m_height_estim,
-         m_bmi_estim
-         )|>
-  arrange(year) |>
-  kable() |>
-  collapse_rows(columns = 1, valign = "top") |>
-  # footnote(general = c("ST = self-talk", "CON = non-intervention control")) |>
-  row_spec(0, bold = TRUE) |>
-  kable_classic(full_width = FALSE) 
-```
+A total of `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` labs including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
+Descriptive characteristics of the arms and participants in these studies are reported in Table 1.
 
@@
 
-The main model for the between participant standard deviation of REE resulted in a posterior distribution for the contrast ratio between control and PCOS conditions with a variance point estimate of `r round(main_arm_variance_effects_contrast_condition$draw)` with a 95% quantile interval ranging from `r round(main_arm_variance_effects_contrast_condition$.lower)` to `r round(main_arm_variance_effects_contrast_condition$.upper)` suggesting there is a 95% probability that the true ratio of standard deviations lies between these values given our priors and the data from included studies.
-The corresponding conditional estimates for the standard deviations of the control condition and PCOS condition respectively were `r round(main_arm_variance_effects_preds_condition$draw[1])` [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` to `r round(main_arm_variance_effects_preds_condition$.upper[1])` ] and `r round(main_arm_variance_effects_preds_condition$draw[2])` [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` to `r round(main_arm_variance_effects_preds_condition$.upper[2])` ].
+The main model for the between participant standard deviation of REE resulted in a posterior distribution for the contrast ratio between control and PCOS conditions with a mean point estimate of `r round(main_arm_variance_effects_contrast_condition$draw,2)` with a 95% quantile interval ranging from `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)` suggesting there is a 95% probability that the true ratio of standard deviations lies between these values given our priors and the data from included studies.
+The corresponding conditional estimates for the standard deviations of the control condition and PCOS condition respectively were `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day] and `r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day].
 These results including the full visualisation of the posterior distribution, in addition to the conditional estimates by study, can be seen in @fig-variance-REE.
@@
 ```
+
+## Sensitivity Analysis with Pairwise Models
+```{r}
+targets::tar_load(c(pairwise_mean_effects_model, pairwise_variance_effects_model))
+
+```
+For mean REE the pairwise model resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(pairwise_mean_effects_model)[1])` kcal/day with a 95% quantile interval ranging from  `r round(brms::fixef(pairwise_mean_effects_model)[3])` kcal/day to `r round(brms::fixef(pairwise_mean_effects_model)[4])` kcal/day.
+This was similar for the standard deviation of REE with the pairwise model resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(pairwise_variance_effects_model)[1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(pairwise_variance_effects_model)[3]),2)` to `r round(exp(brms::fixef(pairwise_variance_effects_model)[4]),2)`.

### 2025-10-14 R/functions/main_data_functions.R

@@ -187,6 +187,88 @@ impute_demographic_estimates <- function(data) {
     
 }
 
+create_descriptives_table <- function(data) {
+  
+  table <- data |> 
+    filter(timepoint == "baseline") |>
+    mutate(across(where(is.numeric), \(x) round(x, 2))) |>
+    mutate(
+      Age = paste0(m_age, " (", sd_age, ")"),
+      `Body mass` = paste0(m_body_mass, " (", sd_body_mass, ")"),
+      Height = paste0(m_height, " (", sd_height, ")"),
+      BMI = paste0(m_bmi, " (", sd_bmi, ")"),
+      `Fat mass` = paste0(m_fat_mass, " (", sd_fat_mass, ")"),
+      `Fat free mass` = paste0(m_fat_free_mass, " (", sd_fat_free_mass, ")")
+    ) |>
+    select(authors, year, title, cond, n,
+           Age, 
+           `Body mass`,
+           Height,
+           BMI,
+           `Fat mass`,
+           `Fat free mass`,
+           race,
+           physical_activity_level,
+           country,
+           -c(m_age, sd_age, 
+              m_body_mass, sd_body_mass,
+              m_height, sd_height,
+              m_bmi, sd_bmi,
+              m_fat_mass, sd_fat_mass,
+              m_fat_free_mass, sd_fat_free_mass),
+           insulin_resistant_description,
+           m_body_mass_estim,
+           m_height_estim,
+           m_bmi_estim
+    )|>
+    rename(
+      Authors = "authors",
+      Year = "year",
+      `Article title` = "title",
+      `Country of study` = "country",
+      Condition = "cond",
+      `Sample size` = "n",
+      `Race/Ethnicity` = "race",
+      `Physical activity` = "physical_activity_level",
+      `Metabolic health` = "insulin_resistant_description"
+    ) |>
+    mutate(Year = as.numeric(Year)) |>
+    arrange(Year) |>
+    # kable() |>
+    # collapse_rows(columns = 1, valign = "top") |>
+    # # footnote(general = c("ST = self-talk", "CON = non-intervention control")) |>
+    # row_spec(0, bold = TRUE) |>
+    # kable_classic(full_width = FALSE) 
+    flextable() |>
+    bold(part = "header") |>
+    merge_v(j = "Authors") |>
+    colformat_num(j = "Year", big.mark = "", digits = 0) |>
+    autofit() |>
+    fontsize(size = 8) |>
[…]

### 2025-10-14 _targets.R

@@ -16,7 +16,9 @@ tar_option_set(
     "brms",
     "marginaleffects",
     "tidybayes",
-    "patchwork"
+    "patchwork",
+    "flextable",
+    "officer"
   ),
   memory = "transient",
   format = "qs",
@@ -129,6 +131,11 @@ list(
   tar_target(
     main_arm_data_imputed_demographics,
     impute_demographic_estimates(main_arm_data)
+  ),
+  
+  tar_target(
+    descriptives_table_docx,
+    create_descriptives_table(main_arm_data_imputed_demographics)
   ),
   
   # Fitting main analysis models

### 2025-10-20 manuscript/manuscript_preprint.qmd
tried installing aft but deleted - will sort later

--- 2025-10-14 methods_results.qmd
@@
 ---
-title: "Methods and Results"
-format: docx
+title: |
+  Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
+format:
+  pdf: default
+  aft-pdf:
+    keep-tex: true
+author:
+  - name: James Steele
+    affiliations:
+      - ref: 1
+      - ref: 2
+      - ref: 3
+    orcid: 0000-0002-8003-0757
+    corresponding: true
+    email: james@steele-research.com
+  - name: James Fisher
+    affiliations:
+      - ref: 3
+      - ref: 4
+    orcid: 0000-0002-6013-8402
+    email: james.fisher.phd@outlook.com
+  - name: Tim Dettmann
+    affiliations:
+      - ref: 1
+    email: tim.dettmann@gmail.com
+affiliations:
+      - id: 1
+        name: Kieser Australia, Melbourne, Australia
+      - id: 2
+        name: Steele Research Limited, Eastleigh, UK
+      - id: 3
+        name: Department of Sport and Health, Solent University, Southampton, UK
+      - id: 4
+        name: The Exercise Coach, Tennessee, USA
+abstract: |
+  To write
+pdf-engine: xelatex
 bibliography: 

### 2025-10-20 manuscript/manuscript_preprint.qmd
update mylibrary.bib file, update eqs in manuscript, forgot to commit message previous intro add

--- 2025-10-14 methods_results.qmd
@@
 ---
-title: "Methods and Results"
-format: docx
+title: |
+  Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
+format:
+  pdf: default
+  aft-pdf:
+    keep-tex: true
+author:
+  - name: Richard Kirwan
+  - name: Leigh Peele
+  - name: Greg Nuckols
+  - name: Georgia Kohlhoff
+  - name: Hannah Cabré
+  - name: Alyssa Olenick
+  - name: James Steele
+abstract: |
+  Background: Polycystic ovary syndrome is the most common endocrine condition affecting women of reproductive age.
+Overweight and obesity in women with PCOS is commonly deemed a result of a lower basal metabolic rate (BMR), which has influenced lifestyle-based interventions in this group.
+Evidence supporting this reduced, however, is not consistent.
+Objectives: To systematically search and meta-analyse the existing literature to determine if resting energy expenditure differs between women with and without polycystic ovary syndrome (PCOS).
+Design: A systematic search was conducted using PubMed, Medline and Web of Science databases from January 1990 to January 2025.
+Studies that assessed resting energy expenditure in women with and without PCOS were included.
+Results: 18 studies were identified.
+Study populations ranged from X to X participants, with mean ages ranging from X to X years.
+Conclusions: TO ADD
+keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
+pdf-engine: xelatex
 bibliography: 
@@
 
+# Introduction
+Polycystic ovary syndrome (PCOS) affects approximately 10% of women of reproductive age worldwide, making it the most common endocrine disorder affecting this population [@salariGlobalPrevalencePolycystic2024].
+Due to a number of factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015].
+Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into these BMI categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
+
+The elevated incidence of overweight and obesity in PCOS is likely multifactorial with proposed mechanisms including blunted postprandial appetite hormone responses leading to reduced satiety and increased food cravings [@hirschbergImpairedCholecystokininSecretion2004; @moranGhrelinMeasuresSatiety2004; @stefanakiFoodCravingsObesity2024] and a reduced basal metabolic rate (BMR) [@georgopoulosBasalMetabolicRate2009].
+Indeed, the relevance of this reduced BMR, has become widely accepted amongst the scientific and medical community, with the original publication which reported this from Georgopoulos et al.
+[-@georgopoulosBasalMetabolicRate2009] having 109 citations according to Google Scholar metrics, as of writing (Add date and link!!!!).
+Furthermore, this publication has been widely disseminated amongst the lay community interested in PCOS, particularly in relation to lifestyle interventions, featuring in the top seven Google search results as of writing.
+However, despite the wide acceptance of this reported reduced BMR in PCOS, multiple other publications have found no such difference between women with and without PCOS [@graffSaturatedFatIntake2017; @larssonDietaryIntakeResting2016; @segalRestingMetabolicRate1990].
+
+The consequences of this wide acceptance of a reduced BMR in women with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
+It has been well documented that women with PCOS suffer with anxiety, depression and a lower quality of life (QOL) as a result of negative body image due to, amongst other reasons, body weight concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023].
+Consequently, women with PCOS have been observed to engage in more frequent weight-loss attempts than those without [@pesonenPolycysticOvarySyndrome2023].
+Furthermore, intentional weight-loss is recognised as efficacious for improving the symptoms of PCOS [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and, as such, is often recommended [@teedeRecommendations2023International2023].
+However, while weight-loss is frequently pursued for body image and/or health-related reasons, women with PCOS may be discouraged from doing so due to the commonly held belief of lower BMR making weight loss attempts more difficult.
+
+Therefore, to determine if there is indeed a difference in resting energy expenditure (REE) between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
+
 # Methods
-This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024.
+This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@liberatiPRISMAStatementReporting2009].
 The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".
@@
 Searches were limited to publications up until May 2025 when the search was completed, limited to English language articles, and Rayyan was used to manage the search and screening process.
+Two reviewers (RK and GK) independently screened all titles and abstracts against the predefined inclusion and exclusion criteria.
+Articles deemed potentially eligible by either reviewer were retrieved in full text.
+Full texts were then independently assessed by RK and LP to determine final eligibility.
+Any disagreements at either stage were resolved through discussion, and when consensus could not be reached, a third reviewer acted as an adjudicator.
 
@@
 
-<!-- ## Risk of bias (quality) assessment -->
-<!-- The assessment will be performed at the study level with the Joanna Briggs Institute (JBI) Critical Appraisal Checklist for Analytical Cross Sectional Studies, which assesses the methodological quality of a study and determines the extent to which a study has addressed the possibility of bias in its design, conduct and analysis.
--->
-
 ## Statistical Analysis
@@
 $$
+\small
 \begin{align*}
@@
 $$
-
 where the prior for $\beta_{0}$, which corresponded to the model intercept and mean REE in the control condition, was set based on meta-analysis of the mean REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this.
@@
 $$
+
 where $\log(\hat{\theta}_{ijkl})$ is the $l\textrm{th}$ natural logarithm of the standard deviation of REE estimate from the $k\textrm{th}$ arm, for the $j\textrm{th}$ study, conducted by the $i\textrm{th}$ lab and $\sigma^2_{ijkl}$ is the corresponding sampling error for that estimate.
@@
 $$
+
 where the prior for $\beta_{0}$, which corresponded to the model intercept and standard deviation of REE in the control condition, was again set based on meta-analysis of the standard deviation of REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this.

### 2025-10-20 manuscript/pre_print.qmd
deleted and cleaned up to have qmd render to pdf with equations working

--- 2025-10-20 manuscript/manuscript_preprint.qmd
@@
   Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
-format:
-  pdf: default
-  aft-pdf:
-    keep-tex: true
+date: 'format(Sys.time(), "%b %d, %Y")'
+format: pdf
 author:
   - name: Richard Kirwan
+    affiliations:
+      - ref: 1
+    orcid: 0000-0003-4645-0077
+    corresponding: true
+    email: R.P.Kirwan@ljmu.ac.uk
   - name: Leigh Peele
+    affiliations:
+      - ref: 2
+    email: leigh@macrofactorapp.com
   - name: Greg Nuckols
+    affiliations:
+      - ref: 2
+    orcid: 0009-0004-2679-7967
+    email: greg@macrofactorapp.com
   - name: Georgia Kohlhoff
+    affiliations:
+      - ref: 3
+    # orcid: 0000-0002-8003-0757
+    email: g_kohlhoff@hotmail.co.uk
   - name: Hannah Cabré
+    affiliations:
+      - ref: 4
+    orcid: 0000-0002-3235-6867
+    email: Hannah.Cabre@pbrc.edu
   - name: Alyssa Olenick
+    affiliations:
+      - ref: 5
+    orcid: 0000-0002-2217-8067
+    corresponding: true
+    email: alyssaolenickphd@gmail.com
   - name: James Steele
+    affiliations:
+      - ref: 6
+      - ref: 2
+      - ref: 7
+    orcid: 0000-0002-8003-0757
+    corresponding: true
+    email: james@steele-research.com
+affiliations:
+  - id: 1
+    name: Research Institute for Sports and Exercise Sciences, Liverpool John Moores Universtiy, Liverpool, UK
+  - id: 2
+    name: MacroFactor, Stronger by Science LLC, Raleigh, North Carolina, USA
+  - id: 3
+    name: Flourishing Health, Liverpool, UK
+  - id: 4
+    name: Reproductive Endocrinology and Women’s Health Research Program, Pennington Biomedical Research Center, Baton Rouge, Louisiana, USA
+  - id: 5
+    name: Dr Alyssa Olenick LLC
+  - id: 6
+    name: Steele Research Limited, Eastleigh, Hampshire, UK
+  - id: 7
+    name: School of Health and Biomedical Sciences, Royal Melbourne Institute of Technology, Melbourne, VIC, Australia
 abstract: |
@@
 keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
-pdf-engine: xelatex
 bibliography: 
@@
 $$
-\small
-\begin{align*}
+\scriptsize
+\begin{aligned}
 \hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
-
 \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl]
@@
        + \alpha_{0,\mathrm{effect}[l]} \\[10pt]
-
 \begin{pmatrix}
@@
 \end{pmatrix}
-
 \text{, for lab i = 1,} \dots \text{,I} \\[12pt]
-
 \begin{pmatrix}
@@
 \text{, for study j = 1,} \dots \text{,J} \\[12pt]
-
 \alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})
@@
 \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
-\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
-
-\end{align*}
+\text{, for effect l = 1,} \dots \text{,L}
+\end{aligned}
 $$
@@
 \small
-\begin{align*}
+\begin{aligned}
 \beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
 \beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]
-
 \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
@@
 \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]
-
 \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
 \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\small
-\begin{align*}
+\scriptsize
+\begin{aligned}
 \log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
-
 \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]}
@@
        + \alpha_{0,\mathrm{effect}[l]} \\[8pt]
-
 m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt]
 \tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt]
-
 \begin{pmatrix}
@@
 \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
-\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
-\end{align*}
+\text{, for effect l = 1,} \dots \text{,L}
+\end{aligned}
 $$
@@
 \small
-\begin{align*}
+\begin{aligned}
 \beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt]
@@
 \beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt]
-
 \text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt]
 \text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt]
-
 \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt]
@@
 \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt]
-
 \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
 \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
-\end{align*}
-
+\end{aligned}
 $$
@@
 \small
-\begin{align*}
+\begin{aligned}
 \hat{\theta}_{ij} &\sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij}) \\[6pt]
-
 \mu_{ij} &= \beta_0
@@
        + \alpha_{0,\mathrm{study}[j]} \\[10pt]
-
 \alpha_{0,\mathrm{lab}[i]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}})
@@
 \alpha_{0,\mathrm{study}[j]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}})
-\text{, for study j = 1,} \dots \text{,J} \\[12pt]
-
-\end{align*}
+\text{, for study j = 1,} \dots \text{,J}
+\end{aligned}
 $$
@@
 #| fig-width: 10
-#| fig-height: 5
+#| fig-height: 7.5
 #| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.
@@
 #| fig-width: 10
-#| fig-height: 5
+#| fig-height: 7.5
 #| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.
@@
 This was similar for the standard deviation of REE with the pairwise model resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(pairwise_variance_effects_model)[1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(pairwise_variance_effects_model)[3]),2)` to `r round(exp(brms::fixef(pairwise_variance_effects_model)[4]),2)`.
+
+# References

### 2025-10-21 R/functions/main_data_functions.R
Updated demographics table functions for systematic review results.

@@ -189,16 +189,58 @@ impute_demographic_estimates <- function(data) {
 
 create_descriptives_table <- function(data) {
   
-  table <- data |> 
+  table <- main_arm_data_imputed_demographics |> 
     filter(timepoint == "baseline") |>
     mutate(across(where(is.numeric), \(x) round(x, 2))) |>
     mutate(
-      Age = paste0(m_age, " (", sd_age, ")"),
-      `Body mass` = paste0(m_body_mass, " (", sd_body_mass, ")"),
-      Height = paste0(m_height, " (", sd_height, ")"),
-      BMI = paste0(m_bmi, " (", sd_bmi, ")"),
-      `Fat mass` = paste0(m_fat_mass, " (", sd_fat_mass, ")"),
-      `Fat free mass` = paste0(m_fat_free_mass, " (", sd_fat_free_mass, ")")
+      authors = paste0(authors, " (", year, ")"),
+      Age = if_else(
+        is.na(m_age) & is.na(sd_age),
+        "",
+        paste0(m_age, " (", sd_age, ")")
+      ),
+      `Body mass` = if_else(
+        is.na(m_body_mass) & is.na(sd_body_mass),
+        "",
+        paste0(
+          if_else(!is.na(m_body_mass), as.character(m_body_mass), ""),
+          if_else(m_body_mass_estim == "y", "\u002A", ""),
+          if_else(!is.na(sd_body_mass), paste0(" (", sd_body_mass, ")"), "")
+        )
+      ),
+      Height = if_else(
+        is.na(m_height) & is.na(sd_height),
+        "",
+        paste0(
+          if_else(!is.na(m_height), as.character(m_height), ""),
+          if_else(m_height_estim == "y", "\u002A", ""),
+          if_else(!is.na(sd_height), paste0(" (", sd_height, ")"), "")
+        )
+      ),
+      BMI = if_else(
+        is.na(m_bmi) & is.na(sd_bmi),
+        "",
+        paste0(
+          if_else(!is.na(m_bmi), as.character(m_bmi), ""),
+          if_else(m_bmi_estim == "y", "\u002A", ""),
+          if_else(!is.na(sd_bmi), paste0(" (", sd_bmi, ")"), "")
+        )
+      ),
+      `Fat mass` = if_else(
+        is.na(m_fat_mass) & is.na(sd_fat_mass),
+        "",
+        paste0(m_fat_mass, " (", sd_fat_mass, ")")
[…]

### 2025-10-21 manuscript/pre_print.qmd
Added to data all median, se, range, iqr cols (need to carefully document all extraction)

@@
 date: 'format(Sys.time(), "%b %d, %Y")'
-format: pdf
+format:
+  pdf:
+    include-in-header:
+      text: |
+       \usepackage[font=scriptsize]{caption}
+       \usepackage{lscape}
+        \newcommand{\blandscape}{\begin{landscape}}
+        \newcommand{\elandscape}{\end{landscape}}
 author:
@@
 # Results
+## Systematic Review
 ```{r}
@@
 A total of `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` labs including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
-Descriptive characteristics of the arms and participants in these studies are reported in Table 1.
+Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
+
+\newpage
+
+\blandscape
+
+```{r}
+#| label: tbl-review 
+#| tbl-cap: Descriptive characteristics of arms and participants for included studies
+
+main_arm_data_imputed_demographics
+```
+
+\elandscape
 

### 2025-10-21 _targets.R

@@ -134,6 +134,11 @@ list(
   ),
   
   tar_target(
+    descriptives_table,
+    create_descriptives_table(main_arm_data_imputed_demographics)
+  ),
+  
+  tar_target(
     descriptives_table_docx,
     create_descriptives_table(main_arm_data_imputed_demographics)
   ),

### 2025-10-23 R/functions/main_data_functions.R
prepare_data() now calculating all demographics from exactly reported summary stats

@@ -18,8 +18,63 @@ prepare_data <- function(file) {
     )
   ) |>
     
+    # add sample size for demographics where this is the same as for REE
+    mutate(
+      n_age = case_when(
+        is.na(n_age) ~ n,
+        .default = n_age
+      ),
+      n_body_mass = case_when(
+        is.na(n_body_mass) ~ n,
+        .default = n_body_mass
+      ),
+      n_fat_mass = case_when(
+        is.na(n_fat_mass) ~ n,
+        .default = n_fat_mass
+      ),
+      n_fat_free_mass = case_when(
+        is.na(n_fat_free_mass) ~ n,
+        .default = n_fat_free_mass
+      ),
+      n_height = case_when(
+        is.na(n_height) ~ n,
+        .default = n_height
+      ),
+      n_bmi = case_when(
+        is.na(n_bmi) ~ n,
+        .default = n_bmi
+      )
+    ) |>
+    
     # convert standard errors to standard deviations
     mutate(
+      # demographics
+      sd_age = case_when(
+        is.na(sd_age) == TRUE ~ se_age * sqrt(n_age),
+        .default = sd_age
+      ),
+      sd_body_mass = case_when(
+        is.na(sd_body_mass) == TRUE ~ se_body_mass * sqrt(n_body_mass),
+        .default = sd_body_mass
+      ),
+      sd_fat_mass = case_when(
+        is.na(sd_fat_mass) == TRUE ~ se_fat_mass * sqrt(n_fat_mass),
+        .default = sd_fat_mass
+      ),
+      sd_fat_free_mass = case_when(
+        is.na(sd_fat_free_mass) == TRUE ~ se_fat_free_mass * sqrt(n_fat_free_mass),
+        .default = sd_fat_free_mass
+      ),
+      sd_height = case_when(
+        is.na(sd_height) == TRUE ~ se_height * sqrt(n_height),
+        .default = sd_height
+      ),
+      sd_bmi = case_when(
+        is.na(sd_bmi) == TRUE ~ se_bmi * sqrt(n_bmi),
+        .default = sd_bmi
+      ),
+      
+      # REE variables
       sd = case_when(
         is.na(sd) == TRUE ~ se * sqrt(n),
         .default = sd
@@ -34,25 +89,194 @@ prepare_data <- function(file) {
[…]

### 2025-10-23 R/functions/main_data_functions.R
rename impute_demographic_estimates to impute_bmi_estimates

@@ -401,7 +401,7 @@ calculate_arm_effects <- function(data) {
   return(data)
 }
 
-impute_demographic_estimates <- function(data) {
+impute_bmi_estimates <- function(data) {
   data <- data |>
     
     # convert height to cm for reporting

### 2025-10-23 _targets.R

@@ -130,7 +130,7 @@ list(
   
   tar_target(
     main_arm_data_imputed_demographics,
-    impute_demographic_estimates(main_arm_data)
+    impute_bmi_estimates(main_arm_data)
   ),
   
   tar_target(

### 2025-10-23 R/functions/main_data_functions.R
create_descriptives_table() has been updated to include device/diagnosis

@@ -450,7 +450,7 @@ impute_bmi_estimates <- function(data) {
 
 create_descriptives_table <- function(data) {
   
-  table <- main_arm_data_imputed_demographics |> 
+  table <- data |> 
     filter(timepoint == "baseline") |>
     mutate(across(where(is.numeric), \(x) round(x, 2))) |>
     mutate(
@@ -501,9 +501,10 @@ create_descriptives_table <- function(data) {
         method == "indirect_calorimetry" ~ "Indirect Calorimetry",
         method == "doubly_labelled_water" ~ "Doubly Labelled Water",
         .default = method
-      )
+      ),
+      method = paste0(method, ": ", device)
     ) |>
-    select(authors, year, title, cond, n,
+    select(authors, year, title, cond, diagnostic_criteria, n,
            Age, 
            `Body mass`,
            Height,
@@ -520,13 +521,14 @@ create_descriptives_table <- function(data) {
               m_fat_mass, sd_fat_mass,
               m_fat_free_mass, sd_fat_free_mass),
            insulin_resistant_description,
-           method
+           method,
     )|>
     rename(
       Authors = "authors",
       `Article title` = "title",
       `Country of study` = "country",
       Condition = "cond",
+      `Diagnostic criteria` = diagnostic_criteria,
       `Sample size` = "n",
       `Race/Ethnicity` = "race",
       `Physical activity` = "physical_activity_level",
@@ -589,7 +591,7 @@ create_descriptives_table <- function(data) {
       value = as_paragraph(
         c(
           "PCOS = polycystic ovary syndrome; BMI = body mass index; OGTT = oral glucose tolerance test; HOMA-IR = homeostatic model assessment of insulin resistance",
-          "Values are Mean (SD) unless otherwise specified",
+          "Values are Mean (SD) unless otherwise specified; note, some have been calculated/estimated from corresponding standard error, range, iqr, median, and sample size (see data and code)",
           "Indicates that this mean was estimated from the corresponding means for body mass/height/BMI"
         )
       ),

### 2025-10-23 _targets.R
Pipeline rerun with no-greek/baseline sensitivity models

@@ -285,6 +285,83 @@ list(
     fit_pairwise_variance_model(
       pairwise_data_effects
     )
+  ),
+  
+  #### Sensitivity analysis with only baseline data ----
+  
+  # Fitting baseline arm analysis models
+  
+  tar_target(
+    baseline_arm_mean_effects_model,
+    fit_arm_mean_effects_model(
+      main_arm_data_effects |>
+        filter(timepoint == "baseline"), # just take baseline
+      prior_arm_mean_effects
+    )
+  ),
+  
+  tar_target(
+    baseline_arm_variance_effects_model,
+    fit_arm_variance_effects_model(
+      main_arm_data_effects |>
+        filter(timepoint == "baseline"), # just take baseline
+      prior_arm_variance_effects
+    )
+  ),
+  
+  # Get contrasts from baseline arm models
+  
+  
+  tar_target(
+    baseline_arm_mean_effects_contrast_condition,
+    get_mean_contrast_condition(baseline_arm_mean_effects_model)
+  ),
+  
+  tar_target(
+    baseline_arm_variance_effects_contrast_condition,
+    get_variance_contrast_condition(baseline_arm_variance_effects_model,
+                                    main_arm_data_effects |>
+                                      filter(timepoint == "baseline") # just take baseline
+    )
+  ),
+  
+  #### Sensitivity analysis removing Greek lab studies ----
+  
+  
+  # Fitting baseline arm analysis models
+  
+  tar_target(
+    no_greek_arm_mean_effects_model,
+    fit_arm_mean_effects_model(
+      main_arm_data_effects |>
+        filter(lab != 8), # remove Greek studies
+      prior_arm_mean_effects
+    )
+  ),
+  
+  tar_target(
+    no_greek_arm_variance_effects_model,
+    fit_arm_variance_effects_model(
+      main_arm_data_effects |>
+        filter(lab != 8), # remove Greek studies
+      prior_arm_variance_effects
+    )
+  ),
+  
+  # Get contrasts from no_greek arm models
+  
+  
+  tar_target(
+    no_greek_arm_mean_effects_contrast_condition,
+    get_mean_contrast_condition(no_greek_arm_mean_effects_model)
+  ),
+  
+  tar_target(
+    no_greek_arm_variance_effects_contrast_condition,
+    get_variance_contrast_condition(no_greek_arm_variance_effects_model,
+                                    main_arm_data_effects |>
+                                      filter(lab != 8) # remove Greek studies
+    )
   )
   
   

### 2025-10-23 manuscript/pre_print.qmd
Update to quarto last modified date and long format

@@
   Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
-date: 'format(Sys.time(), "%b %d, %Y")'
+date: last-modified
+date-format: long
 format:

### 2025-10-23 manuscript/pre_print.qmd
Hannah's feedback on intro/methods addressed

@@
 abstract: |
-  Background: Polycystic ovary syndrome is the most common endocrine condition affecting women of reproductive age.
-Overweight and obesity in women with PCOS is commonly deemed a result of a lower basal metabolic rate (BMR), which has influenced lifestyle-based interventions in this group.
-Evidence supporting this reduced, however, is not consistent.
-Objectives: To systematically search and meta-analyse the existing literature to determine if resting energy expenditure differs between women with and without polycystic ovary syndrome (PCOS).
-Design: A systematic search was conducted using PubMed, Medline and Web of Science databases from January 1990 to January 2025.
-Studies that assessed resting energy expenditure in women with and without PCOS were included.
-Results: 18 studies were identified.
+  Background: Polycystic ovary syndrome (PCOS) is the most common endocrine condition affecting women of reproductive age.
+Women who are living with PCOS are also more likely to experience overwweight and obesity and this is assumed to be a result of a lower resting energy expenditure (REE).
+The assumption that REE is lower in women living with PCOS has also influenced lifestyle-based interventions in this group.
+Evidence supporting this reduced REE compared to women without PCOS, however, is not consistent.
+Objectives: To systematically search and meta-analyse the existing literature to determine if resting energy expenditure differs between women with and without PCOS.
+Design: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.
+Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included.
+Results: Seventeen independent studies were identified.
 Study populations ranged from X to X participants, with mean ages ranging from X to X years.
@@
 Due to a number of factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015].
-Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into these BMI categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
+Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into overweight or obese BMI categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
 
-The elevated incidence of overweight and obesity in PCOS is likely multifactorial with proposed mechanisms including blunted postprandial appetite hormone responses leading to reduced satiety and increased food cravings [@hirschbergImpairedCholecystokininSecretion2004; @moranGhrelinMeasuresSatiety2004; @stefanakiFoodCravingsObesity2024] and a reduced basal metabolic rate (BMR) [@georgopoulosBasalMetabolicRate2009].
-Indeed, the relevance of this reduced BMR, has become widely accepted amongst the scientific and medical community, with the original publication which reported this from Georgopoulos et al.
-[-@georgopoulosBasalMetabolicRate2009] having 109 citations according to Google Scholar metrics, as of writing (Add date and link!!!!).
-Furthermore, this publication has been widely disseminated amongst the lay community interested in PCOS, particularly in relation to lifestyle interventions, featuring in the top seven Google search results as of writing.
-However, despite the wide acceptance of this reported reduced BMR in PCOS, multiple other publications have found no such difference between women with and without PCOS [@graffSaturatedFatIntake2017; @larssonDietaryIntakeResting2016; @segalRestingMetabolicRate1990].
+The elevated incidence of overweight and obesity in PCOS is likely multifactorial with proposed mechanisms including blunted postprandial appetite hormone responses leading to reduced satiety and increased food cravings [@hirschbergImpairedCholecystokininSecretion2004; @moranGhrelinMeasuresSatiety2004; @stefanakiFoodCravingsObesity2024] and a reduced resting energy expenditure (REE) [@georgopoulosBasalMetabolicRate2009].
+Indeed, the study from Georgopoulos et al.
+[-@georgopoulosBasalMetabolicRate2009] examining REE using indirect calorimetry in women with PCOS reported that they had a significantly lower REE, ~400 kcal per day less, compared with women without PCOS.
+Notably, they also reported insulin resistance further influenced REE among women with PCOS, with those who were insulin resistant exhibiting an additional reduction of about ~500 kcal per day compared to women with PCOS who were not insulin resistant [@georgopoulosBasalMetabolicRate2009].
+Other studies similarly report lower BMR using indirect methods (such as prediction from bioelectrical impedance analysis or accelerometer physical activity data) in women with PCOS, also highlighting factors like insulin resistance or BMI category influence REE [@romualdiRestingMetabolicRate2019; churchillBasalMetabolicRate2015].
+However, despite the widespread acceptance that women living with PCOS exhibit reduced REE based on studies such as these, other research has reported little to no difference in REE between women with and without PCOS [@graffSaturatedFatIntake2017; @larssonDietaryIntakeResting2016; @segalRestingMetabolicRate1990].
 
-The consequences of this wide acceptance of a reduced BMR in women with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
-It has been well documented that women with PCOS suffer with anxiety, depression and a lower quality of life (QOL) as a result of negative body image due to, amongst other reasons, body weight concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023].
-Consequently, women with PCOS have been observed to engage in more frequent weight-loss attempts than those without [@pesonenPolycysticOvarySyndrome2023].
-Furthermore, intentional weight-loss is recognised as efficacious for improving the symptoms of PCOS [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and, as such, is often recommended [@teedeRecommendations2023International2023].
-However, while weight-loss is frequently pursued for body image and/or health-related reasons, women with PCOS may be discouraged from doing so due to the commonly held belief of lower BMR making weight loss attempts more difficult.
+The consequences of widespread acceptance that REE is reduced in women living with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
+From a physiological perspective, if women with PCOS do exhibit a lower REE, this could imply a meaningful metabolic disadvantage that may influence dietary and nutritional guidance for weight management.
+However, if no such difference exists, the belief in a “slower metabolism” could instead serve as a deterrent to weight-loss efforts—efforts that are recognised as efficacious for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and are routinely recommended [@teedeRecommendations2023International2023].
+It has been well documented that women with PCOS already experience higher rates of anxiety, depression, and lower quality of life (QOL) as a result of negative body image and weight-related concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023].
+Consequently, they are often observed to engage in more frequent weight-loss attempts than women without PCOS [@pesonenPolycysticOvarySyndrome2023].
+Clarifying the relationship between REE and PCOS may therefore help guide more accurate clinical recommendations and empower both practitioners and women with PCOS.
 
-Therefore, to determine if there is indeed a difference in resting energy expenditure (REE) between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
+Therefore, to determine if there is indeed a difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
 
 # Methods
-This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@liberatiPRISMAStatementReporting2009].
+This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@pagePRISMA2020Statement2021a].
 The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".
@@
 ## Eligibility Criteria
-The condition being studied was PCOS and we included observational cross-sectional design studies, in addition to studies of intervention where REE was reported for the population (and if included, the comparator) condition of interest.
-Following the PICO framework our eligibility criteria were as follows:
+Studies were included in the systematic review if 1) participants were confirmed as women with PCOS between the ages of 18 to 65 years of age with or without insulin resistance; 2) otherwise healthy (e.g., non-diabetic, no cardiovascular disease); 3) had a measure of REE measured via multiple methods including direct/indirect calorimetry, doubly labelled water; and 4) trials were not retracted at the time of this analysis.
+Studies were excluded if they 1) used invalid or non-standard methods for measuring REE (e.g., predicted REE from body composition or accelerometer data); 2) non-peer-reviewed journal articles (including grey literature sources such as conference abstracts, theses and dissertations); and 3) were secondary analyses with the same primary outcome data as another included study.
+
+The condition being studied was PCOS and we included observational cross-sectional design studies, in addition to studies of intervention where REE was reported for the population (and if present, the comparator i.e., women without PCOS) condition of interest.
+For clarity, studies of any design were included if they reported the REE using the methods indicated for a sample of adult women with PCOS and who were otherwise healthy.
+This included both studies with and without samples of healthy control women without PCOS.
+As detailed in the statistical analysis section below, a Bayesian model with informative priors based on normative data for REE in health women without PCOS was included to provide control information indirectly where this was missing.
+The use of such priors is an efficient tool to incorporate historical information on a particular population in a conservative manner [@weberApplyingMetaAnalyticPredictivePriors2021].
+
+Following the PICO framework our eligibility criteria can be defined as follows:
 
@@
     - With or without insulin resistance (IS)
-  + Exclusion criteria:
-    -Studies using invalid or non-standard methods for measuring REE (e.g., predicted REE from body composition)
 
@@
 * Outcome
- + Resting energy expenditure (REE) measured via multiple methods including direct/indirect calorimetry, doubly labelled water.
-
-Studies using invalid or non-standard methods for measuring BMR (e.g., predicted BMR from body composition) were excluded.
+ + Inclusion criteria
+    - Resting energy expenditure (REE) measured via multiple methods including direct/indirect calorimetry, doubly labelled water.
+ + Exclusion criteria:
+    - Studies using invalid or non-standard methods for measuring REE (e.g., predicted REE from body composition or accelerometer data)
 

### 2025-10-23 manuscript/pre_print.qmd
Alyssa/Hannah's comments addressed, plus sensitivity analyses expanded in method

@@
 Due to a number of factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015].
-Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into overweight or obese BMI categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
+Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into overweight or obese body mass index (BMI) categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
 
@@
 
-Therefore, to determine if there is indeed a difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
+Therefore, to estimate the magnitude of difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
 
@@
 The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".
+We summarise and describe the studies in addition to quantitatively synthesising their results via meta-analysis.
 
@@
 Bibliometric data including authors, journal, and title of article were extracted.
-Descriptive statistics for age, body mass, fat mass, fat free mass, height, BMI, race, physical activity levels, country of investigation, information regarding glucose/insulin regulation and insulin resistance status (where available) were extracted for each arm within each study in addition to sample size.
+Descriptive statistics for age, body mass, fat mass, fat free mass, height, BMI, race, physical activity levels, country of investigation, information regarding glucose/insulin regulation and insulin resistance status (where available), diagnostic criteria for PCOS, and measurement method and device were extracted for each arm within each study in addition to sample size.
 Descriptive characteristics were then tabulated across studies for reporting.
 
-The method of measuring REE was extracted and the units of measurement for which REE was reported.
 For each arm, and observation time point if multiple observations reported (e.g., before and after an intervention), depending on what was reported by the authors we extracted the means, medians, standard deviations, standard errors, lower and upper range values, and interquartile range for the unadjusted and/or body mass adjusted and/or fat free mass adjusted REE values.
@@
 [-@wanEstimatingSampleMean2014].
-All REE values were converted to kcal/day.
+Further, where missing, height/body mass/BMI where estimated based on the means reported for these.
+The units of measurement for which REE was extracted and all REE values were converted to kcal/day.
+In one case [@pohlmeierEffectLowstarchLowdairy2014] REE was reported relative to body mass and the unadjusted values were no longer available (confirmed by the authors) as such in this case we used the mean body mass to convert back to estimated REE unadjusted.
+
+### Studies with possible reporting errors
+During data extraction it was noted that several studies from the same lab/research group [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009] contained a number of discrepancies that seemed to be possible reporting errors.
+This included, based on taking the authors results as written, standard errors that implied impossible or at least incredibly unlikely standard deviations, and discrepancies in sample size reporting throughout for most variables without explanation or where this was explained the sample sizes were discrepant with the text Further, data was not reported for the healthy control women without PCOS in three of the studies [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009], and REE was reported as an "adjusted" value whereby $\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})$ and $\textrm{REE}_{predicted}$ was obtained by substituting the individual lean body mass, fat mass, gender, and age in the linear regression equation generated by the data of all patients.
+In correspondence with the senior author we were unable to clarify the reporting discrepancies as the person responsible for the data/results was no longer contactable.
+The original data were also no longer available and so we could not calculated the unadjusted REE.
+
+Given these issues we made a decision to extract the results from these studies as reported, but to conduct analyses both with and without their inclusion.
+Though not pre-registered, due to a lack of confidence in the results as reported, we decided to include the analysis omitting these studies as our main models in the results reported below.
+The results of the analysis including them are however reported in the sensitivity analysis section.
 
@@
 
-We adopted an arm-based multiple condition comparison (i.e., network) type model given that the studies included had arms of women with PCOS both with, and without, a non-PCOS control arm [@hongBayesianMissingData2016], and also in some cases multiple observations of REE in the different arms included in the study.
+We adopted an arm-based multiple condition comparison (i.e., network) type model given that the studies included had arms of women with PCOS both with, and without, a non-PCOS control arm [@hongBayesianMissingData2016], and also in some cases multiple observations of REE in the different arms included in the study (for example, where an intervention was conducted and pre- and post-intervention REE was reported).
 In typical contrast-based meta-analyses data is limited to the effect sizes for paired contrasts between arms and thus studies that include both arms (i.e., relative effects between non-PCOS control vs PCOS arms); however, in arm-based analyses the data are the absolute effects within each arm and information is borrowed across studies to enable both within condition absolute, and between condition relative contrasts to be estimated.
+We made use of historical information regarding REE in healthy control women without PCOS by setting informative priors based on meta-analysis of large scale studies reporting normative data for REE in this population.
+This was included to provide control information indirectly where this was missing from particular studies.
+The use of historical priors like this is an efficient tool to incorporate historical information on a particular population in a conservative manner in meta-analyses [@weberApplyingMetaAnalyticPredictivePriors2021].
 From this model we focus on reporting the global grand mean estimate for the fixed between condition relative contrast for non-PCOS control vs PCOS arms as our primary estimand of interest (i.e., $\beta_1$ in both mean and standard deviation models).
@@
 
-### Sensitivity analysis using pairwise contrast based models
+### Sensitivity analyses 
+#### Pairwise contrast based models
 
-By way of sensitivity analysis we also conducted pairwise contrast based models where we limited the included effects to those extracted from studies including only a directly comparable control and PCOS arm at baseline.
+By way of pre-registered sensitivity analysis we also conducted pairwise contrast based models where we limited the included effects to those extracted from studies including only a directly comparable control and PCOS arm at baseline.
 These models were both as follows:
@@
 From these models we calculated the mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for the $\beta_0$ (comparable to the $\beta_1$ from the corresponding mean and standard devation of REE arm-based models) for each effect size providing the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles given our priors and data.
+
+#### Additional sensitivity analyses
+As noted in the section above, *"Studies with possible reporting errors"*, we opted to conduct analysis with and without the inclusion of four studies with possible reporting errors we could not resolve [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009].
+Thus the main models described above were run with and without these studies, the main results are presented without them and the sensitivity results with them are reported separately below.
+
+As a further sensitivity analysis, given the inclusion of some studies with interventions in women with PCOS reporting REE at multiple timepoints [@moranShorttermMealReplacements2006; @brunerEffectsExerciseNutritional2006; @pohlmeierEffectLowstarchLowdairy2014] but the lack of control women without PCOS, we opted to also conduct sensitivity analysis excluding these and only examining cross-sectional studies.
+As such, the main models described above were run with and without these studies.
 

### 2025-10-23 manuscript/pre_print.qmd
Rejigged pipeline to have main models without Greek studies

@@
     - 18-65 y
-    - With or without insulin resistance (IS)
+    - With or without insulin resistance (IR)
 
@@
 During data extraction it was noted that several studies from the same lab/research group [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009] contained a number of discrepancies that seemed to be possible reporting errors.
-This included, based on taking the authors results as written, standard errors that implied impossible or at least incredibly unlikely standard deviations, and discrepancies in sample size reporting throughout for most variables without explanation or where this was explained the sample sizes were discrepant with the text Further, data was not reported for the healthy control women without PCOS in three of the studies [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009], and REE was reported as an "adjusted" value whereby $\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})$ and $\textrm{REE}_{predicted}$ was obtained by substituting the individual lean body mass, fat mass, gender, and age in the linear regression equation generated by the data of all patients.
+This included, based on taking the authors results as written, standard errors that implied impossible or at least incredibly unlikely standard deviations, and discrepancies in sample size reporting throughout for most variables without explanation or where this was explained the sample sizes were discrepant with the text Further, data was not reported for the healthy control women without PCOS in three of the studies [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009], and REE was reported as an "adjusted" value whereby $\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})$ and the $\textrm{REE}_{predicted}$ was obtained by substituting the individual lean body mass, fat mass, gender, and age in the linear regression equation generated by the data of all patients.
 In correspondence with the senior author we were unable to clarify the reporting discrepancies as the person responsible for the data/results was no longer contactable.

### 2025-10-23 _targets.R

@@ -120,7 +120,8 @@ list(
   
   tar_target(
     main_arm_data,
-    prepare_data(main_arm_data_file)
+    prepare_data(main_arm_data_file) |>
+      filter(lab != 8) # remove Greek studies for main model
   ),
   
   tar_target(
@@ -309,62 +310,35 @@ list(
     )
   ),
   
-  # Get contrasts from baseline arm models
-  
+  #### Sensitivity analysis including Greek lab studies ----
   
   tar_target(
-    baseline_arm_mean_effects_contrast_condition,
-    get_mean_contrast_condition(baseline_arm_mean_effects_model)
+    main_plus_greek_arm_data,
+    prepare_data(main_arm_data_file)
   ),
   
   tar_target(
-    baseline_arm_variance_effects_contrast_condition,
-    get_variance_contrast_condition(baseline_arm_variance_effects_model,
-                                    main_arm_data_effects |>
-                                      filter(timepoint == "baseline") # just take baseline
-    )
+    main_plus_greek_arm_data_effects,
+    calculate_arm_effects(main_plus_greek_arm_data)
   ),
   
-  #### Sensitivity analysis removing Greek lab studies ----
-  
-  
-  # Fitting baseline arm analysis models
+  # Fitting main arm analysis models including Greek studies
   
   tar_target(
-    no_greek_arm_mean_effects_model,
+    main_plus_greek_arm_mean_effects_model,
     fit_arm_mean_effects_model(
-      main_arm_data_effects |>
-        filter(lab != 8), # remove Greek studies
+      main_plus_greek_arm_data_effects, 
       prior_arm_mean_effects
     )
   ),
   
   tar_target(
-    no_greek_arm_variance_effects_model,
+    main_plus_greek_arm_variance_effects_model,
     fit_arm_variance_effects_model(
-      main_arm_data_effects |>
-        filter(lab != 8), # remove Greek studies
+      main_plus_greek_arm_data_effects,
       prior_arm_variance_effects
     )
-  ),
-  
-  # Get contrasts from no_greek arm models
-  
-  
-  tar_target(
-    no_greek_arm_mean_effects_contrast_condition,
-    get_mean_contrast_condition(no_greek_arm_mean_effects_model)
-  ),
-  
-  tar_target(
-    no_greek_arm_variance_effects_contrast_condition,
-    get_variance_contrast_condition(no_greek_arm_variance_effects_model,
-                                    main_arm_data_effects |>
[…]

### 2025-10-23 manuscript/pre_print.qmd
Pipeline all up to date

@@
 The random effects were set similarly scaled to the expected response values but using a $\textrm{half-student-}t$ distribution centred on zero.
-From these models we calculated the mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for the $\beta_0$ (comparable to the $\beta_1$ from the corresponding mean and standard devation of REE arm-based models) for each effect size providing the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles given our priors and data.
+From these models we calculated the mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for the $\beta_0$ (comparable to the $\beta_1$ from the corresponding mean and standard deviation of REE arm-based models) for each effect size providing the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles given our priors and data.
 
@@
 
-As a further sensitivity analysis, given the inclusion of some studies with interventions in women with PCOS reporting REE at multiple timepoints [@moranShorttermMealReplacements2006; @brunerEffectsExerciseNutritional2006; @pohlmeierEffectLowstarchLowdairy2014] but the lack of control women without PCOS, we opted to also conduct sensitivity analysis excluding these and only examining cross-sectional studies.
-As such, the main models described above were run with and without these studies.
+As a further sensitivity analysis, given the inclusion of some studies with interventions in women with PCOS reporting REE at multiple timepoints such as mid- and post-intervention [@moranShorttermMealReplacements2006; @brunerEffectsExerciseNutritional2006; @pohlmeierEffectLowstarchLowdairy2014] but lacking of control women without PCOS, we opted to also conduct sensitivity analysis excluding these and only examining their baseline results in addition to the cross-sectional studies.
+As such, the main models described above were run without the follow-up (i.e., mid- or post-intervention timepoints) from these studies and only using the baseline results in addition to other cross-sectional studies.
 
@@
 ```{r}
-targets::tar_load(main_arm_data_imputed_demographics)
+targets::tar_load(c(
+  descriptives_table,
+  main_arm_data_imputed_demographics
+))
 
@@
 Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
+Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved (they are however summarised in @tbl-review).
 
@@
 
-main_arm_data_imputed_demographics
+descriptives_table
 ```
@@
 
-## Sensitivity Analysis with Pairwise Models
+## Sensitivity Analyses
+### Pairwise Models
 ```{r}
@@
 
+### Models including studies with possible reporting issues
+```{r}
+targets::tar_load(c(main_plus_greek_arm_mean_effects_model, main_plus_greek_arm_variance_effects_model))
+
+```
+For mean REE the model including the studies noted with possible reporting issues that were not resolved [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009] still resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,1])` kcal/day with a 95% quantile interval ranging from  `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,3])` kcal/day to `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,4])` kcal/day.
+This was similar for the standard deviation of REE with the model including these studies resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,4]),2)`.
+
+### Basline REE Measurement Models
+```{r}
+targets::tar_load(c(baseline_arm_mean_effects_model, baseline_arm_variance_effects_model))
+
+```
+For mean REE the model which included only baseline REE measurements from studies involving interventions [@moranShorttermMealReplacements2006; @brunerEffectsExerciseNutritional2006; @pohlmeierEffectLowstarchLowdairy2014], in addition to other cross-sectional studies, also resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(baseline_arm_mean_effects_model)[2,1])` kcal/day with a 95% quantile interval ranging from  `r round(brms::fixef(baseline_arm_mean_effects_model)[2,3])` kcal/day to `r round(brms::fixef(baseline_arm_mean_effects_model)[2,4])` kcal/day.
+This was the case also for the standard deviation of REE with this model resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,4]),2)`.
+
+# Discussion
+
+TO WRITE
+
 # References

### 2025-10-24 R/functions/main_data_functions.R
Changed table to Latin Modern Roman

@@ -556,7 +556,9 @@ create_descriptives_table <- function(data) {
     ) |>
     bold(part = "header") |>
     autofit() |>
-    fontsize(size = 8) |>
+    fontsize(size = 8, part = "header") |>
+    fontsize(size = 8, part = "footer") |>
+    fontsize(size = 6, part = "body") |>
     align(
       j = c(
         "Condition",
@@ -574,7 +576,7 @@ create_descriptives_table <- function(data) {
       align = "center",
       part = "all"
     ) |>
-    width(j = NULL, width = 0.5) |>
+    # width(j = NULL, width = 1, unit = "cm") |>
     delete_columns(j = "is_last_row") |>
     merge_v(j = c("Authors", "Article title")) |>
     footnote(
@@ -597,7 +599,8 @@ create_descriptives_table <- function(data) {
       ),
       ref_symbols = c(" ", "†", "†*", "†*", "†*", "†", "†"),   # no superscript
       part = "header"        
-    )
+    ) |>
+    font(fontname = "Latin Modern Roman", part = "all")
   
 }
 
@@ -619,7 +622,13 @@ convert_descriptives_table_to_docx <- function(table) {
     body_end_section_landscape()
   
   # Save to file
-  print(doc, target = "descriptives_table_landscape.docx")
+  print(doc, target = "tables/descriptives_table.docx")
+}
+
+convert_descriptives_table_to_image <- function(table) {
+  
+  # Save as image
+  save_as_image(table, path = "tables/descriptives_table.svg", res=300)
 }
 
 # Pairwise data preparation for sensitivity analysis

### 2025-10-24 _targets.R

@@ -141,7 +141,12 @@ list(
   
   tar_target(
     descriptives_table_docx,
-    create_descriptives_table(main_arm_data_imputed_demographics)
+    convert_descriptives_table_to_docx(descriptives_table)
+  ),
+  
+  tar_target(
+    descriptives_table_image,
+    convert_descriptives_table_to_image(descriptives_table)
   ),
   
   # Fitting main analysis models

### 2025-10-25 manuscript/pre_print.qmd
Table as png now in manuscript

@@
   pdf:
+    # documentclass: article
+    pdf-engine: xelatex
+    mainfont: "Latin Modern Roman"
     include-in-header:
@@
 targets::tar_load(c(
-  descriptives_table,
   main_arm_data_imputed_demographics
@@
 ```
-A total of `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` labs including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
+Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
 Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
@@
 
-\newpage
-
-\blandscape
-
-```{r}
-#| label: tbl-review 
-#| tbl-cap: Descriptive characteristics of arms and participants for included studies
-
-descriptives_table
-```
-
-\elandscape
+::: {.landscape}
+![Descriptive characteristics of arms and participants for included studies](descriptives_table.png){#tbl-review}
+:::
 

### 2025-10-25 R/functions/main_data_functions.R

@@ -555,7 +555,7 @@ create_descriptives_table <- function(data) {
       border = officer::fp_border(color = "black")
     ) |>
     bold(part = "header") |>
-    autofit() |>
+    # autofit() |>
     fontsize(size = 8, part = "header") |>
     fontsize(size = 8, part = "footer") |>
     fontsize(size = 6, part = "body") |>
@@ -576,7 +576,7 @@ create_descriptives_table <- function(data) {
       align = "center",
       part = "all"
     ) |>
-    # width(j = NULL, width = 1, unit = "cm") |>
+    width(j = NULL, width = 1.5, unit = "in") |>
     delete_columns(j = "is_last_row") |>
     merge_v(j = c("Authors", "Article title")) |>
     footnote(
@@ -628,7 +628,9 @@ convert_descriptives_table_to_docx <- function(table) {
 convert_descriptives_table_to_image <- function(table) {
   
   # Save as image
-  save_as_image(table, path = "tables/descriptives_table.svg", res=300)
+  save_as_image(table, path = "tables/descriptives_table.png", res=600)
+  
+  save_as_image(table, path = "manuscript/descriptives_table.png", res=600)
 }
 
 # Pairwise data preparation for sensitivity analysis

### 2025-10-25 manuscript/pre_print.qmd
quarto for preprint updated with table

@@
 ```
-Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
+
+The PRISMA flow diagram can be seen Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
 Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
-Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved (they are however summarised in @tbl-review).
+Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved (they are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated).
 
-::: {.landscape}
+```{r}
+
+summary_descriptives <- main_arm_data_imputed_demographics_baseline |>
+  group_by(cond) |>
+  summarise(
+    min_age = round(min(m_age, na.rm = TRUE)),
+    max_age = round(max(m_age, na.rm = TRUE)),
+    min_bmi = round(min(m_bmi, na.rm = TRUE),1),
+    max_bmi = round(max(m_bmi, na.rm = TRUE),1)
+  )
+
+```
+
+All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry.
+The specific devices reported by these studies are included in @tbl-review.
+We originally considered in our pre-registration that, given sufficient data, we would compare sub groups of women with PCOS who did and did not have accompanying insulin resistance.
+However, based on the metabolic health variables reported in studies with this information (see @tbl-review) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance.
+Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`.
+Across those studies where BMI was reported or it was possible to estimate the mean BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]`.
+
 ![Descriptive characteristics of arms and participants for included studies](descriptives_table.png){#tbl-review}
-:::
 

### 2025-10-25 R/functions/main_data_functions.R

@@ -628,9 +628,9 @@ convert_descriptives_table_to_docx <- function(table) {
 convert_descriptives_table_to_image <- function(table) {
   
   # Save as image
-  save_as_image(table, path = "tables/descriptives_table.png", res=600)
+  save_as_image(table, path = "tables/descriptives_table.png", res=900)
   
-  save_as_image(table, path = "manuscript/descriptives_table.png", res=600)
+  save_as_image(table, path = "manuscript/descriptives_table.png", res=900)
 }
 
 # Pairwise data preparation for sensitivity analysis

### 2025-10-25 _targets.R

@@ -134,9 +134,26 @@ list(
     impute_bmi_estimates(main_arm_data)
   ),
   
+  # Make descriptives table with all studies including Greek ones
+  
+  tar_target(
+    main_plus_greek_arm_data,
+    prepare_data(main_arm_data_file)
+  ),
+  
+  tar_target(
+    main_plus_greek_arm_data_effects,
+    calculate_arm_effects(main_plus_greek_arm_data)
+  ),
+  
+  tar_target(
+    main_plus_greek_arm_data_imputed_demographics,
+    impute_bmi_estimates(main_plus_greek_arm_data)
+  ),
+  
   tar_target(
     descriptives_table,
-    create_descriptives_table(main_arm_data_imputed_demographics)
+    create_descriptives_table(main_plus_greek_arm_data_imputed_demographics)
   ),
   
   tar_target(
@@ -317,15 +334,6 @@ list(
   
   #### Sensitivity analysis including Greek lab studies ----
   
-  tar_target(
-    main_plus_greek_arm_data,
-    prepare_data(main_arm_data_file)
-  ),
-  
-  tar_target(
-    main_plus_greek_arm_data_effects,
-    calculate_arm_effects(main_plus_greek_arm_data)
-  ),
   
   # Fitting main arm analysis models including Greek studies
   

### 2025-10-25 manuscript/pre_print.qmd
manuscript sent to team for discussion to be written

@@
         \newcommand{\elandscape}{\end{landscape}}
+geometry:
+  - left=1in
+  - right=1in
+  - top=1in
+  - bottom=1in
 author:
@@
 targets::tar_load(c(
-  main_arm_data_imputed_demographics
+  c(main_arm_data_imputed_demographics,
+    main_plus_greek_arm_data_imputed_demographics)
 ))
@@
 
-The PRISMA flow diagram can be seen Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
+Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
 Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
-Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved (they are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated).
+Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved.
+These studies  are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al.
+[-@saltamavrosAlpha2Beta2007].
 

### 2025-10-25 R/functions/main_data_functions.R

@@ -625,12 +625,19 @@ convert_descriptives_table_to_docx <- function(table) {
   print(doc, target = "tables/descriptives_table.docx")
 }
 
-convert_descriptives_table_to_image <- function(table) {
+convert_descriptives_table_to_png <- function(table) {
   
   # Save as image
-  save_as_image(table, path = "tables/descriptives_table.png", res=900)
+  save_as_image(table, path = "tables/descriptives_table.png", res=300)
   
-  save_as_image(table, path = "manuscript/descriptives_table.png", res=900)
+  save_as_image(table, path = "manuscript/descriptives_table.png", res=300)
+}
+
+convert_descriptives_table_to_svg <- function(table) {
+  
+  # Save as image
+  save_as_image(table, path = "tables/descriptives_table.svg", res=300)
+  
 }
 
 # Pairwise data preparation for sensitivity analysis

### 2025-10-25 _targets.R

@@ -155,15 +155,15 @@ list(
     descriptives_table,
     create_descriptives_table(main_plus_greek_arm_data_imputed_demographics)
   ),
-  
+
   tar_target(
-    descriptives_table_docx,
-    convert_descriptives_table_to_docx(descriptives_table)
+    descriptives_table_png,
+    convert_descriptives_table_to_png(descriptives_table)
   ),
   
   tar_target(
-    descriptives_table_image,
-    convert_descriptives_table_to_image(descriptives_table)
+    descriptives_table_svg,
+    convert_descriptives_table_to_svg(descriptives_table)
   ),
   
   # Fitting main analysis models

### 2025-11-25 manuscript/pre_print.qmd
Edits up to discussion addressed

@@
 Polycystic ovary syndrome (PCOS) affects approximately 10% of women of reproductive age worldwide, making it the most common endocrine disorder affecting this population [@salariGlobalPrevalencePolycystic2024].
-Due to a number of factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015].
-Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into overweight or obese body mass index (BMI) categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
+Due to several factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015].
+Furthermore, epidemiological data has consistently demonstrated that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into overweight or obese body mass index (BMI) categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].
 
@@
 Indeed, the study from Georgopoulos et al.
-[-@georgopoulosBasalMetabolicRate2009] examining REE using indirect calorimetry in women with PCOS reported that they had a significantly lower REE, ~400 kcal per day less, compared with women without PCOS.
-Notably, they also reported insulin resistance further influenced REE among women with PCOS, with those who were insulin resistant exhibiting an additional reduction of about ~500 kcal per day compared to women with PCOS who were not insulin resistant [@georgopoulosBasalMetabolicRate2009].
-Other studies similarly report lower BMR using indirect methods (such as prediction from bioelectrical impedance analysis or accelerometer physical activity data) in women with PCOS, also highlighting factors like insulin resistance or BMI category influence REE [@romualdiRestingMetabolicRate2019; churchillBasalMetabolicRate2015].
+[-@georgopoulosBasalMetabolicRate2009] examined REE in women with and without PCOS using indirect calorimetry and reported that resting energy expenditure was approximately ~400 kcal/day lower in women with PCOS.
+Notably, they also reported that insulin resistance further reduced REE among women with PCOS, with insulin-resistant women exhibiting an additional reduction nearly 500 kcal per day compared to women with PCOS who were not insulin resistant [@georgopoulosBasalMetabolicRate2009].
+Other studies similarly report lower REE in women with PCOS using indirect methods (such as prediction from bioelectrical impedance analysis or accelerometer physical activity data), also suggesting that factors including insulin resistance and BMI category influence REE in women with PCOS [@romualdiRestingMetabolicRate2019; @churchillBasalMetabolicRate2015].
 However, despite the widespread acceptance that women living with PCOS exhibit reduced REE based on studies such as these, other research has reported little to no difference in REE between women with and without PCOS [@graffSaturatedFatIntake2017; @larssonDietaryIntakeResting2016; @segalRestingMetabolicRate1990].
 
-The consequences of widespread acceptance that REE is reduced in women living with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
+The consequences of widespread acceptance that REE is substantially lower in women living with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
 From a physiological perspective, if women with PCOS do exhibit a lower REE, this could imply a meaningful metabolic disadvantage that may influence dietary and nutritional guidance for weight management.
-However, if no such difference exists, the belief in a “slower metabolism” could instead serve as a deterrent to weight-loss efforts—efforts that are recognised as efficacious for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and are routinely recommended [@teedeRecommendations2023International2023].
+However, if no such difference exists, the belief in a “slower metabolism” could instead serve as a deterrent to energy restriction based weight-loss approaches that are recognised as efficacious for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and are routinely recommended [@teedeRecommendations2023International2023].
 It has been well documented that women with PCOS already experience higher rates of anxiety, depression, and lower quality of life (QOL) as a result of negative body image and weight-related concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023].
-Consequently, they are often observed to engage in more frequent weight-loss attempts than women without PCOS [@pesonenPolycysticOvarySyndrome2023].
+Consequently, they are often observed to engage in more frequent weight-loss attempts than women without PCOS, though it was unclear whether this is focused on energy restriction approaches [@pesonenPolycysticOvarySyndrome2023].
 Clarifying the relationship between REE and PCOS may therefore help guide more accurate clinical recommendations and empower both practitioners and women with PCOS.
 
-Therefore, to estimate the magnitude of difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
+Therefore, to estimate and describe the magnitude of difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.
 
@@
 
-The condition being studied was PCOS and we included observational cross-sectional design studies, in addition to studies of intervention where REE was reported for the population (and if present, the comparator i.e., women without PCOS) condition of interest.
+The condition being studied was PCOS and we included observational cross-sectional design studies, in addition to intervention studies where REE was reported for the population (and if present, the comparator i.e., women without PCOS) condition of interest.
 For clarity, studies of any design were included if they reported the REE using the methods indicated for a sample of adult women with PCOS and who were otherwise healthy.
 This included both studies with and without samples of healthy control women without PCOS.
-As detailed in the statistical analysis section below, a Bayesian model with informative priors based on normative data for REE in health women without PCOS was included to provide control information indirectly where this was missing.
-The use of such priors is an efficient tool to incorporate historical information on a particular population in a conservative manner [@weberApplyingMetaAnalyticPredictivePriors2021].
+As detailed in the statistical analysis section below, a Bayesian model with informative priors based on normative data for REE in healthy women without PCOS was included to provide control information indirectly where this was missing.
+The use of such priors is an efficient tool for incorporating historical information on a particular population in a conservative manner [@weberApplyingMetaAnalyticPredictivePriors2021].
 
@@
 ## Data extraction (selection and coding)
-Bibliometric data including authors, journal, and title of article were extracted.
+Bibliometric data including authors, journal, and article titles were extracted.
 Descriptive statistics for age, body mass, fat mass, fat free mass, height, BMI, race, physical activity levels, country of investigation, information regarding glucose/insulin regulation and insulin resistance status (where available), diagnostic criteria for PCOS, and measurement method and device were extracted for each arm within each study in addition to sample size.
@@
 Where REE values adjusted for body mass and/or fat free mass were reported we used the reported body mass and/or fat free mass mean values for that arm to convert them to unadjusted REE values (i.e., multiplied them by body mass and/or fat free mass mean values).
-Where means and/or standard deviations were missing the latter were either calculated from standard errors and sample size, or all both were estimated from lower and upper range, interquartile range, median, and sample size depending on the information available using the methods of Wan et al.
+Where means and/or standard deviations were missing the latter were either calculated from standard errors and sample size, or all both were estimated from lower and upper range, interquartile range, median, and sample size depending on the available information using the methods of Wan et al.
 [-@wanEstimatingSampleMean2014].
-Further, where missing, height/body mass/BMI where estimated based on the means reported for these.
+Further, where missing, height/body mass/BMI where estimated based on the reported means.
 The units of measurement for which REE was extracted and all REE values were converted to kcal/day.
-In one case [@pohlmeierEffectLowstarchLowdairy2014] REE was reported relative to body mass and the unadjusted values were no longer available (confirmed by the authors) as such in this case we used the mean body mass to convert back to estimated REE unadjusted.
+In one case [@pohlmeierEffectLowstarchLowdairy2014] REE was reported relative to body mass and the unadjusted values were no longer available (confirmed by the authors).
+As such, in this case we used the mean body mass to convert back to estimated REE unadjusted.
 
@@
 
-Given these issues we made a decision to extract the results from these studies as reported, but to conduct analyses both with and without their inclusion.
-Though not pre-registered, due to a lack of confidence in the results as reported, we decided to include the analysis omitting these studies as our main models in the results reported below.
-The results of the analysis including them are however reported in the sensitivity analysis section.
+Given these issues we decided to extract the results from these studies as reported and to conduct analyses both with and without their inclusion.
+Though not pre-registered, due to a lack of confidence in the reported results, we decided to include the analysis omitting these studies as our main models in the results reported below.
+The results of the analysis including them are reported in the sensitivity analysis section.
 
@@
 We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf>.
-The statistical analysis plan was linked in our pre-registration (PROSPERO: [CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) and available at the accompanying github repository.
+The statistical analysis plan was linked in our pre-registration (PROSPERO: [CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) and available at the accompanying GitHub repository.
 Any deviations from the pre-registration are noted below.
@@
 We made use of historical information regarding REE in healthy control women without PCOS by setting informative priors based on meta-analysis of large scale studies reporting normative data for REE in this population.
-This was included to provide control information indirectly where this was missing from particular studies.
-The use of historical priors like this is an efficient tool to incorporate historical information on a particular population in a conservative manner in meta-analyses [@weberApplyingMetaAnalyticPredictivePriors2021].
+This was included to provide indirect control information  where it was missing from particular studies.
+The use of historical priors like this is an efficient tool to incorporate historical information about a particular population in a conservative manner in meta-analyses [@weberApplyingMetaAnalyticPredictivePriors2021].
 From this model we focus on reporting the global grand mean estimate for the fixed between condition relative contrast for non-PCOS control vs PCOS arms as our primary estimand of interest (i.e., $\beta_1$ in both mean and standard deviation models).
 We examined both raw mean REE (i.e., the absolute mean REE in kcals per day for each arm) in addition to the between person standard deviation in REE (i.e., the absolute standard deviation in REE in kcals per day for each arm).
-Both models were multilevel in that they included nested random intercepts for both study, and arm within study, and in addition and in deviation from our pre-registration we also included lab as a random intercept as in some cases we had multiple studies from the same lab or research group included.
-Lastly, the inclusion of a random intercept for each effect size was also accidentally omitted from our pre-registration and so this is also included in the model.
+Both models were multilevel in that they included nested random intercepts for both study and arm within study.
+In addition, and in deviation from our pre-registration, we also included lab as a random intercept as in some cases we had multiple studies from the same lab or research group.
+Lastly, the inclusion of a random intercept for each effect size was accidentally omitted from our pre-registration, and so this is also included in the model.
 
@@
 
+Note, the numbers in this section exclude the four studies previously noted with possible reporting issues that were unresolved.
+These studies are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al.
+[-@saltamavrosAlpha2Beta2007].
+
 Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
 Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
-Note, these numbers exclude the four studies previously noted with possible reporting issues that were unresolved.
-These studies  are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al.
-[-@saltamavrosAlpha2Beta2007].
 
@@
 
+countries <- main_plus_greek_arm_data_imputed_demographics |>
+  group_by(study) |>
+  slice_head(n=1) |>
+  ungroup() |>
+  group_by(country) |>
+  count()
+
 ```
 
+Studies were carried out in multiple countries: Brazil (k = `r filter(countries, country == "Brazil")$n`) ,and USA (k = `r filter(countries, country == "USA")$n` studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK (all k = 1 study).
+The four studies with noted reporting errors were carried out in Greece.
+A total of 11 studies used the Rotterdam criteria (or a modified version there of) for diagnosing PCOS [@robinsonPostprandialThermogenesisReduced1992; @brunerEffectsExerciseNutritional2006; @moranShorttermMealReplacements2006; @cosarRestingMetabolicRate2008; @graffDietaryGlycemicIndex2013; @pohlmeierEffectLowstarchLowdairy2014; @dohRelationshipAdiposityInsulin2016; @larssonDietaryIntakeResting2016; @graffSaturatedFatIntake2017; @rodriguesLowValidityPredictive2018; @tosiInsulinMediatedSubstrateUse2021].
+One study used the 1990 National Institutes of Health criteria [@broskeyAssessingEnergyRequirements2017], and two studies [@segalRestingMetabolicRate1990; @robinsonPostprandialThermogenesisReduced1992] diagnosed PCOS via the presence of oligomenorrhea or amenorrheaammenorrhea alongside additional criteria including plasma androgen levels, hirsutism or polycystic ovaries on ultrasound scanning.
 All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry.
@@
 However, based on the metabolic health variables reported in studies with this information (see @tbl-review) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance.
-Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`.
+Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging from `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`.
 Across those studies where BMI was reported or it was possible to estimate the mean BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]`.
@@
 The corresponding conditional estimates for the control condition and PCOS condition respectively were `r round(main_arm_mean_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[1])` kcal/day] and `r round(main_arm_mean_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[2])` kcal/day].
-These results including the full visualisation of the posterior distribution, in addition to the conditional estimates by study, can be seen in @fig-mean-REE.
+These results, including the full visualisation of the posterior distribution and the conditional estimates by study, can be seen in @fig-mean-REE.
 
@@
 The corresponding conditional estimates for the standard deviations of the control condition and PCOS condition respectively were `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day] and `r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day].
-These results including the full visualisation of the posterior distribution, in addition to the conditional estimates by study, can be seen in @fig-variance-REE.
+These results, including the full visualisation of the posterior distribution and the conditional estimates by study, can be seen in @fig-variance-REE.
 
@@
 For mean REE the pairwise model resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(pairwise_mean_effects_model)[1])` kcal/day with a 95% quantile interval ranging from `r round(brms::fixef(pairwise_mean_effects_model)[3])` kcal/day to `r round(brms::fixef(pairwise_mean_effects_model)[4])` kcal/day.
-This was similar for the standard deviation of REE with the pairwise model resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(pairwise_variance_effects_model)[1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(pairwise_variance_effects_model)[3]),2)` to `r round(exp(brms::fixef(pairwise_variance_effects_model)[4]),2)`.
+This was similar for the standard deviation of REE with the pairwise model resulting in a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(pairwise_variance_effects_model)[1]),2)` and a 95% quantile interval ranging from `r round(exp(brms::fixef(pairwise_variance_effects_model)[3]),2)` to `r round(exp(brms::fixef(pairwise_variance_effects_model)[4]),2)`.
 
@@
 For mean REE the model including the studies noted with possible reporting issues that were not resolved [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009] still resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,1])` kcal/day with a 95% quantile interval ranging from `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,3])` kcal/day to `r round(brms::fixef(main_plus_greek_arm_mean_effects_model)[2,4])` kcal/day.
-This was similar for the standard deviation of REE with the model including these studies resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,4]),2)`.
+This was similar for the standard deviation of REE with the model including these studies resulting in a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,1]),2)` and a 95% quantile interval ranging from `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(main_plus_greek_arm_variance_effects_model)[2,4]),2)`.
 
@@
 For mean REE the model which included only baseline REE measurements from studies involving interventions [@moranShorttermMealReplacements2006; @brunerEffectsExerciseNutritional2006; @pohlmeierEffectLowstarchLowdairy2014], in addition to other cross-sectional studies, also resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(brms::fixef(baseline_arm_mean_effects_model)[2,1])` kcal/day with a 95% quantile interval ranging from `r round(brms::fixef(baseline_arm_mean_effects_model)[2,3])` kcal/day to `r round(brms::fixef(baseline_arm_mean_effects_model)[2,4])` kcal/day.
-This was the case also for the standard deviation of REE with this model resulting in a contrast ratio between control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,1]),2)` with a 95% quantile interval ranging from  `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,4]),2)`.
+This was also the case for the standard deviation of REE with this model resulting in a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,1]),2)` and a 95% quantile interval ranging from `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,3]),2)` to `r round(exp(brms::fixef(baseline_arm_variance_effects_model)[2,4]),2)`.
 

### 2025-11-26 manuscript/pre_print.qmd
Edits to intro to better explain impact on recommendations

@@
 The consequences of widespread acceptance that REE is substantially lower in women living with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population.
-From a physiological perspective, if women with PCOS do exhibit a lower REE, this could imply a meaningful metabolic disadvantage that may influence dietary and nutritional guidance for weight management.
-However, if no such difference exists, the belief in a “slower metabolism” could instead serve as a deterrent to energy restriction based weight-loss approaches that are recognised as efficacious for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and are routinely recommended [@teedeRecommendations2023International2023].
+Women with PCOS typically engage in more frequent weight-loss attempts than women without PCOS [@pesonenPolycysticOvarySyndrome2023].
+From a physiological perspective, if women with PCOS do exhibit a lower REE, this could imply a meaningful metabolic disadvantage that may influence dietary and nutritional guidance for weight management; for example, recommending a slightly more severe energy restriction to overcome the belief that they have a lower REE [@ozgensaydamWeightManagementStrategies2021].
+Recommendations such as this could influence the well documented prevalence of eating disorders in women with PCOS [@lalonde-besterPrevalenceEtiologyEating2024; @jeanesBingeEatingBehaviours2017].
+Contrastingly, belief in a “slower metabolism” could instead serve as a deterrent to energy restriction based weight-loss approaches for some women in line with typical general population recommendations that are also recognised as efficacious for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and are routinely recommended [@teedeRecommendations2023International2023].
 It has been well documented that women with PCOS already experience higher rates of anxiety, depression, and lower quality of life (QOL) as a result of negative body image and weight-related concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023].
-Consequently, they are often observed to engage in more frequent weight-loss attempts than women without PCOS, though it was unclear whether this is focused on energy restriction approaches [@pesonenPolycysticOvarySyndrome2023].
 Clarifying the relationship between REE and PCOS may therefore help guide more accurate clinical recommendations and empower both practitioners and women with PCOS.
@@
 
-TO WRITE
+This study sought to estimate and describe the magnitude of difference in REE between women with and without PCOS.
+Most studies identified in the systematic review, and included in the meta-analysis, used indirect calorimetry as the primary measure of REE and assessed women with PCOS who were insulin resistant and categorised as being in overweight or obese BMI categories compared to healthy controls.
+Our results indicate there is only a small magnitude of difference in REE (`r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day [95% quantile interval: `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)`kcal/day]) between women with PCOS and those without and also little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95% quantile interval: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]).
+These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain.
+This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
+Recognising that there may be minimal differences in REE between women with PCOS and those without can inform both clinical and public practices, potentially leading to a shift in focus away from a requirement for more severe caloric restriction as a primary method of treatment towards more comprehensive, individualised, and psychologically safe approaches to care [@teedeRecommendations2023International2023].
+
+The current study suggests that REE may not be the primary barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (1442 kcal/day vs.
+1472 kcal/ day, respectively).
+Similar to our findings, a case-control study comparing women with PCOS to healthy controls demonstrated REE was similar between groups (PMID: 38867472); however, REE adjusted for fat-free mass was lower in women with PCOS suggesting the importance of fat-free tissue in energy regulation.
+Indeed, a recent systematic review of mechanisms for metabolic dysfunction found excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780).
+Taken together, these findings highlight that factors beyond REE, such as hormonal and tissue-specific metabolic effects, likely play a more significant role in weight regulation challenges in women with PCOS.
+
+Women with PCOS are routinely encouraged to participate in caloric restriction for weight management, advice that may inadvertently foster disordered eating.
+The prevalence of any eating disorder in women with PCOS has been reported at up to 62%, with affected women three to six times more likely than controls to experience an eating disorder (PMID: 38408541).
+Specific conditions such as binge-eating disorder (about 17%), bulimia nervosa (about 6 per cent) and night-eating syndrome (about 13%) are all markedly more common than in the general female population (PMID: 28104244, PMID: 27244869).
+Approximately 60% of women with PCOS report binge or comfort eating, with 39% displaying clinically significant behaviours (PMID: 27825940).
+The pathways linking PCOS and disordered eating are multifactorial.
+Biological mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered ghrelin and leptin signalling can heighten hunger, carbohydrate cravings, and appetite variability (PMID: 21421016).
+Frequent hypoglycaemia and associated mood changes have also been observed, which can trigger compensatory eating or binge episodes (PMID: 21421016).
+These physiological processes interact with psychological and social stressors, including infertility concerns, conflicting nutrition advice, chronic dieting, and exposure to idealised body images on social media, which together compound vulnerability to disordered eating (PMID 33153014).
+Moreover, eating disorders themselves can disrupt endocrine function, potentially worsening PCOS symptoms and creating a self-reinforcing cycle (PMID 39115340, PMID 27244869).
+Understanding these intertwined biological, psychological, and social influences underscores the importance of avoiding overly restrictive dietary advice that can exacerbate feelings of failure, hunger dysregulation, and shame (PMID31615157).
+When metabolic rate is not impaired, recommending large caloric deficits is not only unnecessary but may be actively harmful.
+A further barrier to effective management is weight stigma, which remains pervasive in both healthcare and everyday life (PMID: 39462252).
+Women with PCOS often report being labelled as “non-compliant” or “lacking willpower” when they do not lose weight rapidly, despite evidence of complex metabolic and psychosocial contributors (PMID: 37580314).
+Such attitudes are not benign.
+Weight stigma has been linked to poorer self-esteem, depression, anxiety, avoidance of medical care, and increased disordered eating.
+Physiologically, stigma-related stress elevates chronic cortisol, blood pressure, and insulin resistance, thereby worsening cardiometabolic outcomes (PMID: 37580314).
+Repeated cycles of weight loss and regain, driven by shame and unrealistic expectations, are associated with higher long-term weight, adverse lipid profiles, and elevated cardiovascular risk (PMID 37436934, PMID 23046078, PMID 10792339).
+Encouragingly, the 2023 international PCOS guidelines now recognise weight stigma as a determinant of health and call for its reduction across clinical and public health settings.
+Addressing stigma is therefore both an ethical imperative and a practical strategy to improve health outcomes.
+Given these insights, the finding of a similar REE  in women with PCOS compared to those without PCOS invites a paradigm shift towards weight-neutral, individualised, and empowering care.
+Without evidence of metabolic disadvantage, interventions need not centre on extreme energy deficits.
+Instead, practitioners can focus on improving insulin sensitivity, inflammatory markers, and hormonal regulation through balanced nutrition, physical activity for enjoyment, adequate sleep, and stress management.
+Screening for disordered eating and providing evidence-based psychological support should be standard components of PCOS care.
+Approaches grounded in intuitive or mindful eating encourage self-trust and attunement to hunger and fullness cues, while dismantling moral judgements about food.
+This can stabilise glycaemic control, reduce binge episodes, and promote sustainable behaviour change (PMID: 32006391).
+Health coaching and dietetic practice should also include ongoing reflection on personal biases, the use of non-stigmatising language, and the creation of inclusive environments for people in all body sizes.
+Importantly, this reframing is empowering for women with PCOS.
+Understanding that they are not “metabolically disadvantaged” enables them to move forward in their health journey without the burden of restrictive calorie targets or the belief that they are inherently at fault.
+By shifting attention towards metabolic-health parameters such as insulin resistance, stress management, and healthy lifestyle, women can pursue improvements in energy, mood, fertility, and quality of life on their own terms.
+When the focus becomes nourishment rather than restriction, self-care rather than self-control, and progress rather than perfection, sustainable wellbeing becomes achievable.
+
+Strengths and Limitations
+The current study has multiple strengths stemming from its preregistered, comprehensive methodology and Bayesian approach.
+This statistical framework allowed us to incorporate studies with and without control groups to better estimate mean REE in women with and without PCOS and to perform multiple sensitivity analyses that confirmed the stability of our findings.
+However, a limitation here is that variability in methods across studies, such as differences in PCOS diagnostic criteria and REE testing protocols, may have influenced results.
+Furthermore, some studies controlled for body weight or body composition when reporting REE values, while others did not.
+We accounted for this by estimating or converting reported data to obtain unadjusted REE values across all groups, thereby reducing this variability.
+Another limitation is that, due to fewer total control groups than PCOS groups, informed priors were required in several statistical models.
+Additionally, some studies reported data inconsistencies that could not be clarified (e.g., Georgopoulos et al.) and were accounted for in our analysis.
+Finally, most included studies were cross-sectional or baseline assessments within intervention trials, which limits causal inference.
+
+Conclusion
+In conclusion, the findings from this meta-analysis indicate that resting energy expenditure does not meaningfully differ between women with and without PCOS.
+Group-level differences in resting energy expenditure were small, insignificant, or not physiologically relevant.
+Additionally, variability in resting energy expenditure was also similar, suggesting that a lower baseline metabolic rate is not a causal factor for the weight-related challenges often associated with PCOS.
+These findings challenge the popular narrative that women with PCOS have a lower resting metabolism and may help better inform dietary interventions and nutritional support for these individuals.
+Future research should include more standardized REE measurement and reporting protocols, greater data transparency, consistent control and reporting of body weight or body composition, the presentation of both absolute and relative REE, and more precise characterization of PCOS phenotypes.
+Overall, these findings support the conclusion that PCOS does not negatively influence resting energy expenditure and may help practitioners and researchers focus on targeted lifestyle interventions rather than negatively framed metabolic assumptions.
 

### 2025-11-26 R/functions/model_functions.R
Add models adjusted for BMI

@@ -589,3 +589,81 @@ fit_pairwise_variance_model <- function(data) {
   
   return(pairwise_model)
 }
+
+# Additional models
+fit_arm_mean_effects_model_moderator <- function(data, prior, moderator) {
+  
+  formula <- as.formula(paste0("yi_mean | se(sqrt(vi_mean)) ~ 0 + Intercept + cond + me(m_",moderator,",se_",moderator,") +", 
+                               "(1 + cond | lab) + (1 + cond | study) + (1 | arm) + (1|effect)"))
+  
+  arm_model <- brm(formula,
+                   data = data,
+                   prior = prior,
+                   chains = 4,
+                   cores = 4,
+                   seed = 1988,
+                   warmup = 2000,
+                   iter = 8000)
+}
+
+fit_arm_variance_effects_model_moderator <- function(data, prior, moderator) {
+  
+  formula <- as.formula(paste0("yi_sd | se(sqrt(vi_sd)) ~ 0 + Intercept + cond + me(m_",moderator,",se_",moderator,") +", 
+  "me(log_yi_mean, se_log_yi_mean) + (1 + cond | lab) + (1 + cond | study) + (1 | arm) + (1|effect)"))
+  
+  arm_model <- brm(formula,
+                   data = data,
+                   prior = prior,
+                   chains = 4,
+                   cores = 4,
+                   seed = 1988,
+                   warmup = 2000,
+                   iter = 8000)
+}
+
+get_mean_contrast_condition_bmi <- function(model, data) {
+  
+  # use median BMI in controls
+  m_bmi <- data |>
+    filter(cond == "Control") |>
+    group_by(arm) |>
+    slice_head(n=1) |>
+    ungroup() |>
+    summarise(m_bmi = median(m_bmi, na.rm = TRUE))
+  
+  # pooled preds per condition
+  preds <- avg_comparisons(
+    model,
+    newdata = datagrid(
+      m_bmi = m_bmi$m_bmi
+    ),
+    re_formula = NA,
+    variables = "cond"
+  ) 
+  
+  return(preds)
+}
+
+get_variance_contrast_condition_bmi <- function(model, data) {
+  
+  # use median BMI in controls
+  m_bmi <- data |>
+    filter(cond == "Control") |>
+    group_by(arm) |>
+    slice_head(n=1) |>
+    ungroup() |>
[…]

### 2025-11-26 _targets.R

@@ -351,7 +351,48 @@ list(
       main_plus_greek_arm_data_effects,
       prior_arm_variance_effects
     )
+  ),
+  
+  #### Additional models including BMI
+  
+  # add imputed demographics for moderators
+  tar_target(
+    main_arm_data_effects_imputed_demographics,
+    impute_bmi_estimates(main_arm_data_effects) |>
+      mutate(se_bmi = sd_bmi/sqrt(n_bmi))
+  ),
+  
+  tar_target(
+    main_arm_mean_effects_model_bmi,
+    fit_arm_mean_effects_model_moderator(
+      main_arm_data_effects_imputed_demographics,
+      prior_arm_mean_effects,
+      "bmi"
+    )
+  ),
+  
+  tar_target(
+    main_arm_variance_effects_model_bmi,
+    fit_arm_variance_effects_model_moderator(
+      main_arm_data_effects_imputed_demographics,
+      prior_arm_variance_effects,
+      "bmi"
+    )
+  ),
+  
+  tar_target(
+    main_arm_mean_effects_contrast_condition_bmi,
+    get_mean_contrast_condition_bmi(main_arm_mean_effects_model_bmi,
+                                    main_arm_data_effects_imputed_demographics)
+  ),
+  
+  tar_target(
+    main_arm_variance_effects_contrast_condition_bmi,
+    get_variance_contrast_condition_bmi(main_arm_variance_effects_model_bmi,
+                                    main_arm_data_effects_imputed_demographics)
   )
+  
+  
 )
 
 

### 2025-11-26 R/functions/model_functions.R
bmi and ffm functions updated

@@ -591,6 +591,32 @@ fit_pairwise_variance_model <- function(data) {
 }
 
 # Additional models
+get_predictor_medians <- function(data) {
+  # use median BMI in controls
+  m_bmi <- data |>
+    filter(cond == "Control") |>
+    group_by(arm) |>
+    slice_head(n=1) |>
+    ungroup() |>
+    summarise(m_bmi = median(m_bmi, na.rm = TRUE))
+  
+  # use median BMI in controls
+  m_fat_free_mass <- data |>
+    filter(cond == "Control") |>
+    group_by(arm) |>
+    slice_head(n=1) |>
+    ungroup() |>
+    summarise(m_fat_free_mass = median(m_fat_free_mass, na.rm = TRUE))
+  
+  predictor_medians <- tibble(
+    m_bmi = m_bmi$m_bmi,
+    m_fat_free_mass = m_fat_free_mass$m_fat_free_mass
+  )
+  
+  return(predictor_medians)
+}
+  
+
 fit_arm_mean_effects_model_moderator <- function(data, prior, moderator) {
   
   formula <- as.formula(paste0("yi_mean | se(sqrt(vi_mean)) ~ 0 + Intercept + cond + me(m_",moderator,",se_",moderator,") +", 
@@ -621,21 +647,14 @@ fit_arm_variance_effects_model_moderator <- function(data, prior, moderator) {
                    iter = 8000)
 }
 
-get_mean_contrast_condition_bmi <- function(model, data) {
-  
-  # use median BMI in controls
-  m_bmi <- data |>
-    filter(cond == "Control") |>
-    group_by(arm) |>
-    slice_head(n=1) |>
-    ungroup() |>
-    summarise(m_bmi = median(m_bmi, na.rm = TRUE))
+get_mean_contrast_condition_moderator <- function(model, predictor_medians) {
   
   # pooled preds per condition
   preds <- avg_comparisons(
     model,
     newdata = datagrid(
-      m_bmi = m_bmi$m_bmi
+      m_bmi = predictor_medians$m_bmi,
+      m_fat_free_mass = predictor_medians$m_fat_free_mass
     ),
     re_formula = NA,
     variables = "cond"
@@ -644,22 +663,15 @@ get_mean_contrast_condition_bmi <- function(model, data) {
   return(preds)
 }
 
-get_variance_contrast_condition_bmi <- function(model, data) {
-  
-  # use median BMI in controls
-  m_bmi <- data |>
[…]

### 2025-11-26 _targets.R

@@ -353,15 +353,23 @@ list(
     )
   ),
   
-  #### Additional models including BMI
+  #### Additional models including BMI and fat free mass
   
-  # add imputed demographics for moderators
+  # Add imputed demographics for moderators
   tar_target(
     main_arm_data_effects_imputed_demographics,
     impute_bmi_estimates(main_arm_data_effects) |>
-      mutate(se_bmi = sd_bmi/sqrt(n_bmi))
+      mutate(se_bmi = sd_bmi/sqrt(n_bmi),
+             se_fat_free_mass = sd_fat_free_mass/sqrt(n_fat_free_mass))
   ),
   
+  # Get control medians for predictors
+  tar_target(
+    predictor_medians,
+    get_predictor_medians(main_arm_data_effects_imputed_demographics)
+  ),
+  
+  # BMI
   tar_target(
     main_arm_mean_effects_model_bmi,
     fit_arm_mean_effects_model_moderator(
@@ -382,16 +390,49 @@ list(
   
   tar_target(
     main_arm_mean_effects_contrast_condition_bmi,
-    get_mean_contrast_condition_bmi(main_arm_mean_effects_model_bmi,
-                                    main_arm_data_effects_imputed_demographics)
+    get_mean_contrast_condition_moderator(main_arm_mean_effects_model_bmi,
+                                          predictor_medians)
   ),
   
   tar_target(
     main_arm_variance_effects_contrast_condition_bmi,
-    get_variance_contrast_condition_bmi(main_arm_variance_effects_model_bmi,
-                                    main_arm_data_effects_imputed_demographics)
-  )
+    get_variance_contrast_condition_moderator(main_arm_variance_effects_model_bmi,
+                                              main_arm_data_effects_imputed_demographics,
+                                        predictor_medians)
+  ),
   
+  # Fat free mass
+  
+  tar_target(
+    main_arm_mean_effects_model_fat_free_mass,
+    fit_arm_mean_effects_model_moderator(
+      main_arm_data_effects_imputed_demographics,
+      prior_arm_mean_effects,
+      "fat_free_mass"
+    )
+  ),
+  
+  tar_target(
+    main_arm_variance_effects_model_fat_free_mass,
+    fit_arm_variance_effects_model_moderator(
+      main_arm_data_effects_imputed_demographics,
+      prior_arm_variance_effects,
+      "fat_free_mass"
+    )
+  ),
+  
+  tar_target(
+    main_arm_mean_effects_contrast_condition_fat_free_mass,
[…]

### 2025-11-27 manuscript/pre_print.qmd
Add front matter for double line spacing and line numbers

@@
       text: |
+        \usepackage{lineno}\linenumbers
+        \usepackage{setspace}\doublespacing
         \usepackage[font=scriptsize]{caption}
@@
   Background: Polycystic ovary syndrome (PCOS) is the most common endocrine condition affecting women of reproductive age.
-Women who are living with PCOS are also more likely to experience overwweight and obesity and this is assumed to be a result of a lower resting energy expenditure (REE).
-The assumption that REE is lower in women living with PCOS has also influenced lifestyle-based interventions in this group.
+Women who are living with PCOS are also more likely to fall into overweight or obese body mass index (BMI) categories and this is assumed to be a result of a lower resting energy expenditure (REE).
+The assumption that REE is lower in women living with PCOS may also influence lifestyle-based interventions in this group.
 Evidence supporting this reduced REE compared to women without PCOS, however, is not consistent.
-Objectives: To systematically search and meta-analyse the existing literature to determine if resting energy expenditure differs between women with and without PCOS.
+Objectives: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS.
 Design: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.

### 2025-11-27 manuscript/pre_print.qmd
Change title to <120 characters for JCEM

@@
 title: |
-  Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
+  Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis
 date: last-modified

### 2025-11-27 manuscript/pre_print.qmd
Abstract and discussion edits (for some reason didn't previously save)

@@
 abstract: |
-  Background: Polycystic ovary syndrome (PCOS) is the most common endocrine condition affecting women of reproductive age.
-Women who are living with PCOS are also more likely to fall into overweight or obese body mass index (BMI) categories and this is assumed to be a result of a lower resting energy expenditure (REE).
-The assumption that REE is lower in women living with PCOS may also influence lifestyle-based interventions in this group.
-Evidence supporting this reduced REE compared to women without PCOS, however, is not consistent.
-Objectives: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS.
-Design: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.
-Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included.
-Results: Seventeen independent studies were identified.
-Study populations ranged from X to X participants, with mean ages ranging from X to X years.
-Conclusions: TO ADD
+  Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age women, who often have higher BMI classification.
+This is assumed to stem from lower resting energy expenditure (REE), influencing lifestyle intervention guidelines.
+However, evidence for reduced REE in women with PCOS compared with those without is inconsistent.
+Objective: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS.
+Data Sources: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.
+Study Selection: Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included.
+Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate.
+Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors was used to compare both mean REE, and between person variation in REE, between women with and without PCOS.
+Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33].
+Conclusions: These findings indicate that REE does not meaningfully differ between women with and without PCOS.
+Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.
 keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
@@
 
-# Methods
+# Materials and Methods
 This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@pagePRISMA2020Statement2021a].
@@
 ## Data extraction (selection and coding)
+Data was extracted by one investigator and checked in triplicate.
 Bibliometric data including authors, journal, and article titles were extracted.
@@
 Most studies identified in the systematic review, and included in the meta-analysis, used indirect calorimetry as the primary measure of REE and assessed women with PCOS who were insulin resistant and categorised as being in overweight or obese BMI categories compared to healthy controls.
-Our results indicate there is only a small magnitude of difference in REE (`r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day [95% quantile interval: `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)`kcal/day]) between women with PCOS and those without and also little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95% quantile interval: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]).
+Our results indicate there is only a small magnitude of difference in REE (`r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)`kcal/day]) between women with PCOS and those without.
+Further, there is little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95%QI: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]) suggesting that, despite individual differences in REE, PCOS is not systematically associated with lesser or greater individual variability.
 These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain.
@@
 
-The current study suggests that REE may not be the primary barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (1442 kcal/day vs.
-1472 kcal/ day, respectively).
-Similar to our findings, a case-control study comparing women with PCOS to healthy controls demonstrated REE was similar between groups (PMID: 38867472); however, REE adjusted for fat-free mass was lower in women with PCOS suggesting the importance of fat-free tissue in energy regulation.
-Indeed, a recent systematic review of mechanisms for metabolic dysfunction found excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780).
-Taken together, these findings highlight that factors beyond REE, such as hormonal and tissue-specific metabolic effects, likely play a more significant role in weight regulation challenges in women with PCOS.
+The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively).
+If anything, REE may be slightly higher in women with PCOS compared to healthy women without PCOS.
+BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]` and this may explain the slightly greater REE in the former group.
+However, for those studies where we could extract or estimate BMI, we fit an additional exploratory model extending the main arm based model and extracted an estimate of the difference adjusted for BMI as a fixed effect.
+Adjusted to the median value of average BMI of the control arms of women without PCOS (BMI = `r main_arm_mean_effects_contrast_condition_bmi$m_bmi` kg/$\textrm{m}^2$) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)`kcal/day]) between women with PCOS and those without.
+However, one of the studies included in our analysis [@tosiRestingEnergyExpenditure2024] reported that, whilst there was little difference in unadjusted REE, when REE was adjusted for fat-free mass it was lower in women with PCOS suggesting the potential importance of fat-free tissue in energy regulation.
+Yet, similarly to the additional exploratory model including BMI, when we adjusted for fat-free mass including those studies reporting this characteristic (fat-free mass = `r main_arm_mean_effects_contrast_condition_fat_free_mass$m_fat_free_mass` kg) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)`kcal/day]).
+As such, even adjusted for both BMI and fat-free mass, there seems to be little difference in REE between women with, and without, PCOS.
+Yet, a recent systematic review of mechanisms for metabolic dysfunction has reported excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780).
+Taken together, these findings highlight that factors beyond REE and typical correlates of this including BMI or fat-free mass, such as hormonal and tissue-specific metabolic effects, may play a more significant role in weight regulation challenges in women with PCOS.
 
-Women with PCOS are routinely encouraged to participate in caloric restriction for weight management, advice that may inadvertently foster disordered eating.
-The prevalence of any eating disorder in women with PCOS has been reported at up to 62%, with affected women three to six times more likely than controls to experience an eating disorder (PMID: 38408541).
-Specific conditions such as binge-eating disorder (about 17%), bulimia nervosa (about 6 per cent) and night-eating syndrome (about 13%) are all markedly more common than in the general female population (PMID: 28104244, PMID: 27244869).
-Approximately 60% of women with PCOS report binge or comfort eating, with 39% displaying clinically significant behaviours (PMID: 27825940).
+As noted, women with PCOS are more likely to engage in weight-loss attempts [@pesonenPolycysticOvarySyndrome2023] and there could be concerns that this could inadvertently further foster the already well documented disordered eating in this population [@lalonde-besterPrevalenceEtiologyEating2024; @jeanesBingeEatingBehaviours2017; @leeIncreasedRiskDisordered2017; @bernadettPrevalenceEatingDisorders2016].
 The pathways linking PCOS and disordered eating are multifactorial.
-Biological mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered ghrelin and leptin signalling can heighten hunger, carbohydrate cravings, and appetite variability (PMID: 21421016).
-Frequent hypoglycaemia and associated mood changes have also been observed, which can trigger compensatory eating or binge episodes (PMID: 21421016).
-These physiological processes interact with psychological and social stressors, including infertility concerns, conflicting nutrition advice, chronic dieting, and exposure to idealised body images on social media, which together compound vulnerability to disordered eating (PMID 33153014).
-Moreover, eating disorders themselves can disrupt endocrine function, potentially worsening PCOS symptoms and creating a self-reinforcing cycle (PMID 39115340, PMID 27244869).
-Understanding these intertwined biological, psychological, and social influences underscores the importance of avoiding overly restrictive dietary advice that can exacerbate feelings of failure, hunger dysregulation, and shame (PMID31615157).
-When metabolic rate is not impaired, recommending large caloric deficits is not only unnecessary but may be actively harmful.
-A further barrier to effective management is weight stigma, which remains pervasive in both healthcare and everyday life (PMID: 39462252).
-Women with PCOS often report being labelled as “non-compliant” or “lacking willpower” when they do not lose weight rapidly, despite evidence of complex metabolic and psychosocial contributors (PMID: 37580314).
-Such attitudes are not benign.
-Weight stigma has been linked to poorer self-esteem, depression, anxiety, avoidance of medical care, and increased disordered eating.
-Physiologically, stigma-related stress elevates chronic cortisol, blood pressure, and insulin resistance, thereby worsening cardiometabolic outcomes (PMID: 37580314).
-Repeated cycles of weight loss and regain, driven by shame and unrealistic expectations, are associated with higher long-term weight, adverse lipid profiles, and elevated cardiovascular risk (PMID 37436934, PMID 23046078, PMID 10792339).
-Encouragingly, the 2023 international PCOS guidelines now recognise weight stigma as a determinant of health and call for its reduction across clinical and public health settings.
-Addressing stigma is therefore both an ethical imperative and a practical strategy to improve health outcomes.
-Given these insights, the finding of a similar REE  in women with PCOS compared to those without PCOS invites a paradigm shift towards weight-neutral, individualised, and empowering care.
-Without evidence of metabolic disadvantage, interventions need not centre on extreme energy deficits.
-Instead, practitioners can focus on improving insulin sensitivity, inflammatory markers, and hormonal regulation through balanced nutrition, physical activity for enjoyment, adequate sleep, and stress management.
-Screening for disordered eating and providing evidence-based psychological support should be standard components of PCOS care.
-Approaches grounded in intuitive or mindful eating encourage self-trust and attunement to hunger and fullness cues, while dismantling moral judgements about food.
-This can stabilise glycaemic control, reduce binge episodes, and promote sustainable behaviour change (PMID: 32006391).
-Health coaching and dietetic practice should also include ongoing reflection on personal biases, the use of non-stigmatising language, and the creation of inclusive environments for people in all body sizes.
-Importantly, this reframing is empowering for women with PCOS.
-Understanding that they are not “metabolically disadvantaged” enables them to move forward in their health journey without the burden of restrictive calorie targets or the belief that they are inherently at fault.
-By shifting attention towards metabolic-health parameters such as insulin resistance, stress management, and healthy lifestyle, women can pursue improvements in energy, mood, fertility, and quality of life on their own terms.
-When the focus becomes nourishment rather than restriction, self-care rather than self-control, and progress rather than perfection, sustainable wellbeing becomes achievable.
+Biological mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered ghrelin and leptin signalling can heighten hunger, carbohydrate cravings, and appetite variability [@barryImpactEatingBehavior2011].
+Frequent hypoglycaemia and associated mood changes have also been observed, which can trigger compensatory eating or binge episodes [@barryImpactEatingBehavior2011].
+These physiological processes interact with psychological and social stressors, including infertility concerns, conflicting nutrition advice, chronic dieting, and exposure to idealised body images on social media, which together compound vulnerability to disordered eating [@steegers-theunissenPolycysticOvarySyndrome2020].
+Moreover, eating disorders themselves can disrupt endocrine function, potentially worsening PCOS symptoms and creating a self-reinforcing cycle [@cooneyIncreasedPrevalenceBinge2024; @bernadettPrevalenceEatingDisorders2016].
 
-Strengths and Limitations
-The current study has multiple strengths stemming from its preregistered, comprehensive methodology and Bayesian approach.
-This statistical framework allowed us to incorporate studies with and without control groups to better estimate mean REE in women with and without PCOS and to perform multiple sensitivity analyses that confirmed the stability of our findings.
-However, a limitation here is that variability in methods across studies, such as differences in PCOS diagnostic criteria and REE testing protocols, may have influenced results.
+Understanding the intertwined biological, psychological, and social influences suggests the importance of considering whether restrictive dietary advice is appropriate given it may exacerbate feelings of failure, hunger dysregulation, and shame [@pirottaDisorderedEatingBehaviours2019].
+These concerns, coupled with the lack of difference in REE between women with and without PCOS might suggest that energy restriction based dietary interventions for weight-loss may be unnecessary.
+But, there is also evidence supporting the effect of energy restricted dietary interventions for improving PCOS symptoms [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and they are recommended in international guidelines [@teedeRecommendations2023International2023].
+Encouragingly though, these guidelines also recognise weight stigma as a determinant of health and call for its reduction across clinical and public health settings.
+Evidently the greater prevalence of women with PCOS falling into overweight and obese BMI categories compared to women without PCOS [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021] is unlikely to be due to differences in REE and so, despite the potential effectiveness of energy restriction dietary interventions, there is potential value in moving towards more weight-neutral, individualised, and empowering care following holistic guidelines recommending multiple approaches to management [@teedeRecommendations2023International2023; @ozgensaydamWeightManagementStrategies2021].
+
+## Strengths and Limitations
+The current study has multiple strengths stemming from its preregistered, comprehensive methodology and Bayesian statistical approach.
+This statistical framework allowed us to incorporate studies with and without control groups to better estimate REE in women with and without PCOS and to perform multiple sensitivity analyses that confirmed the stability of our findings.
+However, a limitation here is that variability in methods across studies, such as differences in PCOS diagnostic criteria and REE testing protocols, may have influenced results and the relatively small number of studies overall limits the extent to which we can explore these potential moderators.
 Furthermore, some studies controlled for body weight or body composition when reporting REE values, while others did not.
-We accounted for this by estimating or converting reported data to obtain unadjusted REE values across all groups, thereby reducing this variability.
+We accounted for this by estimating or converting reported data to obtain unadjusted REE values across all groups, thereby reducing this variability and further as noted above provided estimates adjusted for BMI and fat-free mass both of which had little influence on our conclusions.
 Another limitation is that, due to fewer total control groups than PCOS groups, informed priors were required in several statistical models.
-Additionally, some studies reported data inconsistencies that could not be clarified (e.g., Georgopoulos et al.) and were accounted for in our analysis.
+However, in the context of Bayesian meta-analysis this can also be considered a strength.
+Additionally, some studies reported data inconsistencies that could not be clarified (e.g., [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009; @georgopoulosBasalMetabolicRate2009]) and were dropped from our main analysis though our conclusions again did not qualitative change when we conducted sensitivity analyses including these studies.
 Finally, most included studies were cross-sectional or baseline assessments within intervention trials, which limits causal inference.
+Indeed, we did not pre-register any kind of causal model (e.g., a directed acyclic graph) to inform our analysis approach for causal inference and as such have been explicit about the estimates presented as being descriptive.
 
-Conclusion
-In conclusion, the findings from this meta-analysis indicate that resting energy expenditure does not meaningfully differ between women with and without PCOS.
-Group-level differences in resting energy expenditure were small, insignificant, or not physiologically relevant.
-Additionally, variability in resting energy expenditure was also similar, suggesting that a lower baseline metabolic rate is not a causal factor for the weight-related challenges often associated with PCOS.
-These findings challenge the popular narrative that women with PCOS have a lower resting metabolism and may help better inform dietary interventions and nutritional support for these individuals.
+# Conclusion
+In conclusion, the findings from this meta-analysis indicate that REE does not meaningfully differ between women with and without PCOS.
+Group-level differences in REE were small, insignificant, or not physiologically relevant.
+Additionally, variability in REE between individuals was also similar.
+These results suggest that a lower baseline REE is not associated with the weight-related challenges often associated with PCOS.
+These findings challenge the popular narrative that women with PCOS have a lower REE and may help better inform dietary interventions and nutritional support for these individuals.
 Future research should include more standardized REE measurement and reporting protocols, greater data transparency, consistent control and reporting of body weight or body composition, the presentation of both absolute and relative REE, and more precise characterization of PCOS phenotypes.
-Overall, these findings support the conclusion that PCOS does not negatively influence resting energy expenditure and may help practitioners and researchers focus on targeted lifestyle interventions rather than negatively framed metabolic assumptions.
+Overall, these findings support the conclusion that PCOS is not negatively associated with REE and may help practitioners and researchers focus on individually targeted and holistic lifestyle interventions rather than negatively framed interventions based on unsupported assumptions regarding REE.
+
+#
 

### 2025-11-27 manuscript/pre_print.qmd
add ama csl
Check rendering

@@
   mylibrary.bib
-csl: apa.csl
+csl: american-medical-association.csl
 execute: 
@@
 
+```{r}
+targets::tar_load(c(main_arm_mean_effects_contrast_condition_bmi, main_arm_mean_effects_contrast_condition_fat_free_mass))
+
+```
+
 The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively).
@@
 
-#
+# Data Availability
+All code utilised for data preparation, transformations, analyses, plotting, and reporting are available in the corresponding GitHub repository <https://github.com/jamessteeleii/pcos_ree_meta>.
 

### 2025-11-27 manuscript/submission_manuscript.qmd
preprint version formatted, submission version to finish

--- 2025-11-27 manuscript/pre_print.qmd
@@
         \usepackage{lineno}\linenumbers
-        \usepackage{setspace}\doublespacing
+        \usepackage{setspace}\AtBeginDocument{\spacing{2}}
         \usepackage[font=scriptsize]{caption}
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
+        \usepackage{pdflscape}
 geometry:
@@
 Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate.
-Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors was used to compare both mean REE, and between person variation in REE, between women with and without PCOS.
+Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors to compare both mean REE, and between person variation in REE, between women with and without PCOS.
 Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33].
@@
 This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@pagePRISMA2020Statement2021a].
+The PRISMA flow diagram reported below (@fig-prisma) was produced using the `PRISMA2020` R package and Shiny app [@haddawayPRISMA2020PackageShiny2022].
 The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".
@@
 
+```{r}
+targets::tar_load(c(main_arm_mean_effects_contrast_condition_bmi, main_arm_variance_effects_contrast_condition_bmi,
+                    main_arm_mean_effects_contrast_condition_fat_free_mass,main_arm_variance_effects_contrast_condition_fat_free_mass))
+
+```
+
+Lastly, given the role of body mass and, in particular, highly metabolically active tissue on REE we included models adjusted for these group level characteristics where studies reported them.
+These models were essentially extensions of the main models noted above for BMI and fat-free mass where each was modelled with it's corresponding sampling error similarly to how the log mean REE was modelled in the models for standard deviation of REE i.e., the mean BMI or fat-free mass estimates were modelled as estimated with measurement error.
+When extracting posterior distributions for the contrasts between conditions in these models both BMI and fat-free mass where adjusted to the median values seen in the control conditions i.e., BMI = `r main_arm_mean_effects_contrast_condition_bmi$m_bmi` kg/$\textrm{m}^2$ and fat-free mass = `r main_arm_mean_effects_contrast_condition_fat_free_mass$m_fat_free_mass` kg.
+
 # Results
@@
 Note, the numbers in this section exclude the four studies previously noted with possible reporting issues that were unresolved.
-These studies are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al.
+These studies are however summarised in fully in the descriptive characteristics table in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html) where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al.
 [-@saltamavrosAlpha2Beta2007].
+The PRISMA flow diagram (@fig-prisma) also includes all studies identified including the four with possible errors.
+
+![PRISMA 2020 flow diagram template for systematic reviews.
+Note that a “report” could be a journal article, preprint, conference abstract, study register entry, clinical study report, dissertation, unpublished manuscript, government report or any other document providing relevant information.
+The website noted was a prior narrative review on this topic by some of the authors (https://macrofactorapp.com/pcos-bmr/) which identified two studies not found in our systematic database search.](prisma.png){#fig-prisma}
 
 Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
-Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.
+Descriptive characteristics of the arms and participants in these studies are reported fully in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)).
 
@@
 All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry.
-The specific devices reported by these studies are included in @tbl-review.
+The specific devices reported by these studies are included in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)).
 We originally considered in our pre-registration that, given sufficient data, we would compare sub groups of women with PCOS who did and did not have accompanying insulin resistance.
-However, based on the metabolic health variables reported in studies with this information (see @tbl-review) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance.
+However, based on the metabolic health variables reported in studies with this information (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance.
 Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging from `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`.
 Across those studies where BMI was reported or it was possible to estimate the mean BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]`.
-
-![Descriptive characteristics of arms and participants for included studies](descriptives_table.png){#tbl-review}
 
@@
 #| fig-width: 10
-#| fig-height: 7.5
+#| fig-height: 6.25
 #| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.
@@
 #| fig-width: 10
-#| fig-height: 7.5
+#| fig-height: 6.25
 #| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.
@@
 
+### BMI and Fat-Free Mass Adjusted Models
+For mean REE the model adjusted for BMI, also resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day with a 95% quantile interval ranging from `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)` kcal/day.
+This was also the case for the standard deviation of REE with this model resulting in a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(main_arm_variance_effects_contrast_condition_bmi$estimate),2)` and a 95% quantile interval ranging from `r round(exp(main_arm_variance_effects_contrast_condition_bmi$conf.low),2)` to `r round(exp(main_arm_variance_effects_contrast_condition_bmi$conf.high),2)`.
+This was similar for fat-free mass adjusted models too showing little difference in mean REE between control and PCOS conditions with mean point estimate of `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day with a 95% quantile interval ranging from `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)` kcal/day and a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$estimate),2)` and a 95% quantile interval ranging from `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$conf.low),2)` to `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$conf.high),2)`.
+
 # Discussion
@@
 Further, there is little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95%QI: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]) suggesting that, despite individual differences in REE, PCOS is not systematically associated with lesser or greater individual variability.
+
 These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain.
@@
 
-```{r}
-targets::tar_load(c(main_arm_mean_effects_contrast_condition_bmi, main_arm_mean_effects_contrast_condition_fat_free_mass))
-
-```
-
 The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively).
@@
 BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]` and this may explain the slightly greater REE in the former group.
-However, for those studies where we could extract or estimate BMI, we fit an additional exploratory model extending the main arm based model and extracted an estimate of the difference adjusted for BMI as a fixed effect.
-Adjusted to the median value of average BMI of the control arms of women without PCOS (BMI = `r main_arm_mean_effects_contrast_condition_bmi$m_bmi` kg/$\textrm{m}^2$) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)`kcal/day]) between women with PCOS and those without.
+However, for those studies where we could extract or estimate BMI, our additional exploratory models adjusted for this similarly showed little difference in REE (`r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)`kcal/day]) between women with PCOS and those without.
 However, one of the studies included in our analysis [@tosiRestingEnergyExpenditure2024] reported that, whilst there was little difference in unadjusted REE, when REE was adjusted for fat-free mass it was lower in women with PCOS suggesting the potential importance of fat-free tissue in energy regulation.
-Yet, similarly to the additional exploratory model including BMI, when we adjusted for fat-free mass including those studies reporting this characteristic (fat-free mass = `r main_arm_mean_effects_contrast_condition_fat_free_mass$m_fat_free_mass` kg) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)`kcal/day]).
+Yet, for those studies where fat-free mass was reported, our additional exploratory models adjusted for this similarly showed little difference in REE (`r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)`kcal/day]) between women with PCOS and those without.
 As such, even adjusted for both BMI and fat-free mass, there seems to be little difference in REE between women with, and without, PCOS.

### 2025-11-27 R/functions/main_data_functions.R

@@ -556,9 +556,9 @@ create_descriptives_table <- function(data) {
     ) |>
     bold(part = "header") |>
     # autofit() |>
-    fontsize(size = 8, part = "header") |>
-    fontsize(size = 8, part = "footer") |>
-    fontsize(size = 6, part = "body") |>
+    fontsize(size = 10, part = "header") |>
+    fontsize(size = 10, part = "footer") |>
+    fontsize(size = 8, part = "body") |>
     align(
       j = c(
         "Condition",
@@ -625,19 +625,17 @@ convert_descriptives_table_to_docx <- function(table) {
   print(doc, target = "tables/descriptives_table.docx")
 }
 
-convert_descriptives_table_to_png <- function(table) {
+convert_descriptives_table_to_html <- function(table) {
   
   # Save as image
-  save_as_image(table, path = "tables/descriptives_table.png", res=300)
+  save_as_html(table, path = "tables/descriptives_table.html",
+               title = "Descriptive characteristics of arms and participants for included studies")
   
-  save_as_image(table, path = "manuscript/descriptives_table.png", res=300)
 }
 
-convert_descriptives_table_to_svg <- function(table) {
-  
-  # Save as image
-  save_as_image(table, path = "tables/descriptives_table.svg", res=300)
-  
+convert_descriptives_table_to_pdf <- function(table_html) {
+  # Render to high-resolution PDF
+  webshot("tables/descriptives_table.html", "manuscript/descriptives_table.pdf", zoom = 2, vwidth = 2400)
 }
 
 # Pairwise data preparation for sensitivity analysis

### 2025-11-27 _targets.R

@@ -18,7 +18,8 @@ tar_option_set(
     "tidybayes",
     "patchwork",
     "flextable",
-    "officer"
+    "officer",
+    "webshot2"
   ),
   memory = "transient",
   format = "qs",
@@ -155,17 +156,12 @@ list(
     descriptives_table,
     create_descriptives_table(main_plus_greek_arm_data_imputed_demographics)
   ),
+  
+  tar_target(
+    descriptives_table_html,
+    convert_descriptives_table_to_html(descriptives_table)
+  ),
 
-  tar_target(
-    descriptives_table_png,
-    convert_descriptives_table_to_png(descriptives_table)
-  ),
-  
-  tar_target(
-    descriptives_table_svg,
-    convert_descriptives_table_to_svg(descriptives_table)
-  ),
-  
   # Fitting main analysis models
   
   tar_target(

### 2025-12-01 manuscript/pre_print.qmd
added author affiliation front matter and corresponding plus citation

@@
     # documentclass: article
+    keep-tex: true
+    template-partials: 
+     - title.tex
     pdf-engine: xelatex
@@
       text: |
-        \usepackage{lineno}\linenumbers
-        \usepackage{setspace}\doublespacing
         \usepackage[font=scriptsize]{caption}
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
+        \usepackage[noblocks]{authblk}
+        \renewcommand*{\Authsep}{, }
+        \renewcommand*{\Authand}{, }
+        \renewcommand*{\Authands}{, }
+        \renewcommand\Affilfont{\small}
 geometry:
@@
   - name: Richard Kirwan
-    affiliations:
-      - ref: 1
+    affil-id: 1
     orcid: 0000-0003-4645-0077
-    corresponding: true
     email: R.P.Kirwan@ljmu.ac.uk
   - name: Leigh Peele
-    affiliations:
-      - ref: 2
+    affil-id: 2
     email: leigh@macrofactorapp.com
   - name: Greg Nuckols
-    affiliations:
-      - ref: 2
+    affil-id: 2
     orcid: 0009-0004-2679-7967
@@
   - name: Georgia Kohlhoff
-    affiliations:
-      - ref: 3
-    # orcid: 0000-0002-8003-0757
+    affil-id: 3
     email: g_kohlhoff@hotmail.co.uk
   - name: Hannah Cabré
-    affiliations:
-      - ref: 4
+    affil-id: 4
     orcid: 0000-0002-3235-6867
@@
   - name: Alyssa Olenick
-    affiliations:
-      - ref: 5
+    affil-id: 5
     orcid: 0000-0002-2217-8067
-    corresponding: true
     email: alyssaolenickphd@gmail.com
   - name: James Steele
-    affiliations:
-      - ref: 6
-      - ref: 2
-      - ref: 7
+    affil-id: 6,2,7
     orcid: 0000-0002-8003-0757
@@
 Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.
+thanks: |
+  Preprint, please cite as: Kirwan, R., Peele, L., Nuckols, G., Kohlhoff, G., Cabré, H., Olenick, A., and Steele, J.
+(2025).
+Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis.
+medrχiv DOI: TO ADD.
+Address for correspondence: james@steele-research.com
 keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR

### 2025-12-01 manuscript/pre_print.tex
added author affiliation front matter and corresponding plus citation

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Latin Modern Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[font=scriptsize]{caption}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis},
  pdfauthor={Richard Kirwan; Leigh Peele; Greg Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
  pdfkeywords={Polycystic ovary syndrome, PCOS, resting energy
expenditure, REE, basal metabolic rate, BMR},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Resting energy expenditure of women with and without polycystic
ovary syndrome: a systematic review and meta-analysis\thanks{Preprint,
please cite as: Kirwan, R., Peele, L., Nuckols, G., Kohlhoff, G., Cabré,
H., Olenick, A., and Steele, J. (2025). Resting energy expenditure of
women with and without polycystic ovary syndrome: a systematic review
and meta-analysis. medrχiv DOI: TO ADD. Address for correspondence:
james@steele-research.com}}

\author[1]{Richard Kirwan}
\author[2]{Leigh Peele}
\author[2]{Greg Nuckols}
\author[3]{Georgia Kohlhoff}
\author[4]{Hannah Cabré}
\author[5]{Alyssa Olenick}
\author[6,2,7]{James Steele}

\affil[1]{Research Institute for Sports and Exercise Sciences, Liverpool
John Moores Universtiy, Liverpool, UK}
\affil[2]{MacroFactor, Stronger by Science LLC, Raleigh, North Carolina,
USA}
\affil[3]{Flourishing Health, Liverpool, UK}
\affil[4]{Reproductive Endocrinology and Women's Health Research
Program, Pennington Biomedical Research Center, Baton Rouge, Louisiana,
USA}
\affil[5]{Dr Alyssa Olenick LLC}
\affil[6]{Steele Research Limited, Eastleigh, Hampshire, UK}
\affil[7]{School of Health and Biomedical Sciences, Royal Melbourne
Institute of Technology, Melbourne, VIC, Australia}

\date{December 1, 2025}
\begin{document}
\maketitle
\begin{abstract}
Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age
women, who often have higher BMI classification. This is assumed to stem
from lower resting energy expenditure (REE), influencing lifestyle
intervention guidelines. However, evidence for reduced REE in women with
PCOS compared with those without is inconsistent. Objective: To
systematically search and meta-analyse the existing literature to
estimate and describe the difference in REE between women with and
without PCOS. Data Sources: A systematic search was conducted using
PubMed, Medline and Web of Science databases of published research from
January 1990 to January 2025. Study Selection: Studies that measured REE
in women living with PCOS, both with and without control arms of women
without PCOS, were included. Data Extraction: Bibliometric, demographic,
and REE data was extracted by one investigator and checked in
triplicate. Data Synthesis: Thirteen studies were included in a Bayesian
arm-based multiple condition comparison (i.e., network) type
meta-analysis model with informative priors to compare both mean REE,
and between person variation in REE, between women with and without
PCOS. Mean REE differed between groups by 30 kcal/day {[}95\% quantile
interval: -47 to 113 kcal/day{]} and the contrast ratio for between
person standard deviations was 0.98 {[}95\% quantile interval: 0.71 to
1.33{]}. Conclusions: These findings indicate that REE does not
meaningfully differ between women with and without PCOS. Group-level
differences in resting energy expenditure are small, insignificant, or
not physiologically relevant.
\end{abstract}

\section{Introduction}\label{introduction}

Polycystic ovary syndrome (PCOS) affects approximately 10\% of women of
reproductive age worldwide, making it the most common endocrine disorder
affecting this population\textsuperscript{1}. Due to several factors
including hyperandrogenism and alterations in insulin resistance, PCOS
is believed to contribute to an increased risk of diabetes, metabolic
syndrome and cardiovascular disease\textsuperscript{2--6}, along with
being a leading cause of anovulatory infertility in
women\textsuperscript{7}. Furthermore, epidemiological data has
consistently demonstrated that women with PCOS are significantly more
likely to suffer from overweight or obesity, compared to the general
female population, with estimates ranging from 38\% to 88\% of PCOS
patients falling into overweight or obese body mass index (BMI)
categories\textsuperscript{8,9}.

The elevated incidence of overweight and obesity in PCOS is likely
multifactorial with proposed mechanisms including blunted postprandial
appetite hormone responses leading to reduced satiety and increased food
cravings\textsuperscript{10--12} and a reduced resting energy
expenditure (REE)\textsuperscript{13}. Indeed, the study from
Georgopoulos et al.\textsuperscript{13} examined REE in women with and
without PCOS using indirect calorimetry and reported that resting energy
expenditure was approximately \textasciitilde400 kcal/day lower in women
with PCOS. Notably, they also reported that insulin resistance further
reduced REE among women with PCOS, with insulin-resistant women
exhibiting an additional reduction nearly 500 kcal per day compared to
women with PCOS who were not insulin resistant\textsuperscript{13}.
Other studies similarly report lower REE in women with PCOS using
indirect methods (such as prediction from bioelectrical impedance
analysis or accelerometer physical activity data), also suggesting that
factors including insulin resistance and BMI category influence REE in
women with PCOS\textsuperscript{14,15}. However, despite the widespread
acceptance that women living with PCOS exhibit reduced REE based on
studies such as these, other research has reported little to no
difference in REE between women with and without
PCOS\textsuperscript{16--18}.

The consequences of widespread acceptance that REE is substantially
lower in women living with PCOS should not be underestimated,
particularly in light of the aforementioned incidence of overweight and
obesity in this population. Women with PCOS typically engage in more
frequent weight-loss attempts than women without
PCOS\textsuperscript{19}. From a physiological perspective, if women
with PCOS do exhibit a lower REE, this could imply a meaningful
metabolic disadvantage that may influence dietary and nutritional
guidance for weight management; for example, recommending a slightly
more severe energy restriction to overcome the belief that they have a
lower REE\textsuperscript{20}. Recommendations such as this could
influence the well documented prevalence of eating disorders in women
with PCOS\textsuperscript{21,22}. Contrastingly, belief in a ``slower
metabolism'' could instead serve as a deterrent to energy restriction
based weight-loss approaches for some women in line with typical general
population recommendations that are also recognised as efficacious for
improving PCOS symptoms\textsuperscript{23,24} and are routinely
recommended\textsuperscript{25}. It has been well documented that women
with PCOS already experience higher rates of anxiety, depression, and
lower quality of life (QOL) as a result of negative body image and
weight-related concerns\textsuperscript{26--29}. Clarifying the
relationship between REE and PCOS may therefore help guide more accurate
clinical recommendations and empower both practitioners and women with
PCOS.

Therefore, to estimate and describe the magnitude of difference in REE
between women with and without PCOS, we completed a systematic review
and meta-analysis of studies reporting REE in these populations.

\section{Materials and Methods}\label{materials-and-methods}

This systematic review and meta-analysis was pre-registered on PROSPERO
(\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
initially on the \(3^{rd}\) of December 2024 and performed in accordance
with the Preferred Reporting Items for Systematic Reviews and
Meta-Analyses (PRISMA) statement guidelines\textsuperscript{30}. The
PRISMA flow diagram reported below (Figure~\ref{fig-prisma}) was
produced using the \texttt{PRISMA2020} R package and Shiny
app\textsuperscript{31}. The primary aim of this review was to examine
the descriptive question ``Does resting energy expenditure (REE) differ
between women with and without polycystic ovary syndrome (PCOS)?''. We
summarise and describe the studies in addition to quantitatively
synthesising their results via meta-analysis.

\subsection{Search Strategy}\label{search-strategy}

PubMed, Web of Science, and MEDLINE databases were searched using the
following Boolean search string: ((``Basal Metabolic Rate''{[}MeSH{]} OR
``Energy Metabolism''{[}MeSH{]} OR ``Resting Metabolic Rate'' OR RMR OR
``Resting Energy Expenditure'' OR REE OR ``Basal Metabolic Rate'' OR BMR
OR ``resting energy'' OR ``basal energy expenditure'') AND (``Polycystic
Ovary Syndrome''{[}MeSH{]} OR ``Polycystic Ovary Syndrome'' OR PCOS OR
``Polycystic Ovarian Disease'' OR ``Stein-Leventhal Syndrome'')).
Searches were limited to publications up until May 2025 when the search
was completed, limited to English language articles, and Rayyan was used
to manage the search and screening process. Two reviewers (RK and GK)
independently screened all titles and abstracts against the predefined
inclusion and exclusion criteria. Articles deemed potentially eligible
by either reviewer were retrieved in full text. Full texts were then
independently assessed by RK and LP to determine final eligibility. Any
disagreements at either stage were resolved through discussion, and when
consensus could not be reached, a third reviewer acted as an
adjudicator.

\subsection{Eligibility Criteria}\label{eligibility-criteria}

Studies were included in the systematic review if 1) participants were
confirmed as women with PCOS between the ages of 18 to 65 years of age
with or without insulin resistance; 2) otherwise healthy (e.g.,
non-diabetic, no cardiovascular disease); 3) had a measure of REE
measured via multiple methods including direct/indirect calorimetry,
doubly labelled water; and 4) trials were not retracted at the time of
this analysis. Studies were excluded if they 1) used invalid or
non-standard methods for measuring REE (e.g., predicted REE from body
composition or accelerometer data); 2) non-peer-reviewed journal
articles (including grey literature sources such as conference
abstracts, theses and dissertations); and 3) were secondary analyses
with the same primary outcome data as another included study.

The condition being studied was PCOS and we included observational
cross-sectional design studies, in addition to intervention studies
where REE was reported for the population (and if present, the
comparator i.e., women without PCOS) condition of interest. For clarity,
studies of any design were included if they reported the REE using the
methods indicated for a sample of adult women with PCOS and who were
otherwise healthy. This included both studies with and without samples
of healthy control women without PCOS. As detailed in the statistical
analysis section below, a Bayesian model with informative priors based
on normative data for REE in healthy women without PCOS was included to
provide control information indirectly where this was missing. The use
of such priors is an efficient tool for incorporating historical
information on a particular population in a conservative
manner\textsuperscript{32}.

Following the PICO framework our eligibility criteria can be defined as
follows:

\begin{itemize}
\tightlist
\item
  Population

  \begin{itemize}
  \tightlist
  \item
    Inclusion criteria:

    \begin{itemize}
    \tightlist
    \item
      Women
    \item
      18-65 y
    \item
      With or without insulin resistance (IR)
    \end{itemize}
  \end{itemize}
\item
  Intervention(s) or exposure(s)

  \begin{itemize}
  \tightlist
  \item
    Otherwise healthy women with PCOS
  \end{itemize}
\item
  Comparator(s) or control(s)

  \begin{itemize}
  \tightlist
  \item
    Otherwise healthy control women without PCOS
  \end{itemize}
\item
  Outcome

  \begin{itemize}
  \tightlist
  \item
    Inclusion criteria

    \begin{itemize}
    \tightlist
    \item
      Resting energy expenditure (REE) measured via multiple methods
      including direct/indirect calorimetry, doubly labelled water.
    \end{itemize}
  \item
    Exclusion criteria:

    \begin{itemize}
    \tightlist
    \item
      Studies using invalid or non-standard methods for measuring REE
      (e.g., predicted REE from body composition or accelerometer data)
    \end{itemize}
  \end{itemize}
\end{itemize}

\subsection{Data extraction (selection and
coding)}\label{data-extraction-selection-and-coding}

Data was extracted by one investigator and checked in triplicate.
Bibliometric data including authors, journal, and article titles were
extracted. Descriptive statistics for age, body mass, fat mass, fat free
mass, height, BMI, race, physical activity levels, country of
investigation, information regarding glucose/insulin regulation and
insulin resistance status (where available), diagnostic criteria for
PCOS, and measurement method and device were extracted for each arm
within each study in addition to sample size. Descriptive
characteristics were then tabulated across studies for reporting.

For each arm, and observation time point if multiple observations
reported (e.g., before and after an intervention), depending on what was
reported by the authors we extracted the means, medians, standard
deviations, standard errors, lower and upper range values, and
interquartile range for the unadjusted and/or body mass adjusted and/or
fat free mass adjusted REE values. Where REE values adjusted for body
mass and/or fat free mass were reported we used the reported body mass
and/or fat free mass mean values for that arm to convert them to
unadjusted REE values (i.e., multiplied them by body mass and/or fat
free mass mean values). Where means and/or standard deviations were
missing the latter were either calculated from standard errors and
sample size, or all both were estimated from lower and upper range,
interquartile range, median, and sample size depending on the available
information using the methods of Wan et al.\textsuperscript{33}.
Further, where missing, height/body mass/BMI where estimated based on
the reported means. The units of measurement for which REE was extracted
and all REE values were converted to kcal/day. In one
case\textsuperscript{34} REE was reported relative to body mass and the
unadjusted values were no longer available (confirmed by the authors).
As such, in this case we used the mean body mass to convert back to
estimated REE unadjusted.

\subsubsection{Studies with possible reporting
errors}\label{studies-with-possible-reporting-errors}

During data extraction it was noted that several studies from the same
lab/research group\textsuperscript{13,35--37} contained a number of
discrepancies that seemed to be possible reporting errors. This
included, based on taking the authors results as written, standard
errors that implied impossible or at least incredibly unlikely standard
deviations, and discrepancies in sample size reporting throughout for
most variables without explanation or where this was explained the
sample sizes were discrepant with the text Further, data was not
reported for the healthy control women without PCOS in three of the
studies\textsuperscript{35--37}, and REE was reported as an ``adjusted''
value whereby
\(\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})\)
and the \(\textrm{REE}_{predicted}\) was obtained by substituting the
individual lean body mass, fat mass, gender, and age in the linear
regression equation generated by the data of all patients. In
correspondence with the senior author we were unable to clarify the
reporting discrepancies as the person responsible for the data/results
was no longer contactable. The original data were also no longer
available and so we could not calculated the unadjusted REE.

Given these issues we decided to extract the results from these studies
as reported and to conduct analyses both with and without their
inclusion. Though not pre-registered, due to a lack of confidence in the
reported results, we decided to include the analysis omitting these
studies as our main models in the results reported below. The results of
the analysis including them are reported in the sensitivity analysis
section.

\subsection{Statistical Analysis}\label{statistical-analysis}

Statistical analysis of the data extracted was be performed in R, (v
4.3.3; R Core Team, https://www.r-project.org/) and RStudio (v
2023.06.1; Posit, https://posit.co/). All code utilised for data
preparation, transformations, analyses, plotting, and reporting are
available in the corresponding GitHub repository
\url{https://github.com/jamessteeleii/pcos_ree_meta}. We cite all
software and packages used in the analysis pipeline using the
\texttt{grateful} package\textsuperscript{38} which can be seen here:
\url{https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf}.
The statistical analysis plan was linked in our pre-registration
(PROSPERO:
\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
and available at the accompanying GitHub repository. Any deviations from
the pre-registration are noted below.

Given our research question our analysis was aimed at parameter
estimation\textsuperscript{39} within a Bayesian meta-analytic
framework\textsuperscript{40}. For all analyses model parameter
estimates and their precision, along with conclusions based upon them,
are interpreted continuously and probabilistically, considering data
quality, plausibility of effect, and previous literature, all within the
context of each model. The \texttt{renv} package\textsuperscript{41} was
used for package version reproducibility and a function based analysis
pipeline using the \texttt{targets} package\textsuperscript{42} was
employed (the analysis pipeline can be viewed by downloading the R
Project and running the function \texttt{targets::tar\_visnetwork()}).
Effect sizes and their variances were all calculated using the
\texttt{metafor} packages \texttt{escalc()}
function\textsuperscript{43}. The main package
\texttt{brms}\textsuperscript{44} was used in fitting all the Bayesian
meta-analysis models. Prior and posterior draws were taken using
\texttt{marginaleffects}\textsuperscript{45} and
\texttt{tidybayes}\textsuperscript{46} packages. All visualisations are
created using \texttt{ggplot2}\textsuperscript{47}, \texttt{tidybayes},
and the \texttt{patchwork}\textsuperscript{48} packages.

\subsubsection{Main Pre-registered
Models}\label{main-pre-registered-models}

We adopted an arm-based multiple condition comparison (i.e., network)
type model given that the studies included had arms of women with PCOS
both with, and without, a non-PCOS control arm\textsuperscript{49}, and
also in some cases multiple observations of REE in the different arms
included in the study (for example, where an intervention was conducted
and pre- and post-intervention REE was reported). In typical
contrast-based meta-analyses data is limited to the effect sizes for
paired contrasts between arms and thus studies that include both arms
(i.e., relative effects between non-PCOS control vs PCOS arms); however,
in arm-based analyses the data are the absolute effects within each arm
and information is borrowed across studies to enable both within
condition absolute, and between condition relative contrasts to be
estimated. We made use of historical information regarding REE in
healthy control women without PCOS by setting informative priors based
on meta-analysis of large scale studies reporting normative data for REE
in this population. This was included to provide indirect control
information where it was missing from particular studies. The use of
historical priors like this is an efficient tool to incorporate
historical information about a particular population in a conservative
manner in meta-analyses\textsuperscript{32}. From this model we focus on
reporting the global grand mean estimate for the fixed between condition
relative contrast for non-PCOS control vs PCOS arms as our primary
estimand of interest (i.e., \(\beta_1\) in both mean and standard
deviation models). We examined both raw mean REE (i.e., the absolute
mean REE in kcals per day for each arm) in addition to the between
person standard deviation in REE (i.e., the absolute standard deviation
in REE in kcals per day for each arm). Both models were multilevel in
that they included nested random intercepts for both study and arm
within study. In addition, and in deviation from our pre-registration,
we also included lab as a random intercept as in some cases we had
multiple studies from the same lab or research group. Lastly, the
inclusion of a random intercept for each effect size was accidentally
omitted from our pre-registration, and so this is also included in the
model.

\paragraph{Mean REE Model}\label{mean-ree-model}

The main model for mean REE with \(\textrm{cond}\) representing the
condition (either control or PCOS) was as follows: \[
\scriptsize
\begin{aligned}
\hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl]
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[10pt]
\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}
\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix}
\text{, for study j = 1,} \dots \text{,J} \\[12pt]
\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
\text{, for effect l = 1,} \dots \text{,L}
\end{aligned}
\]

where \(\hat{\theta}_{ijkl}\) is the \(l\textrm{th}\) mean REE estimate
from the \(k\textrm{th}\) arm, for the \(j\textrm{th}\) study, conducted
by the \(i\textrm{th}\) lab and \(\sigma^2_{ijkl}\) is the corresponding
sampling error for that estimate. The random intercepts for the
\(i\textrm{th}\) lab, \(j\textrm{th}\) study, \(k\textrm{th}\) arm, and
\(l\textrm{th}\) mean REE estimate are \(\alpha_{0,lab[i]}\),
\(\alpha_{0,study[j]}\), \(\alpha_{0,arm[k]}\), and
\(\alpha_{0,effect[l]}\) respectively each with standard deviation of
\(\sigma^2_{0,lab[i]}\), \(\sigma^2_{0,study[j]}\),
\(\sigma^2_{0,arm[k]}\), and \(\sigma^2_{0,effect[l]}\). The parameter
\(\beta_0\) represents the fixed effect estimate of REE for control
conditions and \(\beta_1\) the fixed effect estimate for the offset from
this for the PCOS conditions (i.e., the difference between conditions).
The estimated offset was allowed to vary across both labs and studies
each reflected by \(\alpha_{1,lab[i]}\) and \(\alpha_{1,study[j]}\)
respectively, and these effects were also modelled as correlated with
the corresponding random intercepts with covariance \(\Sigma_{lab}\) and
\(\Sigma_{study}\), and \(\mathrm{corr}_{\mathrm{lab}}\) and
\(\mathrm{corr}_{\mathrm{study}}\) correlation matrices.

The priors for this model were as follows:

\[
\small
\begin{aligned}
\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]
\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]
\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{aligned}
\] where the prior for \(\beta_{0}\), which corresponded to the model
intercept and mean REE in the control condition, was set based on
meta-analysis of the mean REEs for women from two large studies of
healthy people\textsuperscript{50,51} though set with a conservative
degrees of freedom for the \(\textrm{Student-}t\) distribution.The
random intercept \(\sigma^2_{0,study}\) was set similarly to this. The
prior for the fixed effect \(\beta_{1}\), reflecting the difference
between control and PCOS conditions was set based on a wide range of
possible values considering the minimum and and maximum values of the
ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We
then set a prior that permits values approximately across this range of
values with the majority of it's mass centred around zero. The remaining
random effects were set based on the default weakly regularising priors
for \texttt{brms} and scaled to the expected response values using a
\(\textrm{half-student-}t\) distribution with 3 degrees of freedom and
\(\mu=0\), and both correlation matrices \(\textrm{corr}_{lab}\) and
\(\textrm{corr}_{study}\) were set with an \(\textrm{LKJcorr(1)}\)
distribution.

\paragraph{Standard Deviation of REE
Model}\label{standard-deviation-of-ree-model}

The main model for the standard deviation of REE with \(\textrm{cond}\)
representing the condition (either control or PCOS) was as follows:

\[
\scriptsize
\begin{aligned}
\log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]
\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]}
       + \beta_2\,\tilde{m}_{[ijkl]}
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[8pt]
m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt]
\tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt]
\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}
\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix}
\text{, for study j = 1,} \dots \text{,J} \\[12pt]
\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
\text{, for effect l = 1,} \dots \text{,L}
\end{aligned}
\]

where \(\log(\hat{\theta}_{ijkl})\) is the \(l\textrm{th}\) natural
logarithm of the standard deviation of REE estimate from the
\(k\textrm{th}\) arm, for the \(j\textrm{th}\) study, conducted by the
\(i\textrm{th}\) lab and \(\sigma^2_{ijkl}\) is the corresponding
sampling error for that estimate. The random intercepts for the
\(i\textrm{th}\) lab, \(j\textrm{th}\) study, \(k\textrm{th}\) arm, and
\(l\textrm{th}\) standard deviation of REE estimate are
\(\alpha_{0,lab[i]}\), \(\alpha_{0,study[j]}\), \(\alpha_{0,arm[k]}\),
and \(\alpha_{0,effect[l]}\) respectively each with standard deviation
of \(\sigma^2_{0,lab[i]}\), \(\sigma^2_{0,study[j]}\),
\(\sigma^2_{0,arm[k]}\), and \(\sigma^2_{0,effect[l]}\). The parameter
\(\beta_0\) represents the fixed effect estimate of standard deviation
of REE for control conditions and \(\beta_1\) the fixed effect estimate
for the offset from this for the PCOS conditions (i.e., the difference
between conditions). The estimated offset was allowed to vary across
both labs and studies each reflected by \(\alpha_{1,lab[i]}\) and
\(\alpha_{1,study[j]}\) respectively, and these effects were also
modelled as correlated with the corresponding random intercepts with
covariance \(\Sigma_{lab}\) and \(\Sigma_{study}\), and
\(\mathrm{corr}_{\mathrm{lab}}\) and \(\mathrm{corr}_{\mathrm{study}}\)
correlation matrices. Finally, \(\beta_{2}\) represents the fixed effect
of the natural logarithm of the corresponding mean REE estimate
\(\tilde{m}\) which is modelled as estimated with measurement error
i.e., \(m\) represents the point estimate for the \(l\textrm{th}\)
natural logarithm of the mean REE estimate from the \(k\textrm{th}\)
arm, for the \(j\textrm{th}\) study, conducted by the \(i\textrm{th}\)
lab and \(\sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2\) is the
corresponding sampling error for that estimate.

The priors for this model were as follows: \[
\small
\begin{aligned}
\beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt]
\beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt]
\text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt]
\text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt]
\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt]
\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{aligned}
\]

where the prior for \(\beta_{0}\), which corresponded to the model
intercept and standard deviation of REE in the control condition, was
again set based on meta-analysis of the standard deviation of REEs for
women from two large studies of healthy people\textsuperscript{50,51}
though set with a conservative degrees of freedom for the
\(\textrm{Student-}t\) distribution.The random intercept
\(\sigma^2_{0,study}\) was set similarly to this. The prior for the
fixed effect \(\beta_{1}\), reflecting the difference between control
and PCOS conditions was set based on a wide range of possible values.
Given that in many cases of variables in the field there is an
approximate relationship of \textasciitilde1 for the natural logarithm
of the standard deviation conditioned upon the natural logarithm of the
mean\textsuperscript{52} we set this prior to reflect the range of
differences on the on the log scale (i.e., \(\log(1584)\)). We then set
a prior that permits values approximately across this range of values
with the majority of it's mass centred around zero. The prior for the
fixed effect \(\beta_{2}\), reflecting the relationship between the
natural logarithm of the mean REE with the natural logarithm of the
standard deviation of REE, was set it to be weakly informative centred
on zero with a wide scale to indicate uncertainty in this outcome
specifically despite the typical relationship close to \textasciitilde1.
Priors for the measurement error of the natural logarithm of the mean
REE estimate \(\tilde{m}\) were again based upon meta-analysis of the
aforementioned studies, though measurement error has to be positive, as
does the corresponding standard deviation of this error, so we set these
to conservative wide \(\textrm{half-student-}t\) distributions. The
remaining random effects were set based on the default weakly
regularising priors for \texttt{brms} and scaled to the expected
response values using a \(\textrm{half-student-}t\) distribution with 3
degrees of freedom and \(\mu=0\), and both correlation matrices
\(\textrm{corr}_{lab}\) and \(\textrm{corr}_{study}\) were set with an
\(\textrm{LKJcorr(1)}\) distribution.

\paragraph{Post-processing of models}\label{post-processing-of-models}

For both models we examined trace plots along with \(\hat{R}\) values to
examine whether chains have converged, and posterior predictive checks
for each model to understand the model implied distributions. From each
model we took draws from the posterior distributions for the conditional
absolute estimates for each condition (i.e., controls and PCOS) by study
incorporating random effects, the global grand mean absolute estimates
for each condition ignoring random effects, and the global grand mean
between condition relative contrast for controls vs PCOS conditions
ignoring random effects. The between condition relative contrast for
controls vs PCOS conditions corresponded to \(\beta{1}\) in each model
and was our primary estimand of interest; for the mean REE model this
corresponded to the absolute difference in mean REE, and for the
standard deviation of REE model this corresponded to the natural
logarithm of the ratio of standard deviations of REE which was
exponentiated (note, all log standard deviation of REE model estimates
were exponentiated back to the original scale to aid interpretibility).
We present the full probability density functions for posterior
visually, and also to calculate mean and 95\% quantile intervals (QI:
i.e., `credible' or `compatibility' intervals) for each estimate
providing the most probable value of the parameter in addition to the
range from 2.5\% to 97.5\% percentiles given our priors and data.

\subsubsection{Sensitivity analyses}\label{sensitivity-analyses}

\paragraph{Pairwise contrast based
models}\label{pairwise-contrast-based-models}

By way of pre-registered sensitivity analysis we also conducted pairwise
contrast based models where we limited the included effects to those
extracted from studies including only a directly comparable control and
PCOS arm at baseline. These models were both as follows:

\[
\small
\begin{aligned}
\hat{\theta}_{ij} &\sim \mathcal{N}(\mu_{ij}, \sigma^2_{ij}) \\[6pt]
\mu_{ij} &= \beta_0
       + \alpha_{0,\mathrm{lab}[i]}
       + \alpha_{0,\mathrm{study}[j]} \\[10pt]
\alpha_{0,\mathrm{lab}[i]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{lab}})
\text{, for lab i = 1,} \dots \text{,I} \\[12pt]
\alpha_{0,\mathrm{study}[j]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{study}})
\text{, for study j = 1,} \dots \text{,J}
\end{aligned}
\]

where \(\hat{\theta}_{ijkl}\) is the pairwise effect size, either the
mean difference in REE or the log coefficient of variation ratio
(calculated as PCOS vs control), for the \(j\textrm{th}\) study,
conducted by the \(i\textrm{th}\) lab and \(\sigma^2_{ijkl}\) is the
corresponding sampling error for that effect size estimate. The random
intercepts for the \(i\textrm{th}\) lab, \(j\textrm{th}\) study are
\(\alpha_{0,lab[i]}\) and \(\alpha_{0,study[j]}\) respectively each with
standard deviation of \(\sigma^2_{0,lab[i]}\),
\(\sigma^2_{0,study[j]}\). The parameter \(\beta_0\) represents the
fixed effect estimate of the pairwise effect size i.e., the pooled
estimate of the contrast between conditions. The priors for
\(\beta_{0}\) in these models were set as default weakly regularising
which is set on an intercept that results when internally centering all
population-level predictors around zero to improve sampling efficiency
and scaled to the expected response values using a
\(\textrm{student-}t\) distribution; for the mean difference in REE this
was \(\textrm{student-}t(3,\ 5.5,\ 50.2)\) and for the log coefficient
of variance ratio this was \(\textrm{student-}t(3,\ -0.1,\ 2.5)\). The
random effects were set similarly scaled to the expected response values
but using a \(\textrm{half-student-}t\) distribution centred on zero.
From these models we calculated the mean and 95\% quantile intervals
(i.e., `credible' or `compatibility' intervals) for the \(\beta_0\)
(comparable to the \(\beta_1\) from the corresponding mean and standard
deviation of REE arm-based models) for each effect size providing the
most probable value of the parameter in addition to the range from 2.5\%
to 97.5\% percentiles given our priors and data.

\paragraph{Additional sensitivity
analyses}\label{additional-sensitivity-analyses}

As noted in the section above, \emph{``Studies with possible reporting
errors''}, we opted to conduct analysis with and without the inclusion
of four studies with possible reporting errors we could not
resolve\textsuperscript{13,35--37}. Thus the main models described above
were run with and without these studies, the main results are presented
without them and the sensitivity results with them are reported
separately below.

As a further sensitivity analysis, given the inclusion of some studies
with interventions in women with PCOS reporting REE at multiple
timepoints such as mid- and post-intervention\textsuperscript{34,53,54}
but lacking of control women without PCOS, we opted to also conduct
sensitivity analysis excluding these and only examining their baseline
results in addition to the cross-sectional studies. As such, the main
models described above were run without the follow-up (i.e., mid- or
post-intervention timepoints) from these studies and only using the
baseline results in addition to other cross-sectional studies.

Lastly, given the role of body mass and, in particular, highly
metabolically active tissue on REE we included models adjusted for these
group level characteristics where studies reported them. These models
were essentially extensions of the main models noted above for BMI and
fat-free mass where each was modelled with it's corresponding sampling
error similarly to how the log mean REE was modelled in the models for
standard deviation of REE i.e., the mean BMI or fat-free mass estimates
were modelled as estimated with measurement error. When extracting
posterior distributions for the contrasts between conditions in these
models both BMI and fat-free mass where adjusted to the median values
seen in the control conditions i.e., BMI = 26.03 kg/\(\textrm{m}^2\) and
fat-free mass = 48.8 kg.

\section{Results}\label{results}

\subsection{Systematic Review}\label{systematic-review}

Note, the numbers in this section exclude the four studies previously
noted with possible reporting issues that were unresolved. These studies
are however summarised in fully in the descriptive characteristics table
in the online supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table} where some of the possible reporting errors are seen in the
standard deviations reported or calculated e.g., the standard deviation
calculated for body mass of women with PCOS in Saltamavros et
al.\textsuperscript{36}. The PRISMA flow diagram
(Figure~\ref{fig-prisma}) also includes all studies identified including
the four with possible errors.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{prisma.png}}

}

\caption{\label{fig-prisma}PRISMA 2020 flow diagram template for
systematic reviews. Note that a ``report'' could be a journal article,
preprint, conference abstract, study register entry, clinical study
report, dissertation, unpublished manuscript, government report or any
other document providing relevant information. The website noted was a
prior narrative review on this topic by some of the authors
(https://macrofactorapp.com/pcos-bmr/) which identified two studies not
found in our systematic database search.}

\end{figure}%

Our systematic review identified 13 studies from 12 lab/research groups
including 24 arms (Control arms = 9, PCOS arms = 15) and a total of 918
participants (Controls: minimum n = 9, median n = 29, maximum n = 54;
PCOS: minimum n = 5, median n = 28, maximum n = 266). Descriptive
characteristics of the arms and participants in these studies are
reported fully in the online supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}).

Studies were carried out in multiple countries: Brazil (k = 3) ,and USA
(k = 3 studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK
(all k = 1 study). The four studies with noted reporting errors were
carried out in Greece. A total of 11 studies used the Rotterdam criteria
(or a modified version there of) for diagnosing
PCOS\textsuperscript{16,17,34,53--60}. One study used the 1990 National
Institutes of Health criteria\textsuperscript{61}, and two
studies\textsuperscript{18,55} diagnosed PCOS via the presence of
oligomenorrhea or amenorrheaammenorrhea alongside additional criteria
including plasma androgen levels, hirsutism or polycystic ovaries on
ultrasound scanning. All but one study\textsuperscript{61}, which used
doubly labelled water, measured REE using indirect calorimetry. The
specific devices reported by these studies are included in the online
supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}). We originally considered in our pre-registration that, given
sufficient data, we would compare sub groups of women with PCOS who did
and did not have accompanying insulin resistance. However, based on the
metabolic health variables reported in studies with this information
(see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}) and, where available, considering primary criteria of either
homeostatic model assessment of insulin resistance (HOMA-IR) \(\geq2.5\)
or secondary criteria including fasting insulin \(>12\mu \textrm{U/mL}\)
or fasting glucose \(\geq100\textrm{mg/dL}\), all groups of women with
PCOS in the included studies would be considered to have insulin
resistance. Mean age of women with PCOS in these studies ranged from 23
to 33 which was similar to the control women without PCOS ranging from
23 to 30. Across those studies where BMI was reported or it was possible
to estimate the mean BMI of women with PCOS ranged from 26.4 to 39.9 was
typically greater than women without PCOS which ranged 20.5 to 27.9.

\subsection{Mean REE Model Results}\label{mean-ree-model-results}

The main model for mean REE resulted in a posterior distribution for the
contrast between control and PCOS conditions with a mean point estimate
of 30 kcal/day with a 95\% quantile interval ranging from -47 kcal/day
to 113 kcal/day suggesting there is a 95\% probability that the true
difference lies between these values given our priors and the data from
included studies. The corresponding conditional estimates for the
control condition and PCOS condition respectively were 1442 kcal/day
{[}95\%QI:1334 kcal/day to 1553 kcal/day{]} and 1472 kcal/day
{[}95\%QI:1359 kcal/day to 1587 kcal/day{]}. These results, including
the full visualisation of the posterior distribution and the conditional
estimates by study, can be seen in Figure~\ref{fig-mean-REE}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{pre_print_files/figure-pdf/fig-mean-REE-1.pdf}}

}

\caption{\label{fig-mean-REE}Posterior distribution, mean point
estimates, and 95\% quantile intervals for conditional estimates by
study, global grand mean estimates by condition, and the contrast
between conditions for mean resting energy expenditure of control women
without PCOS and women with PCOS.}

\end{figure}%

\subsection{Standard Deviation of REE Model
Results}\label{standard-deviation-of-ree-model-results}

The main model for the between participant standard deviation of REE
resulted in a posterior distribution for the contrast ratio between
control and PCOS conditions with a mean point estimate of 0.98 with a
95\% quantile interval ranging from 0.71 to 1.33 suggesting there is a
95\% probability that the true ratio of standard deviations lies between
these values given our priors and the data from included studies. The
corresponding conditional estimates for the standard deviations of the
control condition and PCOS condition respectively were 238 kcal/day
{[}95\%QI:178 kcal/day to 312 kcal/day{]} and 229 kcal/day {[}95\%QI:169
kcal/day to 303 kcal/day{]}. These results, including the full
visualisation of the posterior distribution and the conditional
estimates by study, can be seen in Figure~\ref{fig-variance-REE}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{pre_print_files/figure-pdf/fig-variance-REE-1.pdf}}

}

\caption{\label{fig-variance-REE}Posterior distribution, variance point
estimates, and 95\% quantile intervals for conditional estimates by
study, global grand variance estimates by condition, and the contrast
ratio between conditions for the standard deviation of resting energy
expenditure of control women without PCOS and women with PCOS.}

\end{figure}%

\subsection{Sensitivity Analyses}\label{sensitivity-analyses-1}

\subsubsection{Pairwise Models}\label{pairwise-models}

For mean REE the pairwise model resulted in qualitatively similar
inferences suggesting little difference between control and PCOS
conditions with mean point estimate of 16 kcal/day with a 95\% quantile
interval ranging from -36 kcal/day to 70 kcal/day. This was similar for
the standard deviation of REE with the pairwise model resulting in a
contrast ratio between the control and PCOS conditions with a mean point
estimate of 0.9 and a 95\% quantile interval ranging from 0.6 to 1.35.

\subsubsection{Models including studies with possible reporting
issues}\label{models-including-studies-with-possible-reporting-issues}

For mean REE the model including the studies noted with possible
reporting issues that were not resolved\textsuperscript{13,35--37} still
resulted in qualitatively similar inferences suggesting little
difference between control and PCOS conditions with mean point estimate
of 15 kcal/day with a 95\% quantile interval ranging from -69 kcal/day
to 95 kcal/day. This was similar for the standard deviation of REE with
the model including these studies resulting in a contrast ratio between
the control and PCOS conditions with a mean point estimate of 1.21 and a
95\% quantile interval ranging from 0.68 to 2.13.

\subsubsection{Basline REE Measurement
Models}\label{basline-ree-measurement-models}

For mean REE the model which included only baseline REE measurements
from studies involving interventions\textsuperscript{34,53,54}, in
addition to other cross-sectional studies, also resulted in
qualitatively similar inferences suggesting little difference between
control and PCOS conditions with mean point estimate of 32 kcal/day with
a 95\% quantile interval ranging from -46 kcal/day to 118 kcal/day. This
was also the case for the standard deviation of REE with this model
resulting in a contrast ratio between the control and PCOS conditions
with a mean point estimate of 0.97 and a 95\% quantile interval ranging
from 0.71 to 1.36.

\subsubsection{BMI and Fat-Free Mass Adjusted
Models}\label{bmi-and-fat-free-mass-adjusted-models}

For mean REE the model adjusted for BMI, also resulted in qualitatively
similar inferences suggesting little difference between control and PCOS
conditions with mean point estimate of 20 kcal/day with a 95\% quantile
interval ranging from -96 kcal/day to 140 kcal/day. This was also the
case for the standard deviation of REE with this model resulting in a
contrast ratio between the control and PCOS conditions with a mean point
estimate of 0.91 and a 95\% quantile interval ranging from 0.56 to 1.41.
This was similar for fat-free mass adjusted models too showing little
difference in mean REE between control and PCOS conditions with mean
point estimate of -9 kcal/day with a 95\% quantile interval ranging from
-195 kcal/day to 168 kcal/day and a contrast ratio between the control
and PCOS conditions with a mean point estimate of 0.79 and a 95\%
quantile interval ranging from 0.31 to 1.94.

\section{Discussion}\label{discussion}

This study sought to estimate and describe the magnitude of difference
in REE between women with and without PCOS. Most studies identified in
the systematic review, and included in the meta-analysis, used indirect
calorimetry as the primary measure of REE and assessed women with PCOS
who were insulin resistant and categorised as being in overweight or
obese BMI categories compared to healthy controls. Our results indicate
there is only a small magnitude of difference in REE (30 kcal/day
{[}95\%QI: -47 kcal/day to 113kcal/day{]}) between women with PCOS and
those without. Further, there is little difference in between person
variation between the groups based on the ratio of standard deviations
(0.98 {[}95\%QI: 0.71 to 1.33{]}) suggesting that, despite individual
differences in REE, PCOS is not systematically associated with lesser or
greater individual variability.

These findings challenge the widely held belief that PCOS is inherently
associated with a slower metabolism (PMID: 32244780), predisposing women
with PCOS to weight gain. This belief may mistakenly lead to
recommendations centred on slightly more severe calorie restriction to
achieve weight-loss goals, compared with recommendations for the general
population, as primary management strategies for women with
PCOS\textsuperscript{20}. Recognising that there may be minimal
differences in REE between women with PCOS and those without can inform
both clinical and public practices, potentially leading to a shift in
focus away from a requirement for more severe caloric restriction as a
primary method of treatment towards more comprehensive, individualised,
and psychologically safe approaches to care\textsuperscript{25}.

The current study suggests that REE may not a barrier to weight
regulation in PCOS given small group-level differences between women
with PCOS and those without (229 kcal/day {[}95\%QI:169 kcal/day to 303
kcal/day{]} versus 238 kcal/day {[}95\%QI:178 kcal/day to 312
kcal/day{]}, respectively). If anything, REE may be slightly higher in
women with PCOS compared to healthy women without PCOS. BMI of women
with PCOS ranged from 26.4 to 39.9 was typically greater than women
without PCOS which ranged 20.5 to 27.9 and this may explain the slightly
greater REE in the former group. However, for those studies where we
could extract or estimate BMI, our additional exploratory models
adjusted for this similarly showed little difference in REE (20 kcal/day
{[}95\%QI: -96 kcal/day to 140kcal/day{]}) between women with PCOS and
those without. However, one of the studies included in our
analysis\textsuperscript{62} reported that, whilst there was little
difference in unadjusted REE, when REE was adjusted for fat-free mass it
was lower in women with PCOS suggesting the potential importance of
fat-free tissue in energy regulation. Yet, for those studies where
fat-free mass was reported, our additional exploratory models adjusted
for this similarly showed little difference in REE (-9 kcal/day
{[}95\%QI: -195 kcal/day to 168kcal/day{]}) between women with PCOS and
those without. As such, even adjusted for both BMI and fat-free mass,
there seems to be little difference in REE between women with, and
without, PCOS. Yet, a recent systematic review of mechanisms for
metabolic dysfunction has reported excess androgen drives metabolic
issues within adipose tissue and muscle tissue contributing to
complications like obesity and insulin resistance (PMID: 32244780).
Taken together, these findings highlight that factors beyond REE and
typical correlates of this including BMI or fat-free mass, such as
hormonal and tissue-specific metabolic effects, may play a more
significant role in weight regulation challenges in women with PCOS.

As noted, women with PCOS are more likely to engage in weight-loss
attempts\textsuperscript{19} and there could be concerns that this could
inadvertently further foster the already well documented disordered
eating in this population\textsuperscript{21,22,63,64}. The pathways
linking PCOS and disordered eating are multifactorial. Biological
mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered
ghrelin and leptin signalling can heighten hunger, carbohydrate
cravings, and appetite variability\textsuperscript{65}. Frequent
hypoglycaemia and associated mood changes have also been observed, which
can trigger compensatory eating or binge episodes\textsuperscript{65}.
These physiological processes interact with psychological and social
stressors, including infertility concerns, conflicting nutrition advice,
chronic dieting, and exposure to idealised body images on social media,
which together compound vulnerability to disordered
eating\textsuperscript{66}. Moreover, eating disorders themselves can
disrupt endocrine function, potentially worsening PCOS symptoms and
creating a self-reinforcing cycle\textsuperscript{64,67}.

Understanding the intertwined biological, psychological, and social
influences suggests the importance of considering whether restrictive
dietary advice is appropriate given it may exacerbate feelings of
failure, hunger dysregulation, and shame\textsuperscript{68}. These
concerns, coupled with the lack of difference in REE between women with
and without PCOS might suggest that energy restriction based dietary
interventions for weight-loss may be unnecessary. But, there is also
evidence supporting the effect of energy restricted dietary
interventions for improving PCOS symptoms\textsuperscript{23,24} and
they are recommended in international guidelines\textsuperscript{25}.
Encouragingly though, these guidelines also recognise weight stigma as a
determinant of health and call for its reduction across clinical and
public health settings. Evidently the greater prevalence of women with
PCOS falling into overweight and obese BMI categories compared to women
without PCOS\textsuperscript{8,9} is unlikely to be due to differences
in REE and so, despite the potential effectiveness of energy restriction
dietary interventions, there is potential value in moving towards more
weight-neutral, individualised, and empowering care following holistic
guidelines recommending multiple approaches to
management\textsuperscript{20,25}.

\subsection{Strengths and Limitations}\label{strengths-and-limitations}

The current study has multiple strengths stemming from its
preregistered, comprehensive methodology and Bayesian statistical
approach. This statistical framework allowed us to incorporate studies
with and without control groups to better estimate REE in women with and
without PCOS and to perform multiple sensitivity analyses that confirmed
the stability of our findings. However, a limitation here is that
variability in methods across studies, such as differences in PCOS
diagnostic criteria and REE testing protocols, may have influenced
results and the relatively small number of studies overall limits the
extent to which we can explore these potential moderators. Furthermore,
some studies controlled for body weight or body composition when
reporting REE values, while others did not. We accounted for this by
estimating or converting reported data to obtain unadjusted REE values
across all groups, thereby reducing this variability and further as
noted above provided estimates adjusted for BMI and fat-free mass both
of which had little influence on our conclusions. Another limitation is
that, due to fewer total control groups than PCOS groups, informed
priors were required in several statistical models. However, in the
context of Bayesian meta-analysis this can also be considered a
strength. Additionally, some studies reported data inconsistencies that
could not be clarified (e.g.,\textsuperscript{13,35--37}) and were
dropped from our main analysis though our conclusions again did not
qualitative change when we conducted sensitivity analyses including
these studies. Finally, most included studies were cross-sectional or
baseline assessments within intervention trials, which limits causal
inference. Indeed, we did not pre-register any kind of causal model
(e.g., a directed acyclic graph) to inform our analysis approach for
causal inference and as such have been explicit about the estimates
presented as being descriptive.

\section{Conclusion}\label{conclusion}

In conclusion, the findings from this meta-analysis indicate that REE
does not meaningfully differ between women with and without PCOS.
Group-level differences in REE were small, insignificant, or not
physiologically relevant. Additionally, variability in REE between
individuals was also similar. These results suggest that a lower
baseline REE is not associated with the weight-related challenges often
associated with PCOS. These findings challenge the popular narrative
that women with PCOS have a lower REE and may help better inform dietary
interventions and nutritional support for these individuals. Future
research should include more standardized REE measurement and reporting
protocols, greater data transparency, consistent control and reporting
of body weight or body composition, the presentation of both absolute
and relative REE, and more precise characterization of PCOS phenotypes.
Overall, these findings support the conclusion that PCOS is not
negatively associated with REE and may help practitioners and
researchers focus on individually targeted and holistic lifestyle
interventions rather than negatively framed interventions based on
unsupported assumptions regarding REE.

\section{Data Availability}\label{data-availability}

All code utilised for data preparation, transformations, analyses,
plotting, and reporting are available in the corresponding GitHub
repository \url{https://github.com/jamessteeleii/pcos_ree_meta}.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-salariGlobalPrevalencePolycystic2024}
\CSLLeftMargin{1. }%
\CSLRightInline{Salari N, Nankali A, Ghanbari A, et al. Global
prevalence of polycystic ovary syndrome in women worldwide: A
comprehensive systematic review and meta-analysis. \emph{Archives of
Gynecology and Obstetrics}. 2024;310(3):1303-1314.
doi:\href{https://doi.org/10.1007/s00404-024-07607-x}{10.1007/s00404-024-07607-x}}

\bibitem[\citeproctext]{ref-glintborgCardiovascularDiseaseNationwide2018}
\CSLLeftMargin{2. }%
\CSLRightInline{Glintborg D, Rubin KH, Nybo M, Abrahamsen B, Andersen M.
Cardiovascular disease in a nationwide population of {Danish} women with
polycystic ovary syndrome. \emph{Cardiovascular Diabetology}.
2018;17(1):37.
doi:\href{https://doi.org/10.1186/s12933-018-0680-5}{10.1186/s12933-018-0680-5}}

\bibitem[\citeproctext]{ref-glintborgProspectiveRiskType2022}
\CSLLeftMargin{3. }%
\CSLRightInline{Glintborg D, Kolster ND, Ravn P, Andersen MS.
Prospective {Risk} of {Type} 2 {Diabetes} in {Normal Weight Women} with
{Polycystic Ovary Syndrome}. \emph{Biomedicines}. 2022;10(6):1455.
doi:\href{https://doi.org/10.3390/biomedicines10061455}{10.3390/biomedicines10061455}}

\bibitem[\citeproctext]{ref-perssonHigherRiskType2021}
\CSLLeftMargin{4. }%
\CSLRightInline{Persson S, Elenis E, Turkmen S, Kramer MS, Yong EL,
Poromaa IS. Higher risk of type 2 diabetes in women with hyperandrogenic
polycystic ovary syndrome. \emph{Fertility and Sterility}.
2021;116(3):862-871.
doi:\href{https://doi.org/10.1016/j.fertnstert.2021.04.018}{10.1016/j.fertnstert.2021.04.018}}

\bibitem[\citeproctext]{ref-glintborgProspectiveRiskType2024}
\CSLLeftMargin{5. }%
\CSLRightInline{Glintborg D, Ollila MM, Møller JJK, et al. Prospective
risk of {Type} 2 diabetes in 99~892 {Nordic} women with polycystic ovary
syndrome and 446~055 controls: National cohort study from {Denmark},
{Finland}, and {Sweden}. \emph{Human Reproduction (Oxford, England)}.
2024;39(8):1823-1834.
doi:\href{https://doi.org/10.1093/humrep/deae124}{10.1093/humrep/deae124}}

\bibitem[\citeproctext]{ref-limMetabolicSyndromePolycystic2019}
\CSLLeftMargin{6. }%
\CSLRightInline{Lim SS, Kakoly NS, Tan JWJ, et al. Metabolic syndrome in
polycystic ovary syndrome: A systematic review, meta-analysis and
meta-regression. \emph{Obesity Reviews: An Official Journal of the
International Association for the Study of Obesity}. 2019;20(2):339-352.
doi:\href{https://doi.org/10.1111/obr.12762}{10.1111/obr.12762}}

\bibitem[\citeproctext]{ref-johamPrevalenceInfertilityUse2015}
\CSLLeftMargin{7. }%
\CSLRightInline{Joham AE, Teede HJ, Ranasinha S, Zoungas S, Boyle J.
Prevalence of infertility and use of fertility treatment in women with
polycystic ovary syndrome: Data from a large community-based cohort
study. \emph{Journal of Women's Health (2002)}. 2015;24(4):299-307.
doi:\href{https://doi.org/10.1089/jwh.2014.5000}{10.1089/jwh.2014.5000}}

\bibitem[\citeproctext]{ref-barberWhyAreWomen2022}
\CSLLeftMargin{8. }%
\CSLRightInline{Barber TM. Why are women with polycystic ovary syndrome
obese? \emph{British Medical Bulletin}. 2022;143(1):4-15.
doi:\href{https://doi.org/10.1093/bmb/ldac007}{10.1093/bmb/ldac007}}

\bibitem[\citeproctext]{ref-barberObesityPolycysticOvary2021}
\CSLLeftMargin{9. }%
\CSLRightInline{Barber TM, Franks S. Obesity and polycystic ovary
syndrome. \emph{Clinical Endocrinology}. 2021;95(4):531-541.
doi:\href{https://doi.org/10.1111/cen.14421}{10.1111/cen.14421}}

\bibitem[\citeproctext]{ref-hirschbergImpairedCholecystokininSecretion2004}
\CSLLeftMargin{10. }%
\CSLRightInline{Hirschberg AL, Naessén S, Stridsberg M, Byström B,
Holtet J. Impaired cholecystokinin secretion and disturbed appetite
regulation in women with polycystic ovary syndrome. \emph{Gynecological
Endocrinology: The Official Journal of the International Society of
Gynecological Endocrinology}. 2004;19(2):79-87.
doi:\href{https://doi.org/10.1080/09513590400002300}{10.1080/09513590400002300}}

\bibitem[\citeproctext]{ref-moranGhrelinMeasuresSatiety2004}
\CSLLeftMargin{11. }%
\CSLRightInline{Moran LJ, Noakes M, Clifton PM, et al. Ghrelin and
measures of satiety are altered in polycystic ovary syndrome but not
differentially affected by diet composition. \emph{The Journal of
Clinical Endocrinology and Metabolism}. 2004;89(7):3337-3344.
doi:\href{https://doi.org/10.1210/jc.2003-031583}{10.1210/jc.2003-031583}}

\bibitem[\citeproctext]{ref-stefanakiFoodCravingsObesity2024}
\CSLLeftMargin{12. }%
\CSLRightInline{Stefanaki K, Karagiannakis DS, Peppa M, et al. Food
{Cravings} and {Obesity} in {Women} with {Polycystic Ovary Syndrome}:
{Pathophysiological} and {Therapeutic Considerations}. \emph{Nutrients}.
2024;16(7):1049.
doi:\href{https://doi.org/10.3390/nu16071049}{10.3390/nu16071049}}

\bibitem[\citeproctext]{ref-georgopoulosBasalMetabolicRate2009}
\CSLLeftMargin{13. }%
\CSLRightInline{Georgopoulos NA, Saltamavros AD, Vervita V, et al. Basal
metabolic rate is decreased in women with polycystic ovary syndrome and
biochemical hyperandrogenemia and is associated with insulin resistance.
\emph{Fertility and Sterility}. 2009;92(1):250-255.
doi:\href{https://doi.org/10.1016/j.fertnstert.2008.04.067}{10.1016/j.fertnstert.2008.04.067}}

\bibitem[\citeproctext]{ref-romualdiRestingMetabolicRate2019}
\CSLLeftMargin{14. }%
\CSLRightInline{Romualdi D, Versace V, Tagliaferri V, et al. The resting
metabolic rate in women with polycystic ovary syndrome and its relation
to the hormonal milieu, insulin metabolism, and body fat distribution: A
cohort study. \emph{Journal of Endocrinological Investigation}.
2019;42(9):1089-1097.
doi:\href{https://doi.org/10.1007/s40618-019-01029-2}{10.1007/s40618-019-01029-2}}

\bibitem[\citeproctext]{ref-churchillBasalMetabolicRate2015}
\CSLLeftMargin{15. }%
\CSLRightInline{Churchill SJ, Wang ET, Bhasin G, et al. Basal metabolic
rate in women with {PCOS} compared to eumenorrheic controls.
\emph{Clinical Endocrinology}. 2015;83(3):384-388.
doi:\href{https://doi.org/10.1111/cen.12740}{10.1111/cen.12740}}

\bibitem[\citeproctext]{ref-graffSaturatedFatIntake2017}
\CSLLeftMargin{16. }%
\CSLRightInline{Graff SK, Mario FM, Magalhães JA, Moraes RS, Spritzer
PM. Saturated {Fat Intake Is Related} to {Heart Rate Variability} in
{Women} with {Polycystic Ovary~Syndrome}. \emph{Annals of Nutrition \&
Metabolism}. 2017;71(3-4):224-233.
doi:\href{https://doi.org/10.1159/000484325}{10.1159/000484325}}

\bibitem[\citeproctext]{ref-larssonDietaryIntakeResting2016}
\CSLLeftMargin{17. }%
\CSLRightInline{Larsson I, Hulthén L, Landén M, Pålsson E, Janson P,
Stener-Victorin E. Dietary intake, resting energy expenditure, and
eating behavior in women with and without polycystic ovary syndrome.
\emph{Clinical Nutrition (Edinburgh, Scotland)}. 2016;35(1):213-218.
doi:\href{https://doi.org/10.1016/j.clnu.2015.02.006}{10.1016/j.clnu.2015.02.006}}

\bibitem[\citeproctext]{ref-segalRestingMetabolicRate1990}
\CSLLeftMargin{18. }%
\CSLRightInline{Segal KR, Dunaif A.
\href{https://www.ncbi.nlm.nih.gov/pubmed/2228390}{Resting metabolic
rate and postprandial thermogenesis in polycystic ovarian syndrome}.
\emph{International Journal of Obesity}. 1990;14(7):559-567.}

\bibitem[\citeproctext]{ref-pesonenPolycysticOvarySyndrome2023}
\CSLLeftMargin{19. }%
\CSLRightInline{Pesonen E, Nurkkala M, Niemelä M, et al. Polycystic
ovary syndrome is associated with weight-loss attempts and perception of
overweight independent of {BMI}: A population-based cohort study.
\emph{Obesity (Silver Spring, Md)}. 2023;31(4):1108-1120.
doi:\href{https://doi.org/10.1002/oby.23681}{10.1002/oby.23681}}

\bibitem[\citeproctext]{ref-ozgensaydamWeightManagementStrategies2021}
\CSLLeftMargin{20. }%
\CSLRightInline{Ozgen Saydam B, Yildiz BO. Weight management strategies
for patients with {PCOS}: Current perspectives. \emph{Expert Review of
Endocrinology \& Metabolism}. 2021;16(2):49-62.
doi:\href{https://doi.org/10.1080/17446651.2021.1896966}{10.1080/17446651.2021.1896966}}

\bibitem[\citeproctext]{ref-lalonde-besterPrevalenceEtiologyEating2024}
\CSLLeftMargin{21. }%
\CSLRightInline{Lalonde-Bester S, Malik M, Masoumi R, et al. Prevalence
and {Etiology} of {Eating Disorders} in {Polycystic Ovary Syndrome}: {A
Scoping Review}. \emph{Advances in Nutrition (Bethesda, Md)}.
2024;15(4):100193.
doi:\href{https://doi.org/10.1016/j.advnut.2024.100193}{10.1016/j.advnut.2024.100193}}

\bibitem[\citeproctext]{ref-jeanesBingeEatingBehaviours2017}
\CSLLeftMargin{22. }%
\CSLRightInline{Jeanes YM, Reeves S, Gibson EL, Piggott C, May VA, Hart
KH. Binge eating behaviours and food cravings in women with {Polycystic
Ovary Syndrome}. \emph{Appetite}. 2017;109:24-32.
doi:\href{https://doi.org/10.1016/j.appet.2016.11.010}{10.1016/j.appet.2016.11.010}}

\bibitem[\citeproctext]{ref-aleneziImpactDietInducedWeight2024}
\CSLLeftMargin{23. }%
\CSLRightInline{Alenezi SA, Elkmeshi N, Alanazi A, Alanazi ST, Khan R,
Amer S. The {Impact} of {Diet-Induced Weight Loss} on {Inflammatory
Status} and {Hyperandrogenism} in {Women} with {Polycystic Ovarian
Syndrome} ({PCOS})-{A Systematic Review} and {Meta-Analysis}.
\emph{Journal of Clinical Medicine}. 2024;13(16):4934.
doi:\href{https://doi.org/10.3390/jcm13164934}{10.3390/jcm13164934}}

\bibitem[\citeproctext]{ref-holteRestoredInsulinSensitivity1995}
\CSLLeftMargin{24. }%
\CSLRightInline{Holte J, Bergh T, Berne C, Wide L, Lithell H. Restored
insulin sensitivity but persistently increased early insulin secretion
after weight loss in obese women with polycystic ovary syndrome.
\emph{The Journal of Clinical Endocrinology and Metabolism}.
1995;80(9):2586-2593.
doi:\href{https://doi.org/10.1210/jcem.80.9.7673399}{10.1210/jcem.80.9.7673399}}

\bibitem[\citeproctext]{ref-teedeRecommendations2023International2023}
\CSLLeftMargin{25. }%
\CSLRightInline{Teede HJ, Tay CT, Laven JJE, et al. Recommendations
{From} the 2023 {International Evidence-based Guideline} for the
{Assessment} and {Management} of {Polycystic Ovary Syndrome}. \emph{The
Journal of Clinical Endocrinology and Metabolism}.
2023;108(10):2447-2469.
doi:\href{https://doi.org/10.1210/clinem/dgad463}{10.1210/clinem/dgad463}}

\bibitem[\citeproctext]{ref-himeleinDepressionBodyImage2006}
\CSLLeftMargin{26. }%
\CSLRightInline{Himelein MJ, Thatcher SS. Depression and body image
among women with polycystic ovary syndrome. \emph{Journal of Health
Psychology}. 2006;11(4):613-625.
doi:\href{https://doi.org/10.1177/1359105306065021}{10.1177/1359105306065021}}

\bibitem[\citeproctext]{ref-hofmannBodyImageMental2025}
\CSLLeftMargin{27. }%
\CSLRightInline{Hofmann K, Decrinis C, Bitterlich N, Tropschuh K, Stute
P, Bachmann A. Body image and mental health in women with polycystic
ovary syndrome-a cross-sectional study. \emph{Archives of Gynecology and
Obstetrics}. 2025;312(1):177-190.
doi:\href{https://doi.org/10.1007/s00404-024-07913-4}{10.1007/s00404-024-07913-4}}

\bibitem[\citeproctext]{ref-gellerBodyImageIllness2025}
\CSLLeftMargin{28. }%
\CSLRightInline{Geller S, Levy S, Avitsur R. Body image, illness
perception, and psychological distress in women coping with polycystic
ovary syndrome ({PCOS}). \emph{Health Psychology Open}.
2025;12:20551029251327441.
doi:\href{https://doi.org/10.1177/20551029251327441}{10.1177/20551029251327441}}

\bibitem[\citeproctext]{ref-davitadzeBodyImageConcerns2023}
\CSLLeftMargin{29. }%
\CSLRightInline{Davitadze M, Malhotra K, Khalil H, et al. Body image
concerns in women with polycystic ovary syndrome: A systematic review
and meta-analysis. \emph{European Journal of Endocrinology}.
2023;189(2):R1-R9.
doi:\href{https://doi.org/10.1093/ejendo/lvad110}{10.1093/ejendo/lvad110}}

\bibitem[\citeproctext]{ref-pagePRISMA2020Statement2021a}
\CSLLeftMargin{30. }%
\CSLRightInline{Page MJ, McKenzie JE, Bossuyt PM, et al. The {PRISMA}
2020 statement: An updated guideline for reporting systematic reviews.
\emph{BMJ}. 2021;372:n71.
doi:\href{https://doi.org/10.1136/bmj.n71}{10.1136/bmj.n71}}

\bibitem[\citeproctext]{ref-haddawayPRISMA2020PackageShiny2022}
\CSLLeftMargin{31. }%
\CSLRightInline{Haddaway NR, Page MJ, Pritchard CC, McGuinness LA.
{PRISMA2020}: {An R} package and {Shiny} app for producing {PRISMA}
2020-compliant flow diagrams, with interactivity for optimised digital
transparency and {Open Synthesis}. \emph{Campbell Systematic Reviews}.
2022;18(2):e1230.
doi:\href{https://doi.org/10.1002/cl2.1230}{10.1002/cl2.1230}}

\bibitem[\citeproctext]{ref-weberApplyingMetaAnalyticPredictivePriors2021}
\CSLLeftMargin{32. }%
\CSLRightInline{Weber S, Li Y, Iii JWS, Kakizume T, Schmidli H. Applying
{Meta-Analytic-Predictive Priors} with the {R Bayesian Evidence
Synthesis Tools}. \emph{Journal of Statistical Software}. 2021;100:1-32.
doi:\href{https://doi.org/10.18637/jss.v100.i19}{10.18637/jss.v100.i19}}

\bibitem[\citeproctext]{ref-wanEstimatingSampleMean2014}
\CSLLeftMargin{33. }%
\CSLRightInline{Wan X, Wang W, Liu J, Tong T. Estimating the sample mean
and standard deviation from the sample size, median, range and/or
interquartile range. \emph{BMC Medical Research Methodology}.
2014;14(1):135.
doi:\href{https://doi.org/10.1186/1471-2288-14-135}{10.1186/1471-2288-14-135}}

\bibitem[\citeproctext]{ref-pohlmeierEffectLowstarchLowdairy2014}
\CSLLeftMargin{34. }%
\CSLRightInline{Pohlmeier AM, Phy JL, Watkins P, et al. Effect of a
low-starch/low-dairy diet on fat oxidation in overweight and obese women
with polycystic ovary syndrome. \emph{Applied Physiology, Nutrition, and
Metabolism = Physiologie Appliquee, Nutrition Et Metabolisme}.
2014;39(11):1237-1244.
doi:\href{https://doi.org/10.1139/apnm-2014-0073}{10.1139/apnm-2014-0073}}

\bibitem[\citeproctext]{ref-kritikouA2BV3Adrenergic2006}
\CSLLeftMargin{35. }%
\CSLRightInline{Kritikou S, Saltamavros AD, Adonakis G, et al. The
{\(\alpha\)2B} and {\(\beta\)}3 adrenergic receptor genes polymorphisms
in women with polycystic ovarian syndrome ({PCOS}) and their association
with insulin resistance and basal metabolic rate ({BMR}). \emph{Review
of Clinical Pharmacology and Pharmacokinetics -- International Edition}.
2006;2006(2).}

\bibitem[\citeproctext]{ref-saltamavrosAlpha2Beta2007}
\CSLLeftMargin{36. }%
\CSLRightInline{Saltamavros AD, Adonakis G, Kritikou S, et al. Alpha 2
beta adrenoreceptor 301-303 deletion polymorphism in polycystic ovary
syndrome. \emph{Clinical Autonomic Research: Official Journal of the
Clinical Autonomic Research Society}. 2007;17(2):112-114.
doi:\href{https://doi.org/10.1007/s10286-007-0403-6}{10.1007/s10286-007-0403-6}}

\bibitem[\citeproctext]{ref-koikaAssociationPro12AlaPolymorphism2009}
\CSLLeftMargin{37. }%
\CSLRightInline{Koika V, Marioli DJ, Saltamavros AD, et al. Association
of the {Pro12Ala} polymorphism in peroxisome proliferator-activated
receptor Gamma2 with decreased basic metabolic rate in women with
polycystic ovary syndrome. \emph{European Journal of Endocrinology}.
2009;161(2):317-322.
doi:\href{https://doi.org/10.1530/EJE-08-1014}{10.1530/EJE-08-1014}}

\bibitem[\citeproctext]{ref-rodriguez-sanchezGratefulFacilitateCitation2023}
\CSLLeftMargin{38. }%
\CSLRightInline{Rodriguez-Sanchez F, cre, cph, Jackson CP, Hutchins SD,
Clawson JM. Grateful: {Facilitate Citation} of {R Packages}. Published
online October 2023.}

\bibitem[\citeproctext]{ref-cummingNewStatisticsWhy2014}
\CSLLeftMargin{39. }%
\CSLRightInline{Cumming G. The {New Statistics}: {Why} and {How}.
\emph{Psychological Science}. 2014;25(1):7-29.
doi:\href{https://doi.org/10.1177/0956797613504966}{10.1177/0956797613504966}}

\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
\CSLLeftMargin{40. }%
\CSLRightInline{Kruschke JK, Liddell TM. The {Bayesian New Statistics}:
{Hypothesis} testing, estimation, meta-analysis, and power analysis from
a {Bayesian} perspective. \emph{Psychonomic Bulletin \& Review}.
2018;25:178-206.}

\bibitem[\citeproctext]{ref-usheyRenvProjectEnvironments2023}
\CSLLeftMargin{41. }%
\CSLRightInline{Ushey K, cre, Wickham H, Software P, PBC. Renv: {Project
Environments}. Published online September 2023.}

\bibitem[\citeproctext]{ref-landauTargetsDynamicFunctionOriented2023}
\CSLLeftMargin{42. }%
\CSLRightInline{Landau WM, Warkentin MT, Edmondson M, Oliver S, Mahr T,
Company EL and. Targets: {Dynamic Function-Oriented} '{Make}'-{Like
Declarative Pipelines}. Published online October 2023.}

\bibitem[\citeproctext]{ref-viechtbauerMetaforMetaAnalysisPackage2023}
\CSLLeftMargin{43. }%
\CSLRightInline{Viechtbauer W. Metafor: {Meta-Analysis Package} for {R}.
Published online September 2023.}

\bibitem[\citeproctext]{ref-burknerBrmsBayesianRegression2023}
\CSLLeftMargin{44. }%
\CSLRightInline{Bürkner PC, Gabry J, Weber S, et al. Brms: {Bayesian
Regression Models} using '{Stan}'. Published online September 2023.}

\bibitem[\citeproctext]{ref-arel-bundockMarginaleffectsPredictionsComparisons2023}
\CSLLeftMargin{45. }%
\CSLRightInline{Arel-Bundock V, cre, cph, Diniz MA, Greifer N, Bacher E.
Marginaleffects: {Predictions}, {Comparisons}, {Slopes}, {Marginal
Means}, and {Hypothesis Tests}. Published online October 2023.}

\bibitem[\citeproctext]{ref-kayTidybayesTidyData2023}
\CSLLeftMargin{46. }%
\CSLRightInline{Kay M, Mastny T. Tidybayes: {Tidy Data} and '{Geoms}'
for {Bayesian Models}. Published online August 2023.}

\bibitem[\citeproctext]{ref-wickhamGgplot2CreateElegant2023}
\CSLLeftMargin{47. }%
\CSLRightInline{Wickham H, Chang W, Henry L, et al. Ggplot2: {Create
Elegant Data Visualisations Using} the {Grammar} of {Graphics}.
Published online October 2023.}

\bibitem[\citeproctext]{ref-pedersenPatchworkComposerPlots2023}
\CSLLeftMargin{48. }%
\CSLRightInline{Pedersen TL. Patchwork: {The Composer} of {Plots}.
Published online August 2023.}

\bibitem[\citeproctext]{ref-hongBayesianMissingData2016}
\CSLLeftMargin{49. }%
\CSLRightInline{Hong H, Chu H, Zhang J, Carlin BP. A {Bayesian} missing
data framework for generalized multiple outcome mixed treatment
comparisons. \emph{Research Synthesis Methods}. 2016;7(1):6-22.
doi:\href{https://doi.org/10.1002/jrsm.1153}{10.1002/jrsm.1153}}

\bibitem[\citeproctext]{ref-velasquezUseAmmoniaInhalants2011}
\CSLLeftMargin{50. }%
\CSLRightInline{Velasquez JR. The {Use} of {Ammonia Inhalants Among
Athletes}. \emph{Strength \& Conditioning Journal}. 2011;33(2):33.
doi:\href{https://doi.org/10.1519/SSC.0b013e3181fd5c9b}{10.1519/SSC.0b013e3181fd5c9b}}

\bibitem[\citeproctext]{ref-pavlidouRevisedHarrisBenedict2023}
\CSLLeftMargin{51. }%
\CSLRightInline{Pavlidou E, Papadopoulou SK, Seroglou K, Giaginis C.
Revised {Harris}--{Benedict Equation}: {New Human Resting Metabolic Rate
Equation}. \emph{Metabolites}. 2023;13(2):189.
doi:\href{https://doi.org/10.3390/metabo13020189}{10.3390/metabo13020189}}

\bibitem[\citeproctext]{ref-steeleMetaanalysisVariationSport2023a}
\CSLLeftMargin{52. }%
\CSLRightInline{Steele J, Fisher, Smith, Schoenfeld, Yang, and Nakagawa
S. Meta-analysis of variation in sport and exercise science: {Examples}
of application within resistance training research. \emph{Journal of
Sports Sciences}. 2023;41(17):1617-1634.
doi:\href{https://doi.org/10.1080/02640414.2023.2286748}{10.1080/02640414.2023.2286748}}

\bibitem[\citeproctext]{ref-moranShorttermMealReplacements2006}
\CSLLeftMargin{53. }%
\CSLRightInline{Moran LJ, Noakes M, Clifton PM, Wittert GA, Williams G,
Norman RJ. Short-term meal replacements followed by dietary
macronutrient restriction enhance weight loss in polycystic ovary
syndrome. \emph{The American Journal of Clinical Nutrition}.
2006;84(1):77-87.
doi:\href{https://doi.org/10.1093/ajcn/84.1.77}{10.1093/ajcn/84.1.77}}

\bibitem[\citeproctext]{ref-brunerEffectsExerciseNutritional2006}
\CSLLeftMargin{54. }%
\CSLRightInline{Bruner B, Chad K, Chizen D. Effects of exercise and
nutritional counseling in women with polycystic ovary syndrome.
\emph{Applied Physiology, Nutrition, and Metabolism = Physiologie
Appliquee, Nutrition Et Metabolisme}. 2006;31(4):384-391.
doi:\href{https://doi.org/10.1139/h06-007}{10.1139/h06-007}}

\bibitem[\citeproctext]{ref-robinsonPostprandialThermogenesisReduced1992}
\CSLLeftMargin{55. }%
\CSLRightInline{Robinson S, Chan SP, Spacey S, Anyaoku V, Johnston DG,
Franks S. Postprandial thermogenesis is reduced in polycystic ovary
syndrome and is associated with increased insulin resistance.
\emph{Clinical Endocrinology}. 1992;36(6):537-543.
doi:\href{https://doi.org/10.1111/j.1365-2265.1992.tb02262.x}{10.1111/j.1365-2265.1992.tb02262.x}}

\bibitem[\citeproctext]{ref-cosarRestingMetabolicRate2008}
\CSLLeftMargin{56. }%
\CSLRightInline{Cosar E, Köken G, Sahin FK, et al. Resting metabolic
rate and exercise capacity in women with polycystic ovary syndrome.
\emph{International Journal of Gynaecology and Obstetrics: The Official
Organ of the International Federation of Gynaecology and Obstetrics}.
2008;101(1):31-34.
doi:\href{https://doi.org/10.1016/j.ijgo.2007.10.011}{10.1016/j.ijgo.2007.10.011}}

\bibitem[\citeproctext]{ref-graffDietaryGlycemicIndex2013}
\CSLLeftMargin{57. }%
\CSLRightInline{Graff SK, Mário FM, Alves BC, Spritzer PM. Dietary
glycemic index is associated with less favorable anthropometric and
metabolic profiles in polycystic ovary syndrome women with different
phenotypes. \emph{Fertility and Sterility}. 2013;100(4):1081-1088.
doi:\href{https://doi.org/10.1016/j.fertnstert.2013.06.005}{10.1016/j.fertnstert.2013.06.005}}

\bibitem[\citeproctext]{ref-dohRelationshipAdiposityInsulin2016}
\CSLLeftMargin{58. }%
\CSLRightInline{Doh E, Mbanya A, Kemfang-Ngowa JD, et al. The
{Relationship} between {Adiposity} and {Insulin Sensitivity} in {African
Women Living} with the {Polycystic Ovarian Syndrome}: {A Clamp Study}.
\emph{International Journal of Endocrinology}. 2016;2016:9201701.
doi:\href{https://doi.org/10.1155/2016/9201701}{10.1155/2016/9201701}}

\bibitem[\citeproctext]{ref-rodriguesLowValidityPredictive2018}
\CSLLeftMargin{59. }%
\CSLRightInline{Rodrigues AM dos S, Costa ABP, Campos DL, et al. Low
validity of predictive equations for calculating resting energy
expenditure in overweight and obese women with polycystic ovary
syndrome. \emph{Journal of Human Nutrition and Dietetics}.
2018;31(2):266-275.
doi:\href{https://doi.org/10.1111/jhn.12498}{10.1111/jhn.12498}}

\bibitem[\citeproctext]{ref-tosiInsulinMediatedSubstrateUse2021}
\CSLLeftMargin{60. }%
\CSLRightInline{Tosi F, Villani M, Migazzi M, et al. Insulin-{Mediated
Substrate Use} in {Women With Different Phenotypes} of {PCOS}: The
{Role} of {Androgens}. \emph{The Journal of Clinical Endocrinology and
Metabolism}. 2021;106(9):e3414-e3425.
doi:\href{https://doi.org/10.1210/clinem/dgab380}{10.1210/clinem/dgab380}}

\bibitem[\citeproctext]{ref-broskeyAssessingEnergyRequirements2017}
\CSLLeftMargin{61. }%
\CSLRightInline{Broskey NT, Klempel MC, Gilmore LA, et al. Assessing
{Energy Requirements} in {Women With Polycystic Ovary Syndrome}: {A
Comparison Against Doubly Labeled Water}. \emph{The Journal of Clinical
Endocrinology and Metabolism}. 2017;102(6):1951-1959.
doi:\href{https://doi.org/10.1210/jc.2017-00459}{10.1210/jc.2017-00459}}

\bibitem[\citeproctext]{ref-tosiRestingEnergyExpenditure2024}
\CSLLeftMargin{62. }%
\CSLRightInline{Tosi F, Rosmini F, Gremes V, et al. Resting energy
expenditure in women with polycystic ovary syndrome. \emph{Human
Reproduction (Oxford, England)}. 2024;39(8):1794-1803.
doi:\href{https://doi.org/10.1093/humrep/deae129}{10.1093/humrep/deae129}}

\bibitem[\citeproctext]{ref-leeIncreasedRiskDisordered2017}
\CSLLeftMargin{63. }%
\CSLRightInline{Lee I, Cooney LG, Saini S, et al. Increased risk of
disordered eating in~polycystic ovary syndrome. \emph{Fertility and
Sterility}. 2017;107(3):796-802.
doi:\href{https://doi.org/10.1016/j.fertnstert.2016.12.014}{10.1016/j.fertnstert.2016.12.014}}

\bibitem[\citeproctext]{ref-bernadettPrevalenceEatingDisorders2016}
\CSLLeftMargin{64. }%
\CSLRightInline{Bernadett M, Szemán-N A.
\href{https://www.ncbi.nlm.nih.gov/pubmed/27244869}{{{[}Prevalence of
eating disorders among women with polycystic ovary syndrome{]}}}.
\emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos
Folyoirata}. 2016;31(2):136-145.}

\bibitem[\citeproctext]{ref-barryImpactEatingBehavior2011}
\CSLLeftMargin{65. }%
\CSLRightInline{Barry JA, Bouloux P, Hardiman PJ. The impact of eating
behavior on psychological symptoms typical of reactive hypoglycemia. {A}
pilot study comparing women with polycystic ovary syndrome to controls.
\emph{Appetite}. 2011;57(1):73-76.
doi:\href{https://doi.org/10.1016/j.appet.2011.03.003}{10.1016/j.appet.2011.03.003}}

\bibitem[\citeproctext]{ref-steegers-theunissenPolycysticOvarySyndrome2020}
\CSLLeftMargin{66. }%
\CSLRightInline{Steegers-Theunissen RPM, Wiegel RE, Jansen PW, Laven
JSE, Sinclair KD. Polycystic {Ovary Syndrome}: {A Brain Disorder
Characterized} by {Eating Problems Originating} during {Puberty} and
{Adolescence}. \emph{International Journal of Molecular Sciences}.
2020;21(21):8211.
doi:\href{https://doi.org/10.3390/ijms21218211}{10.3390/ijms21218211}}

\bibitem[\citeproctext]{ref-cooneyIncreasedPrevalenceBinge2024}
\CSLLeftMargin{67. }%
\CSLRightInline{Cooney LG, Gyorfi K, Sanneh A, et al. Increased
{Prevalence} of {Binge Eating Disorder} and {Bulimia Nervosa} in {Women
With Polycystic Ovary Syndrome}: {A Systematic Review} and
{Meta-Analysis}. \emph{The Journal of Clinical Endocrinology and
Metabolism}. 2024;109(12):3293-3305.
doi:\href{https://doi.org/10.1210/clinem/dgae462}{10.1210/clinem/dgae462}}

\bibitem[\citeproctext]{ref-pirottaDisorderedEatingBehaviours2019}
\CSLLeftMargin{68. }%
\CSLRightInline{Pirotta S, Barillaro M, Brennan L, et al. Disordered
{Eating Behaviours} and {Eating Disorders} in {Women} in {Australia}
with and without {Polycystic Ovary Syndrome}: {A Cross-Sectional Study}.
\emph{Journal of Clinical Medicine}. 2019;8(10):1682.
doi:\href{https://doi.org/10.3390/jcm8101682}{10.3390/jcm8101682}}

\end{CSLReferences}

\end{document}

### 2025-12-01 manuscript/pre_print.qmd
JS and HC COIs/Financial disclosures

@@
 
+# Financial Disclosures/Conflicts of Interest
+Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.
+James Steele provides research consultancy through his company Steele Research Limited, is contracted currently by MacroFactor and Kieser Australia through Steele Research Limited, and has also received travel expenses and honoraria for speaking from fit20 International, Exercise School Portugal, and Discover Strength.
+
+The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health or any other funding agencies or institutions noted above.
+
 # Data Availability

### 2025-12-01 manuscript/pre_print.tex
JS and HC COIs/Financial disclosures

@@
 
+\section{Financial Disclosures/Conflicts of
+Interest}\label{financial-disclosuresconflicts-of-interest}
+
+Hannah Cabré was supported by the National Institute of Diabetes and
+Digestive and Kidney Diseases of the National Institutes of Health under
+Award Number T32DK064584. James Steele provides research consultancy
+through his company Steele Research Limited, is contracted currently by
+MacroFactor and Kieser Australia through Steele Research Limited, and
+has also received travel expenses and honoraria for speaking from fit20
+International, Exercise School Portugal, and Discover Strength.
+
+The content is solely the responsibility of the authors and does not
+necessarily represent the official views of the National Institutes of
+Health or any other funding agencies or institutions noted above.
+
 \section{Data Availability}\label{data-availability}

### 2025-12-01 manuscript/pre_print.qmd
update missing PMIDs

@@
 
-These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain.
+These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism [@@teedeRecommendations2023International2023], predisposing women with PCOS to weight gain.
 This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
@@
 As such, even adjusted for both BMI and fat-free mass, there seems to be little difference in REE between women with, and without, PCOS.
-Yet, a recent systematic review of mechanisms for metabolic dysfunction has reported excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780).
+Yet, a recent systematic review of mechanisms for metabolic dysfunction has reported excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance [@sanchez-garridoMetabolicDysfunctionPolycystic2020].
 Taken together, these findings highlight that factors beyond REE and typical correlates of this including BMI or fat-free mass, such as hormonal and tissue-specific metabolic effects, may play a more significant role in weight regulation challenges in women with PCOS.

### 2025-12-01 manuscript/submission_manuscript.qmd
added author affiliation front matter and corresponding plus citation
update missing PMIDs

@@
         \usepackage{pdflscape}
+        \usepackage[noblocks]{authblk}
+        \renewcommand*{\Authsep}{, }
+        \renewcommand*{\Authand}{, }
+        \renewcommand*{\Authands}{, }
+        \renewcommand\Affilfont{\small}
 geometry:
@@
     orcid: 0000-0003-4645-0077
-    corresponding: true
     email: R.P.Kirwan@ljmu.ac.uk
@@
       - ref: 3
-    # orcid: 0000-0002-8003-0757
     email: g_kohlhoff@hotmail.co.uk
@@
     orcid: 0000-0002-2217-8067
-    corresponding: true
     email: alyssaolenickphd@gmail.com

### 2025-12-02 manuscript/pre_print.qmd
Leigh's edit to discussion and COIs

@@
   - name: James Steele
-    affil-id: 6,2,7
+    affil-id: 2,6,7
     orcid: 0000-0002-8003-0757
@@
 Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis.
-medrχiv DOI: TO ADD.
+medrxiv DOI: TO ADD.
 Address for correspondence: james@steele-research.com
@@
 These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism [@@teedeRecommendations2023International2023], predisposing women with PCOS to weight gain.
-This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
+This belief largely stems from a single but influential 2009 study from Georgopoulos et al.
+that reported a significantly reduced BMR in women with PCOS [@georgopoulosBasalMetabolicRate2009], which has been widely cited and reinforced in both academic and clinical contexts.
+However, as we have noted, this study along with others [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009] have numerous reporting errors which led us to drop them from our present analysis (though sensitivity analysis including them did not alter our overall conclusions).
+The mistaken belief that REE is lower for those with PCOS may have mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
 Recognising that there may be minimal differences in REE between women with PCOS and those without can inform both clinical and public practices, potentially leading to a shift in focus away from a requirement for more severe caloric restriction as a primary method of treatment towards more comprehensive, individualised, and psychologically safe approaches to care [@teedeRecommendations2023International2023].
@@
 # Financial Disclosures/Conflicts of Interest
+Greg Nuckols and Leigh Peele are employed by MacroFactor, a nutrition-tracking app.
 Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.
+Alyssa Olenick provides fitness coaching and scientific consulting under ALYSSA OLENICK LLC.
+She receives compensation from scientific consulting, speaking engagments, affiliated social media partnerships, and book publishing.
 James Steele provides research consultancy through his company Steele Research Limited, is contracted currently by MacroFactor and Kieser Australia through Steele Research Limited, and has also received travel expenses and honoraria for speaking from fit20 International, Exercise School Portugal, and Discover Strength.
@@
 The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health or any other funding agencies or institutions noted above.
+The authors declare no other conflicts of interest related to the submitted work
 

### 2025-12-03 manuscript/pre_print.qmd
Greg COI/declaration

@@
     email: leigh@macrofactorapp.com
-  - name: Greg Nuckols
+  - name: Gregory Nuckols
     affil-id: 2
@@
   - id: 1
-    name: Research Institute for Sports and Exercise Sciences, Liverpool John Moores Universtiy, Liverpool, UK
+    name: Research Institute for Sports and Exercise Sciences, Liverpool John Moores University, Liverpool, UK
   - id: 2
-    name: MacroFactor, Stronger by Science LLC, Raleigh, North Carolina, USA
+    name: MacroFactor, Stronger by Science Technologies LLC, Raleigh, North Carolina, USA
   - id: 3
@@
 
-The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively).
+The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_mean_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_mean_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[1])` kcal/day], respectively).
 If anything, REE may be slightly higher in women with PCOS compared to healthy women without PCOS.
@@
 # Financial Disclosures/Conflicts of Interest
-Greg Nuckols and Leigh Peele are employed by MacroFactor, a nutrition-tracking app.
+Leigh Peele is employed by MacroFactor, a nutrition-tracking app.
+Gregory Nuckols is co-owner of MacroFactor.
 Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.

### 2025-12-03 manuscript/pre_print.qmd
Preprint ready for submission

@@
         \renewcommand\Affilfont{\small}
+        \usepackage{etoolbox}
+        \AtEndEnvironment{abstract}{\noindent\textbf{Keywords:} \KeywordList \par}
 geometry:
@@
 Address for correspondence: james@steele-research.com
-keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
+keywords: Polycystic ovary syndrome, Energy Metabolism, Resting Metabolic Rate, Basal Metabolism
 bibliography: 
@@
 A total of 11 studies used the Rotterdam criteria (or a modified version there of) for diagnosing PCOS [@robinsonPostprandialThermogenesisReduced1992; @brunerEffectsExerciseNutritional2006; @moranShorttermMealReplacements2006; @cosarRestingMetabolicRate2008; @graffDietaryGlycemicIndex2013; @pohlmeierEffectLowstarchLowdairy2014; @dohRelationshipAdiposityInsulin2016; @larssonDietaryIntakeResting2016; @graffSaturatedFatIntake2017; @rodriguesLowValidityPredictive2018; @tosiInsulinMediatedSubstrateUse2021].
-One study used the 1990 National Institutes of Health criteria [@broskeyAssessingEnergyRequirements2017], and two studies [@segalRestingMetabolicRate1990; @robinsonPostprandialThermogenesisReduced1992] diagnosed PCOS via the presence of oligomenorrhea or amenorrheaammenorrhea alongside additional criteria including plasma androgen levels, hirsutism or polycystic ovaries on ultrasound scanning.
+One study used the 1990 National Institutes of Health criteria [@broskeyAssessingEnergyRequirements2017], and two studies [@segalRestingMetabolicRate1990; @robinsonPostprandialThermogenesisReduced1992] diagnosed PCOS via the presence of oligomenorrhea or amenorrhea alongside additional criteria including plasma androgen levels, hirsutism or polycystic ovaries on ultrasound scanning.
 All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry.
@@
 # Financial Disclosures/Conflicts of Interest
+Richard Kirwan provides nutrition consultancy through his company Be More Nutrition.
 Leigh Peele is employed by MacroFactor, a nutrition-tracking app.
 Gregory Nuckols is co-owner of MacroFactor.
+Georgia Kohlhoff provides nutrition consultancy through her company Flourishing Health.
 Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.
@@
 The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health or any other funding agencies or institutions noted above.
-The authors declare no other conflicts of interest related to the submitted work
+The authors declare no other conflicts of interest related to the submitted work.
 

### 2025-12-03 manuscript/pre_print.tex
Preprint ready for submission

@@
 \renewcommand\Affilfont{\small}
+\usepackage{etoolbox}
+\AtEndEnvironment{abstract}{\noindent\textbf{Keywords:} \KeywordList \par}
 \makeatletter
@@
   pdftitle={Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis},
-  pdfauthor={Richard Kirwan; Leigh Peele; Greg Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
-  pdfkeywords={Polycystic ovary syndrome, PCOS, resting energy
-expenditure, REE, basal metabolic rate, BMR},
+  pdfauthor={Richard Kirwan; Leigh Peele; Gregory Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
+  pdfkeywords={Polycystic ovary syndrome, Energy Metabolism, Resting
+Metabolic Rate, Basal Metabolism},
   colorlinks=true,
@@
 women with and without polycystic ovary syndrome: a systematic review
-and meta-analysis. medrχiv DOI: TO ADD. Address for correspondence:
+and meta-analysis. medrxiv DOI: TO ADD. Address for correspondence:
 james@steele-research.com}}
@@
 \author[2]{Leigh Peele}
-\author[2]{Greg Nuckols}
+\author[2]{Gregory Nuckols}
 \author[3]{Georgia Kohlhoff}
@@
 \author[5]{Alyssa Olenick}
-\author[6,2,7]{James Steele}
+\author[2,6,7]{James Steele}
 
 \affil[1]{Research Institute for Sports and Exercise Sciences, Liverpool
-John Moores Universtiy, Liverpool, UK}
-\affil[2]{MacroFactor, Stronger by Science LLC, Raleigh, North Carolina,
-USA}
+John Moores University, Liverpool, UK}
+\affil[2]{MacroFactor, Stronger by Science Technologies LLC, Raleigh,
+North Carolina, USA}
 \affil[3]{Flourishing Health, Liverpool, UK}
@@
 
-\date{December 1, 2025}
+% Store the keywords (expanded by Pandoc here)
+\def\KeywordList{Polycystic ovary syndrome, Energy Metabolism, Resting
+Metabolic Rate, Basal Metabolism}
+
+\date{December 3, 2025}
 \begin{document}
@@
 studies\textsuperscript{18,55} diagnosed PCOS via the presence of
-oligomenorrhea or amenorrheaammenorrhea alongside additional criteria
-including plasma androgen levels, hirsutism or polycystic ovaries on
-ultrasound scanning. All but one study\textsuperscript{61}, which used
-doubly labelled water, measured REE using indirect calorimetry. The
-specific devices reported by these studies are included in the online
+oligomenorrhea or amenorrhea alongside additional criteria including
+plasma androgen levels, hirsutism or polycystic ovaries on ultrasound
+scanning. All but one study\textsuperscript{61}, which used doubly
+labelled water, measured REE using indirect calorimetry. The specific
+devices reported by these studies are included in the online
 supplementary materials (see
@@
 These findings challenge the widely held belief that PCOS is inherently
-associated with a slower metabolism (PMID: 32244780), predisposing women
-with PCOS to weight gain. This belief may mistakenly lead to
+associated with a slower metabolism\textsuperscript{@ 25}, predisposing
+women with PCOS to weight gain. This belief largely stems from a single
+but influential 2009 study from Georgopoulos et al.~that reported a
+significantly reduced BMR in women with PCOS\textsuperscript{13}, which
+has been widely cited and reinforced in both academic and clinical
+contexts. However, as we have noted, this study along with
+others\textsuperscript{35--37} have numerous reporting errors which led
+us to drop them from our present analysis (though sensitivity analysis
+including them did not alter our overall conclusions). The mistaken
+belief that REE is lower for those with PCOS may have mistakenly lead to
 recommendations centred on slightly more severe calorie restriction to
@@
 regulation in PCOS given small group-level differences between women
-with PCOS and those without (229 kcal/day {[}95\%QI:169 kcal/day to 303
-kcal/day{]} versus 238 kcal/day {[}95\%QI:178 kcal/day to 312
+with PCOS and those without (1472 kcal/day {[}95\%QI:1359 kcal/day to
+1587 kcal/day{]} versus 1442 kcal/day {[}95\%QI:1334 kcal/day to 1553
 kcal/day{]}, respectively). If anything, REE may be slightly higher in
@@
 issues within adipose tissue and muscle tissue contributing to
-complications like obesity and insulin resistance (PMID: 32244780).
+complications like obesity and insulin resistance\textsuperscript{63}.
 Taken together, these findings highlight that factors beyond REE and
@@
 inadvertently further foster the already well documented disordered
-eating in this population\textsuperscript{21,22,63,64}. The pathways
+eating in this population\textsuperscript{21,22,64,65}. The pathways
 linking PCOS and disordered eating are multifactorial. Biological
@@
 ghrelin and leptin signalling can heighten hunger, carbohydrate
-cravings, and appetite variability\textsuperscript{65}. Frequent
+cravings, and appetite variability\textsuperscript{66}. Frequent
 hypoglycaemia and associated mood changes have also been observed, which
-can trigger compensatory eating or binge episodes\textsuperscript{65}.
+can trigger compensatory eating or binge episodes\textsuperscript{66}.
 These physiological processes interact with psychological and social
@@
 which together compound vulnerability to disordered
-eating\textsuperscript{66}. Moreover, eating disorders themselves can
+eating\textsuperscript{67}. Moreover, eating disorders themselves can
 disrupt endocrine function, potentially worsening PCOS symptoms and
-creating a self-reinforcing cycle\textsuperscript{64,67}.
+creating a self-reinforcing cycle\textsuperscript{65,68}.
 
@@
 dietary advice is appropriate given it may exacerbate feelings of
-failure, hunger dysregulation, and shame\textsuperscript{68}. These
+failure, hunger dysregulation, and shame\textsuperscript{69}. These
 concerns, coupled with the lack of difference in REE between women with
@@
 
-Hannah Cabré was supported by the National Institute of Diabetes and
-Digestive and Kidney Diseases of the National Institutes of Health under
-Award Number T32DK064584. James Steele provides research consultancy
-through his company Steele Research Limited, is contracted currently by
-MacroFactor and Kieser Australia through Steele Research Limited, and
-has also received travel expenses and honoraria for speaking from fit20
-International, Exercise School Portugal, and Discover Strength.
+Richard Kirwan provides nutrition consultancy through his company Be
+More Nutrition. Leigh Peele is employed by MacroFactor, a
+nutrition-tracking app. Gregory Nuckols is co-owner of MacroFactor.
+Georgia Kohlhoff provides nutrition consultancy through her company
+Flourishing Health. Hannah Cabré was supported by the National Institute
+of Diabetes and Digestive and Kidney Diseases of the National Institutes
+of Health under Award Number T32DK064584. Alyssa Olenick provides
+fitness coaching and scientific consulting under ALYSSA OLENICK LLC. She
+receives compensation from scientific consulting, speaking engagments,
+affiliated social media partnerships, and book publishing. James Steele
+provides research consultancy through his company Steele Research
+Limited, is contracted currently by MacroFactor and Kieser Australia
+through Steele Research Limited, and has also received travel expenses
+and honoraria for speaking from fit20 International, Exercise School
+Portugal, and Discover Strength.
 
@@
 necessarily represent the official views of the National Institutes of
-Health or any other funding agencies or institutions noted above.
+Health or any other funding agencies or institutions noted above. The
+authors declare no other conflicts of interest related to the submitted
+work.
 
@@
 
+\bibitem[\citeproctext]{ref-sanchez-garridoMetabolicDysfunctionPolycystic2020}
+\CSLLeftMargin{63. }%
+\CSLRightInline{Sanchez-Garrido MA, Tena-Sempere M. Metabolic
+dysfunction in polycystic ovary syndrome: {Pathogenic} role of androgen
+excess and potential therapeutic strategies. \emph{Molecular
+Metabolism}. 2020;35:100937.
+doi:\href{https://doi.org/10.1016/j.molmet.2020.01.001}{10.1016/j.molmet.2020.01.001}}
+
 \bibitem[\citeproctext]{ref-leeIncreasedRiskDisordered2017}
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Lee I, Cooney LG, Saini S, et al. Increased risk of
@@
 \bibitem[\citeproctext]{ref-bernadettPrevalenceEatingDisorders2016}
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Bernadett M, Szemán-N A.
@@
 \bibitem[\citeproctext]{ref-barryImpactEatingBehavior2011}
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Barry JA, Bouloux P, Hardiman PJ. The impact of eating
@@
 \bibitem[\citeproctext]{ref-steegers-theunissenPolycysticOvarySyndrome2020}
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Steegers-Theunissen RPM, Wiegel RE, Jansen PW, Laven
@@
 \bibitem[\citeproctext]{ref-cooneyIncreasedPrevalenceBinge2024}
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Cooney LG, Gyorfi K, Sanneh A, et al. Increased
@@
 \bibitem[\citeproctext]{ref-pirottaDisorderedEatingBehaviours2019}
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Pirotta S, Barillaro M, Brennan L, et al. Disordered

### 2025-12-03 manuscript/submission_manuscript.qmd
Manuscript ready for submission

@@
         \renewcommand\Affilfont{\small}
+        \usepackage{etoolbox}
 geometry:
@@
   - name: Richard Kirwan
-    affiliations:
-      - ref: 1
+    affil-id: 1
     orcid: 0000-0003-4645-0077
@@
   - name: Leigh Peele
-    affiliations:
-      - ref: 2
+    affil-id: 2
     email: leigh@macrofactorapp.com
-  - name: Greg Nuckols
-    affiliations:
-      - ref: 2
+  - name: Gregory Nuckols
+    affil-id: 2
     orcid: 0009-0004-2679-7967
@@
   - name: Georgia Kohlhoff
-    affiliations:
-      - ref: 3
+    affil-id: 3
     email: g_kohlhoff@hotmail.co.uk
   - name: Hannah Cabré
-    affiliations:
-      - ref: 4
+    affil-id: 4
     orcid: 0000-0002-3235-6867
@@
   - name: Alyssa Olenick
-    affiliations:
-      - ref: 5
+    affil-id: 5
     orcid: 0000-0002-2217-8067
@@
   - name: James Steele
-    affiliations:
-      - ref: 6
-      - ref: 2
-      - ref: 7
+    affil-id: 2,6,7
     orcid: 0000-0002-8003-0757
@@
     name: School of Health and Biomedical Sciences, Royal Melbourne Institute of Technology, Melbourne, VIC, Australia
-abstract: |
-  Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age women, who often have higher BMI classification.
-This is assumed to stem from lower resting energy expenditure (REE), influencing lifestyle intervention guidelines.
-However, evidence for reduced REE in women with PCOS compared with those without is inconsistent.
-Objective: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS.
-Data Sources: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.
-Study Selection: Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included.
-Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate.
-Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors to compare both mean REE, and between person variation in REE, between women with and without PCOS.
-Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33].
-Conclusions: These findings indicate that REE does not meaningfully differ between women with and without PCOS.
-Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.
-keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
+thanks: |
+  Corresponding author: James Steele (james@steele-research.com)
 bibliography: 
@@
 ```
+\newpage
 
+## Financial Disclosures/Conflicts of Interest
+Richard Kirwan provides nutrition consultancy through his company Be More Nutrition.
+Leigh Peele is employed by MacroFactor, a nutrition-tracking app.
+Gregory Nuckols is co-owner of MacroFactor.
+Georgia Kohlhoff provides nutrition consultancy through her company Flourishing Health.
+Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.
+Alyssa Olenick provides fitness coaching and scientific consulting under ALYSSA OLENICK LLC.
+She receives compensation from scientific consulting, speaking engagments, affiliated social media partnerships, and book publishing.
+James Steele provides research consultancy through his company Steele Research Limited, is contracted currently by MacroFactor and Kieser Australia through Steele Research Limited, and has also received travel expenses and honoraria for speaking from fit20 International, Exercise School Portugal, and Discover Strength.
+
+The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health or any other funding agencies or institutions noted above.
+The authors declare no other conflicts of interest related to the submitted work.
+
+\newpage
+
+## Abstract
+Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age women, who often have higher BMI classification.
+This is assumed to stem from lower resting energy expenditure (REE), influencing lifestyle intervention guidelines.
+However, evidence for reduced REE in women with PCOS compared with those without is inconsistent.
+Objective: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS.
+Data Sources: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025.
+Study Selection: Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included.
+Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate.
+Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors to compare both mean REE, and between person variation in REE, between women with and without PCOS.
+Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33].
+Conclusions: These findings indicate that REE does not meaningfully differ between women with and without PCOS.
+Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.
+
+Keywords: Polycystic ovary syndrome, Energy Metabolism, Resting Metabolic Rate, Basal Metabolism
+
+\newpage
 # Introduction
@@
 
-![PRISMA 2020 flow diagram template for systematic reviews.
-Note that a “report” could be a journal article, preprint, conference abstract, study register entry, clinical study report, dissertation, unpublished manuscript, government report or any other document providing relevant information.
-The website noted was a prior narrative review on this topic by some of the authors (https://macrofactorapp.com/pcos-bmr/) which identified two studies not found in our systematic database search.](prisma.png){#fig-prisma}
-
 Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`).
@@
 
-```{r}
-#| label: fig-mean-REE 
-#| fig-width: 10
-#| fig-height: 6.25
-#| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.
-
-combined_mean_plot
-```
-
 ## Standard Deviation of REE Model Results
@@
 These results, including the full visualisation of the posterior distribution and the conditional estimates by study, can be seen in @fig-variance-REE.
-
-```{r}
-#| label: fig-variance-REE 
-#| fig-width: 10
-#| fig-height: 6.25
-#| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.
-
-combined_variance_plot
-```
 
@@
 These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism [@@teedeRecommendations2023International2023], predisposing women with PCOS to weight gain.
-This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
+This belief largely stems from a single but influential 2009 study from Georgopoulos et al.
+that reported a significantly reduced BMR in women with PCOS [@georgopoulosBasalMetabolicRate2009], which has been widely cited and reinforced in both academic and clinical contexts.
+However, as we have noted, this study along with others [@kritikouA2BV3Adrenergic2006; @saltamavrosAlpha2Beta2007; @koikaAssociationPro12AlaPolymorphism2009] have numerous reporting errors which led us to drop them from our present analysis (though sensitivity analysis including them did not alter our overall conclusions).
+The mistaken belief that REE is lower for those with PCOS may have mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021].
 Recognising that there may be minimal differences in REE between women with PCOS and those without can inform both clinical and public practices, potentially leading to a shift in focus away from a requirement for more severe caloric restriction as a primary method of treatment towards more comprehensive, individualised, and psychologically safe approaches to care [@teedeRecommendations2023International2023].
 
-The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively).
+The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_mean_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_mean_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_mean_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_mean_effects_preds_condition$.upper[1])` kcal/day], respectively).
 If anything, REE may be slightly higher in women with PCOS compared to healthy women without PCOS.
@@
 
-# Financial Disclosures/Conflicts of Interest
-Hannah Cabré was supported by the National Institute of Diabetes and Digestive and Kidney Diseases of the National Institutes of Health under Award Number T32DK064584.
-James Steele provides research consultancy through his company Steele Research Limited, is contracted currently by MacroFactor and Kieser Australia through Steele Research Limited, and has also received travel expenses and honoraria for speaking from fit20 International, Exercise School Portugal, and Discover Strength.
-
-The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health or any other funding agencies or institutions noted above.
-
 # Data Availability
@@
 # References
+
+::: {#refs}
+:::
+
+\newpage
+
+![PRISMA 2020 flow diagram template for systematic reviews.
+Note that a “report” could be a journal article, preprint, conference abstract, study register entry, clinical study report, dissertation, unpublished manuscript, government report or any other document providing relevant information.
+The website noted was a prior narrative review on this topic by some of the authors (https://macrofactorapp.com/pcos-bmr/) which identified two studies not found in our systematic database search.](prisma.png){#fig-prisma}
+
+```{r}
+#| label: fig-mean-REE 
+#| fig-width: 10
+#| fig-height: 6.25
+#| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.
+
+combined_mean_plot
+```
+
+```{r}
+#| label: fig-variance-REE 
+#| fig-width: 10
+#| fig-height: 6.25
+#| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.
+
+combined_variance_plot
+```

### 2025-12-03 manuscript/submission_manuscript.tex
Manuscript ready for submission

--- 2025-12-03 manuscript/pre_print.tex
@@
 
+\usepackage{lineno}\linenumbers
+\usepackage{setspace}\AtBeginDocument{\spacing{2}}
 \usepackage[font=scriptsize]{caption}
+\usepackage{pdflscape}
 \usepackage[noblocks]{authblk}
@@
 \usepackage{etoolbox}
-\AtEndEnvironment{abstract}{\noindent\textbf{Keywords:} \KeywordList \par}
 \makeatletter
@@
   pdfauthor={Richard Kirwan; Leigh Peele; Gregory Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
-  pdfkeywords={Polycystic ovary syndrome, Energy Metabolism, Resting
-Metabolic Rate, Basal Metabolism},
   colorlinks=true,
@@
 \title{Resting energy expenditure of women with and without polycystic
-ovary syndrome: a systematic review and meta-analysis\thanks{Preprint,
-please cite as: Kirwan, R., Peele, L., Nuckols, G., Kohlhoff, G., Cabré,
-H., Olenick, A., and Steele, J. (2025). Resting energy expenditure of
-women with and without polycystic ovary syndrome: a systematic review
-and meta-analysis. medrxiv DOI: TO ADD. Address for correspondence:
-james@steele-research.com}}
+ovary syndrome: a systematic review and
+meta-analysis\thanks{Corresponding author: James Steele
+(james@steele-research.com)}}
 
@@
 % Store the keywords (expanded by Pandoc here)
-\def\KeywordList{Polycystic ovary syndrome, Energy Metabolism, Resting
-Metabolic Rate, Basal Metabolism}
+\def\KeywordList{}
 
@@
 \maketitle
-\begin{abstract}
+
+\newpage
+
+\subsection{Financial Disclosures/Conflicts of
+Interest}\label{financial-disclosuresconflicts-of-interest}
+
+Richard Kirwan provides nutrition consultancy through his company Be
+More Nutrition. Leigh Peele is employed by MacroFactor, a
+nutrition-tracking app. Gregory Nuckols is co-owner of MacroFactor.
+Georgia Kohlhoff provides nutrition consultancy through her company
+Flourishing Health. Hannah Cabré was supported by the National Institute
+of Diabetes and Digestive and Kidney Diseases of the National Institutes
+of Health under Award Number T32DK064584. Alyssa Olenick provides
+fitness coaching and scientific consulting under ALYSSA OLENICK LLC. She
+receives compensation from scientific consulting, speaking engagments,
+affiliated social media partnerships, and book publishing. James Steele
+provides research consultancy through his company Steele Research
+Limited, is contracted currently by MacroFactor and Kieser Australia
+through Steele Research Limited, and has also received travel expenses
+and honoraria for speaking from fit20 International, Exercise School
+Portugal, and Discover Strength.
+
+The content is solely the responsibility of the authors and does not
+necessarily represent the official views of the National Institutes of
+Health or any other funding agencies or institutions noted above. The
+authors declare no other conflicts of interest related to the submitted
+work.
+
+\newpage
+
+\subsection{Abstract}\label{abstract}
+
 Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age
@@
 not physiologically relevant.
-\end{abstract}
+
+Keywords: Polycystic ovary syndrome, Energy Metabolism, Resting
+Metabolic Rate, Basal Metabolism
+
+\newpage
 
@@
 
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{prisma.png}}
-
-}
-
-\caption{\label{fig-prisma}PRISMA 2020 flow diagram template for
-systematic reviews. Note that a ``report'' could be a journal article,
-preprint, conference abstract, study register entry, clinical study
-report, dissertation, unpublished manuscript, government report or any
-other document providing relevant information. The website noted was a
-prior narrative review on this topic by some of the authors
-(https://macrofactorapp.com/pcos-bmr/) which identified two studies not
-found in our systematic database search.}
-
-\end{figure}%
-
 Our systematic review identified 13 studies from 12 lab/research groups
@@
 
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{pre_print_files/figure-pdf/fig-mean-REE-1.pdf}}
-
-}
-
-\caption{\label{fig-mean-REE}Posterior distribution, mean point
-estimates, and 95\% quantile intervals for conditional estimates by
-study, global grand mean estimates by condition, and the contrast
-between conditions for mean resting energy expenditure of control women
-without PCOS and women with PCOS.}
-
-\end{figure}%
-
 \subsection{Standard Deviation of REE Model
@@
 estimates by study, can be seen in Figure~\ref{fig-variance-REE}.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{pre_print_files/figure-pdf/fig-variance-REE-1.pdf}}
-
-}
-
-\caption{\label{fig-variance-REE}Posterior distribution, variance point
-estimates, and 95\% quantile intervals for conditional estimates by
-study, global grand variance estimates by condition, and the contrast
-ratio between conditions for the standard deviation of resting energy
-expenditure of control women without PCOS and women with PCOS.}
-
-\end{figure}%
 
@@
 
-\section{Financial Disclosures/Conflicts of
-Interest}\label{financial-disclosuresconflicts-of-interest}
-
-Richard Kirwan provides nutrition consultancy through his company Be
-More Nutrition. Leigh Peele is employed by MacroFactor, a
-nutrition-tracking app. Gregory Nuckols is co-owner of MacroFactor.
-Georgia Kohlhoff provides nutrition consultancy through her company
-Flourishing Health. Hannah Cabré was supported by the National Institute
-of Diabetes and Digestive and Kidney Diseases of the National Institutes
-of Health under Award Number T32DK064584. Alyssa Olenick provides
-fitness coaching and scientific consulting under ALYSSA OLENICK LLC. She
-receives compensation from scientific consulting, speaking engagments,
-affiliated social media partnerships, and book publishing. James Steele
-provides research consultancy through his company Steele Research
-Limited, is contracted currently by MacroFactor and Kieser Australia
-through Steele Research Limited, and has also received travel expenses
-and honoraria for speaking from fit20 International, Exercise School
-Portugal, and Discover Strength.
-
-The content is solely the responsibility of the authors and does not
-necessarily represent the official views of the National Institutes of
-Health or any other funding agencies or institutions noted above. The
-authors declare no other conflicts of interest related to the submitted
-work.
-
 \section{Data Availability}\label{data-availability}
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}
 
@@
 
+\newpage
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{prisma.png}}
+
+}
+
+\caption{\label{fig-prisma}PRISMA 2020 flow diagram template for
+systematic reviews. Note that a ``report'' could be a journal article,
+preprint, conference abstract, study register entry, clinical study
+report, dissertation, unpublished manuscript, government report or any
+other document providing relevant information. The website noted was a
+prior narrative review on this topic by some of the authors
+(https://macrofactorapp.com/pcos-bmr/) which identified two studies not
+found in our systematic database search.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{submission_manuscript_files/figure-pdf/fig-mean-REE-1.pdf}}
+
+}
+
+\caption{\label{fig-mean-REE}Posterior distribution, mean point
+estimates, and 95\% quantile intervals for conditional estimates by
+study, global grand mean estimates by condition, and the contrast
+between conditions for mean resting energy expenditure of control women
+without PCOS and women with PCOS.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{submission_manuscript_files/figure-pdf/fig-variance-REE-1.pdf}}
+
+}
+
+\caption{\label{fig-variance-REE}Posterior distribution, variance point
+estimates, and 95\% quantile intervals for conditional estimates by
+study, global grand variance estimates by condition, and the contrast
+ratio between conditions for the standard deviation of resting energy
+expenditure of control women without PCOS and women with PCOS.}
+
+\end{figure}%
+
 \end{document}

### 2025-12-03 manuscript/pre_print.qmd
Add contributions to preprint and submission

@@
 
+# Contributions
+Gregory Nuckols and Leigh Peele conceived the idea for the project.
+All authors contributed to the design of the project and methods.
+Richie Kirwan and Leigh Peele conducted the systematic search and screening.
+James Steele performed the data extraction, conducted the statistical analyses, and produced the data visualisations.
+All authors contributed to drafting the initial manuscript.
+All authors contributed to editing the manuscript.
+All authors read and approved the final manuscript.
+
 # References

### 2025-12-03 manuscript/pre_print.tex
Add contributions to preprint and submission

@@
 
+\section{Contributions}\label{contributions}
+
+Gregory Nuckols and Leigh Peele conceived the idea for the project. All
+authors contributed to the design of the project and methods. Richie
+Kirwan and Leigh Peele conducted the systematic search and screening.
+James Steele performed the data extraction, conducted the statistical
+analyses, and produced the data visualisations. All authors contributed
+to drafting the initial manuscript. All authors contributed to editing
+the manuscript. All authors read and approved the final manuscript.
+
 \section*{References}\label{references}

### 2025-12-03 manuscript/submission_manuscript.qmd
Add contributions to preprint and submission

@@
 thanks: |
-  Corresponding author: James Steele (james@steele-research.com)
+  Corresponding author: James Steele (Email: james@steele-research.com; ORCID: 0000-0002-8003-0757)
 bibliography: 

### 2025-12-03 manuscript/submission_manuscript.tex
Add contributions to preprint and submission

@@
 ovary syndrome: a systematic review and
-meta-analysis\thanks{Corresponding author: James Steele
-(james@steele-research.com)}}
+meta-analysis\thanks{Corresponding author: James Steele (Email:
+james@steele-research.com; ORCID: 0000-0002-8003-0757)}}
 
@@
 
+\section{Contributions}\label{contributions}
+
+Gregory Nuckols and Leigh Peele conceived the idea for the project. All
+authors contributed to the design of the project and methods. Richie
+Kirwan and Leigh Peele conducted the systematic search and screening.
+James Steele performed the data extraction, conducted the statistical
+analyses, and produced the data visualisations. All authors contributed
+to drafting the initial manuscript. All authors contributed to editing
+the manuscript. All authors read and approved the final manuscript.
+
 \section{References}\label{references}

### 2025-12-03 manuscript/pre_print.qmd
Add Nguo et al (2025) reference

@@
 
+A recent scoping review has highlighted the inconsistency in research examining energy expenditure in women with PCOS [@nguoExploringPhysiologicalFactors2025].
+However, to our knowledge there has been no attempts to systematically review, and quantitatively synthesise via meta-analysis, studies of REE in women with, and without, PCOS.
 Therefore, to estimate and describe the magnitude of difference in REE between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.

### 2025-12-03 manuscript/pre_print.tex
Add Nguo et al (2025) reference

@@
 
-Therefore, to estimate and describe the magnitude of difference in REE
-between women with and without PCOS, we completed a systematic review
-and meta-analysis of studies reporting REE in these populations.
+A recent scoping review has highlighted the inconsistency in research
+examining energy expenditure in women with PCOS\textsuperscript{30}.
+However, to our knowledge there has been no attempts to systematically
+review, and quantitatively synthesise via meta-analysis, studies of REE
+in women with, and without, PCOS. Therefore, to estimate and describe
+the magnitude of difference in REE between women with and without PCOS,
+we completed a systematic review and meta-analysis of studies reporting
+REE in these populations.
 
@@
 with the Preferred Reporting Items for Systematic Reviews and
-Meta-Analyses (PRISMA) statement guidelines\textsuperscript{30}. The
+Meta-Analyses (PRISMA) statement guidelines\textsuperscript{31}. The
 PRISMA flow diagram reported below (Figure~\ref{fig-prisma}) was
 produced using the \texttt{PRISMA2020} R package and Shiny
-app\textsuperscript{31}. The primary aim of this review was to examine
+app\textsuperscript{32}. The primary aim of this review was to examine
 the descriptive question ``Does resting energy expenditure (REE) differ
@@
 information on a particular population in a conservative
-manner\textsuperscript{32}.
+manner\textsuperscript{33}.
 
@@
 interquartile range, median, and sample size depending on the available
-information using the methods of Wan et al.\textsuperscript{33}.
+information using the methods of Wan et al.\textsuperscript{34}.
 Further, where missing, height/body mass/BMI where estimated based on
@@
 and all REE values were converted to kcal/day. In one
-case\textsuperscript{34} REE was reported relative to body mass and the
+case\textsuperscript{35} REE was reported relative to body mass and the
 unadjusted values were no longer available (confirmed by the authors).
@@
 During data extraction it was noted that several studies from the same
-lab/research group\textsuperscript{13,35--37} contained a number of
+lab/research group\textsuperscript{13,36--38} contained a number of
 discrepancies that seemed to be possible reporting errors. This
@@
 reported for the healthy control women without PCOS in three of the
-studies\textsuperscript{35--37}, and REE was reported as an ``adjusted''
+studies\textsuperscript{36--38}, and REE was reported as an ``adjusted''
 value whereby
@@
 software and packages used in the analysis pipeline using the
-\texttt{grateful} package\textsuperscript{38} which can be seen here:
+\texttt{grateful} package\textsuperscript{39} which can be seen here:
 \url{https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf}.
@@
 Given our research question our analysis was aimed at parameter
-estimation\textsuperscript{39} within a Bayesian meta-analytic
-framework\textsuperscript{40}. For all analyses model parameter
+estimation\textsuperscript{40} within a Bayesian meta-analytic
+framework\textsuperscript{41}. For all analyses model parameter
 estimates and their precision, along with conclusions based upon them,
@@
 quality, plausibility of effect, and previous literature, all within the
-context of each model. The \texttt{renv} package\textsuperscript{41} was
+context of each model. The \texttt{renv} package\textsuperscript{42} was
 used for package version reproducibility and a function based analysis
-pipeline using the \texttt{targets} package\textsuperscript{42} was
+pipeline using the \texttt{targets} package\textsuperscript{43} was
 employed (the analysis pipeline can be viewed by downloading the R
@@
 \texttt{metafor} packages \texttt{escalc()}
-function\textsuperscript{43}. The main package
-\texttt{brms}\textsuperscript{44} was used in fitting all the Bayesian
+function\textsuperscript{44}. The main package
+\texttt{brms}\textsuperscript{45} was used in fitting all the Bayesian
 meta-analysis models. Prior and posterior draws were taken using
-\texttt{marginaleffects}\textsuperscript{45} and
-\texttt{tidybayes}\textsuperscript{46} packages. All visualisations are
-created using \texttt{ggplot2}\textsuperscript{47}, \texttt{tidybayes},
-and the \texttt{patchwork}\textsuperscript{48} packages.
+\texttt{marginaleffects}\textsuperscript{46} and
+\texttt{tidybayes}\textsuperscript{47} packages. All visualisations are
+created using \texttt{ggplot2}\textsuperscript{48}, \texttt{tidybayes},
+and the \texttt{patchwork}\textsuperscript{49} packages.
 
@@
 type model given that the studies included had arms of women with PCOS
-both with, and without, a non-PCOS control arm\textsuperscript{49}, and
+both with, and without, a non-PCOS control arm\textsuperscript{50}, and
 also in some cases multiple observations of REE in the different arms
@@
 historical information about a particular population in a conservative
-manner in meta-analyses\textsuperscript{32}. From this model we focus on
+manner in meta-analyses\textsuperscript{33}. From this model we focus on
 reporting the global grand mean estimate for the fixed between condition
@@
 meta-analysis of the mean REEs for women from two large studies of
-healthy people\textsuperscript{50,51} though set with a conservative
+healthy people\textsuperscript{51,52} though set with a conservative
 degrees of freedom for the \(\textrm{Student-}t\) distribution.The
@@
 again set based on meta-analysis of the standard deviation of REEs for
-women from two large studies of healthy people\textsuperscript{50,51}
+women from two large studies of healthy people\textsuperscript{51,52}
 though set with a conservative degrees of freedom for the
@@
 of the standard deviation conditioned upon the natural logarithm of the
-mean\textsuperscript{52} we set this prior to reflect the range of
+mean\textsuperscript{53} we set this prior to reflect the range of
 differences on the on the log scale (i.e., \(\log(1584)\)). We then set
@@
 of four studies with possible reporting errors we could not
-resolve\textsuperscript{13,35--37}. Thus the main models described above
+resolve\textsuperscript{13,36--38}. Thus the main models described above
 were run with and without these studies, the main results are presented
@@
 with interventions in women with PCOS reporting REE at multiple
-timepoints such as mid- and post-intervention\textsuperscript{34,53,54}
+timepoints such as mid- and post-intervention\textsuperscript{35,54,55}
 but lacking of control women without PCOS, we opted to also conduct
@@
 calculated for body mass of women with PCOS in Saltamavros et
-al.\textsuperscript{36}. The PRISMA flow diagram
+al.\textsuperscript{37}. The PRISMA flow diagram
 (Figure~\ref{fig-prisma}) also includes all studies identified including
@@
 (or a modified version there of) for diagnosing
-PCOS\textsuperscript{16,17,34,53--60}. One study used the 1990 National
-Institutes of Health criteria\textsuperscript{61}, and two
-studies\textsuperscript{18,55} diagnosed PCOS via the presence of
+PCOS\textsuperscript{16,17,35,54--61}. One study used the 1990 National
+Institutes of Health criteria\textsuperscript{62}, and two
+studies\textsuperscript{18,56} diagnosed PCOS via the presence of
 oligomenorrhea or amenorrhea alongside additional criteria including
 plasma androgen levels, hirsutism or polycystic ovaries on ultrasound
-scanning. All but one study\textsuperscript{61}, which used doubly
+scanning. All but one study\textsuperscript{62}, which used doubly
 labelled water, measured REE using indirect calorimetry. The specific
@@
 For mean REE the model including the studies noted with possible
-reporting issues that were not resolved\textsuperscript{13,35--37} still
+reporting issues that were not resolved\textsuperscript{13,36--38} still
 resulted in qualitatively similar inferences suggesting little
@@
 For mean REE the model which included only baseline REE measurements
-from studies involving interventions\textsuperscript{34,53,54}, in
+from studies involving interventions\textsuperscript{35,54,55}, in
 addition to other cross-sectional studies, also resulted in
@@
 contexts. However, as we have noted, this study along with
-others\textsuperscript{35--37} have numerous reporting errors which led
+others\textsuperscript{36--38} have numerous reporting errors which led
 us to drop them from our present analysis (though sensitivity analysis
@@
 those without. However, one of the studies included in our
-analysis\textsuperscript{62} reported that, whilst there was little
+analysis\textsuperscript{63} reported that, whilst there was little
 difference in unadjusted REE, when REE was adjusted for fat-free mass it
@@
 issues within adipose tissue and muscle tissue contributing to
-complications like obesity and insulin resistance\textsuperscript{63}.
+complications like obesity and insulin resistance\textsuperscript{64}.
 Taken together, these findings highlight that factors beyond REE and
@@
 inadvertently further foster the already well documented disordered
-eating in this population\textsuperscript{21,22,64,65}. The pathways
+eating in this population\textsuperscript{21,22,65,66}. The pathways
 linking PCOS and disordered eating are multifactorial. Biological
@@
 ghrelin and leptin signalling can heighten hunger, carbohydrate
-cravings, and appetite variability\textsuperscript{66}. Frequent
+cravings, and appetite variability\textsuperscript{67}. Frequent
 hypoglycaemia and associated mood changes have also been observed, which
-can trigger compensatory eating or binge episodes\textsuperscript{66}.
+can trigger compensatory eating or binge episodes\textsuperscript{67}.
 These physiological processes interact with psychological and social
@@
 which together compound vulnerability to disordered
-eating\textsuperscript{67}. Moreover, eating disorders themselves can
+eating\textsuperscript{68}. Moreover, eating disorders themselves can
 disrupt endocrine function, potentially worsening PCOS symptoms and
-creating a self-reinforcing cycle\textsuperscript{65,68}.
+creating a self-reinforcing cycle\textsuperscript{66,69}.
 
@@
 dietary advice is appropriate given it may exacerbate feelings of
-failure, hunger dysregulation, and shame\textsuperscript{69}. These
+failure, hunger dysregulation, and shame\textsuperscript{70}. These
 concerns, coupled with the lack of difference in REE between women with
@@
 strength. Additionally, some studies reported data inconsistencies that
-could not be clarified (e.g.,\textsuperscript{13,35--37}) and were
+could not be clarified (e.g.,\textsuperscript{13,36--38}) and were
 dropped from our main analysis though our conclusions again did not
@@
 
+\bibitem[\citeproctext]{ref-nguoExploringPhysiologicalFactors2025}
+\CSLLeftMargin{30. }%
+\CSLRightInline{Nguo K, McGowan M, Cowan S, et al. Exploring the
+physiological factors relating to energy balance in women with
+polycystic ovary syndrome: A scoping review. \emph{Nutrition Reviews}.
+2025;83(1):160-174.
+doi:\href{https://doi.org/10.1093/nutrit/nuad169}{10.1093/nutrit/nuad169}}
+
 \bibitem[\citeproctext]{ref-pagePRISMA2020Statement2021a}
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Page MJ, McKenzie JE, Bossuyt PM, et al. The {PRISMA}
@@
 \bibitem[\citeproctext]{ref-haddawayPRISMA2020PackageShiny2022}
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Haddaway NR, Page MJ, Pritchard CC, McGuinness LA.
@@
 \bibitem[\citeproctext]{ref-weberApplyingMetaAnalyticPredictivePriors2021}
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Weber S, Li Y, Iii JWS, Kakizume T, Schmidli H. Applying
@@
 \bibitem[\citeproctext]{ref-wanEstimatingSampleMean2014}
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Wan X, Wang W, Liu J, Tong T. Estimating the sample mean
@@
 \bibitem[\citeproctext]{ref-pohlmeierEffectLowstarchLowdairy2014}
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Pohlmeier AM, Phy JL, Watkins P, et al. Effect of a
@@
 \bibitem[\citeproctext]{ref-kritikouA2BV3Adrenergic2006}
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Kritikou S, Saltamavros AD, Adonakis G, et al. The
@@
 \bibitem[\citeproctext]{ref-saltamavrosAlpha2Beta2007}
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Saltamavros AD, Adonakis G, Kritikou S, et al. Alpha 2
@@
 \bibitem[\citeproctext]{ref-koikaAssociationPro12AlaPolymorphism2009}
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Koika V, Marioli DJ, Saltamavros AD, et al. Association
@@
 \bibitem[\citeproctext]{ref-rodriguez-sanchezGratefulFacilitateCitation2023}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Rodriguez-Sanchez F, cre, cph, Jackson CP, Hutchins SD,
@@
 \bibitem[\citeproctext]{ref-cummingNewStatisticsWhy2014}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Cumming G. The {New Statistics}: {Why} and {How}.
@@
 \bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Kruschke JK, Liddell TM. The {Bayesian New Statistics}:
@@
 \bibitem[\citeproctext]{ref-usheyRenvProjectEnvironments2023}
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Ushey K, cre, Wickham H, Software P, PBC. Renv: {Project
@@
 \bibitem[\citeproctext]{ref-landauTargetsDynamicFunctionOriented2023}
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Landau WM, Warkentin MT, Edmondson M, Oliver S, Mahr T,
@@
 \bibitem[\citeproctext]{ref-viechtbauerMetaforMetaAnalysisPackage2023}
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Viechtbauer W. Metafor: {Meta-Analysis Package} for {R}.
@@
 \bibitem[\citeproctext]{ref-burknerBrmsBayesianRegression2023}
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Bürkner PC, Gabry J, Weber S, et al. Brms: {Bayesian
@@
 \bibitem[\citeproctext]{ref-arel-bundockMarginaleffectsPredictionsComparisons2023}
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Arel-Bundock V, cre, cph, Diniz MA, Greifer N, Bacher E.
@@
 \bibitem[\citeproctext]{ref-kayTidybayesTidyData2023}
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Kay M, Mastny T. Tidybayes: {Tidy Data} and '{Geoms}'
@@
 \bibitem[\citeproctext]{ref-wickhamGgplot2CreateElegant2023}
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Wickham H, Chang W, Henry L, et al. Ggplot2: {Create
@@
 \bibitem[\citeproctext]{ref-pedersenPatchworkComposerPlots2023}
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Pedersen TL. Patchwork: {The Composer} of {Plots}.
@@
 \bibitem[\citeproctext]{ref-hongBayesianMissingData2016}
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Hong H, Chu H, Zhang J, Carlin BP. A {Bayesian} missing
@@
 \bibitem[\citeproctext]{ref-velasquezUseAmmoniaInhalants2011}
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Velasquez JR. The {Use} of {Ammonia Inhalants Among
@@
 \bibitem[\citeproctext]{ref-pavlidouRevisedHarrisBenedict2023}
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Pavlidou E, Papadopoulou SK, Seroglou K, Giaginis C.
@@
 \bibitem[\citeproctext]{ref-steeleMetaanalysisVariationSport2023a}
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Steele J, Fisher, Smith, Schoenfeld, Yang, and Nakagawa
@@
 \bibitem[\citeproctext]{ref-moranShorttermMealReplacements2006}
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Moran LJ, Noakes M, Clifton PM, Wittert GA, Williams G,
@@
 \bibitem[\citeproctext]{ref-brunerEffectsExerciseNutritional2006}
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Bruner B, Chad K, Chizen D. Effects of exercise and
@@
 \bibitem[\citeproctext]{ref-robinsonPostprandialThermogenesisReduced1992}
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Robinson S, Chan SP, Spacey S, Anyaoku V, Johnston DG,
@@
 \bibitem[\citeproctext]{ref-cosarRestingMetabolicRate2008}
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Cosar E, Köken G, Sahin FK, et al. Resting metabolic
@@
 \bibitem[\citeproctext]{ref-graffDietaryGlycemicIndex2013}
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Graff SK, Mário FM, Alves BC, Spritzer PM. Dietary
@@
 \bibitem[\citeproctext]{ref-dohRelationshipAdiposityInsulin2016}
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Doh E, Mbanya A, Kemfang-Ngowa JD, et al. The
@@
 \bibitem[\citeproctext]{ref-rodriguesLowValidityPredictive2018}
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Rodrigues AM dos S, Costa ABP, Campos DL, et al. Low
@@
 \bibitem[\citeproctext]{ref-tosiInsulinMediatedSubstrateUse2021}
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Tosi F, Villani M, Migazzi M, et al. Insulin-{Mediated
@@
 \bibitem[\citeproctext]{ref-broskeyAssessingEnergyRequirements2017}
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Broskey NT, Klempel MC, Gilmore LA, et al. Assessing
@@
 \bibitem[\citeproctext]{ref-tosiRestingEnergyExpenditure2024}
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Tosi F, Rosmini F, Gremes V, et al. Resting energy
@@
 \bibitem[\citeproctext]{ref-sanchez-garridoMetabolicDysfunctionPolycystic2020}
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Sanchez-Garrido MA, Tena-Sempere M. Metabolic
@@
 \bibitem[\citeproctext]{ref-leeIncreasedRiskDisordered2017}
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Lee I, Cooney LG, Saini S, et al. Increased risk of
@@
 \bibitem[\citeproctext]{ref-bernadettPrevalenceEatingDisorders2016}
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Bernadett M, Szemán-N A.
@@
 \bibitem[\citeproctext]{ref-barryImpactEatingBehavior2011}
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Barry JA, Bouloux P, Hardiman PJ. The impact of eating
@@
 \bibitem[\citeproctext]{ref-steegers-theunissenPolycysticOvarySyndrome2020}
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Steegers-Theunissen RPM, Wiegel RE, Jansen PW, Laven
@@
 \bibitem[\citeproctext]{ref-cooneyIncreasedPrevalenceBinge2024}
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Cooney LG, Gyorfi K, Sanneh A, et al. Increased
@@
 \bibitem[\citeproctext]{ref-pirottaDisorderedEatingBehaviours2019}
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Pirotta S, Barillaro M, Brennan L, et al. Disordered
